---
title: "targetscan mir-146"
author: "Skim"
date: '2020 6 1 '
output: 
  html_document:
    df_print: paged
    toc: true
    toc_depth: 4
    toc_float: false
    code_folding: hide
---

This report is to check target genes of mir-146a. We expected to see downregulation of mir-146a target genes in ADp40KO compare to AD. However, the analysis results isn't conclusive to say target genes are downregulated. In here, I tested other miRNAs e.g. mir-206, let-7a and mir-1970. And I tested AD vs Ctrl, AD vs ADp40KO and Cluster 8 (DAM) vs 3 (homeostatic). 

```{r warning=FALSE, message=FALSE}
#.libPaths(c("/home/skim/R/usr_lib", .libPaths()))
#.libPaths(c(.libPaths()))
#library(Seurat, lib.loc="/data/rajewsky/shared_libs/R/")
library(Seurat)
library(tidyverse)
library(dplyr)
library(stringr)
library(ggplot2)
library(gridExtra)
library(xlsx)
library(readxl)
```

Tested miRNAs

**mir-146-5p** (broadly conserved microRNA family), **mir-146a-3p** (other miRBase annotations)

Other control: mir-206, let-7a, mir-1970

**miR-1-3p/206-3p** (broadly conserved), **miR-206-5p**(very poorly conserved)

**let-7-5p/miR-98-5p** (broadly conserved), **let-7-3p/miR-98-3p** (very poorly conserved), **let-7a-2-3p** (very poorly conserved)

**mir-1970** (other miRBase annotations)

**miR-206** was shown to enhance LPS-induced inflammation and cause the release of amyloid-β (an essential protein in AD) in microglia [ref](https://www.frontiersin.org/articles/10.3389/fnmol.2019.00125/full)

**Let-7a** is involved in maintaining microglial function in inflammation-mediated injury. During inflammation, Let-7a was shown to inhibit the production of proinflammatory mediators [ref](https://www.frontiersin.org/articles/10.3389/fnmol.2019.00125/full)

**mir-1970** is targeting Il12b

```{r warning=FALSE, message=FALSE, fig.width= 5.5, fig.height=5}
load(file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/miRNA_lncRNA/R_Scripts/20200505_pri_miRNA_dgCMatrix_Create_Seurat_with_intergenic_pri_miRNA.obj")
data9set.SO$cell_type <- data9set.SO@active.ident


mir_146_5p <- read_xlsx("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/miRNA_lncRNA/Targetscan/TargetScan7.2__miR-146-5p.predicted_targets.xlsx")

mir_146a_3p <- read_xlsx("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/miRNA_lncRNA/Targetscan/TargetScan7.2__miR-146a-3p.predicted_targets.xlsx")

#mir_146b_3p <- read_xlsx("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/miRNA_lncRNA/Targetscan/TargetScan7.2__miR-146b-3p.predicted_targets.xlsx")


MG_DE_AD_Ctrl <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/DE_csv_files/Cell_type/20200221_Cell_type_DE_by_MAST_MicrogliaAD_Ctrl.csv")

MG_DE_AD_ADp40KO <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/DE_csv_files/Cell_type/20200221_Cell_type_DE_by_MAST_MicrogliaAD_ADp40KO.csv")

MG.marker8vs3 <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200109_MG_sub_analysis.MG_marker.csv")
```



```{r warning=FALSE, message=FALSE, fig.width= 5.5, fig.height=5}
cumulative_plot <- function(DF, is.target, target, labels){
  ggplot(DF, aes(x = avg_logFC, color = is.target)) +
  stat_ecdf() +
  theme_classic() +
  ylab("Cumulative fraction") +
  xlim(-0.5, 0.5) +
  scale_color_manual(values = c("black", "red"),
                     breaks = c(F, T),
                     labels = c(paste0("nontargets (", sum(!DF$is.target), ")"),
                                paste0(target," targets (", sum(DF$is.target), ")")),
                     name   = "") +
  theme(axis.line.y=element_line(size=.5),
        axis.line.x=element_line(size=.5),
        axis.text.x=element_text(size=15,  family="Helvetica", color="black"),
        axis.text.y=element_text(size=15,  family="Helvetica", color="black"),
        axis.title.x=element_text(size=15, family="Helvetica", color="black"),
        axis.title.y=element_text(size=15, family="Helvetica", color="black"),
        legend.text = element_text(size=10, family="Helvetica", color="black"),
        legend.position = c(0.01, 1),
        legend.justification = c(0, 1),
        aspect.ratio=1,
        plot.title=element_text(size=15, face="bold", family="Helvetica", hjust=0.5)) +
  annotate("text", x = 0.2, y = 0.5, label = labels, colour = "#003366", face = "bold", size = 4, family="Helvetica")
}
```

### 1. mir-146a AD vs WT

mir-146-5p Filter condition: Cumulative weighted context++ score less than -0.2

mir-146a-3p Filter condition: Cumulative weighted context++ score less than -0.3


```{r warning=FALSE, message=FALSE, fig.width= 11, fig.height=10}
#################
# 1. mir-146-5p
#################

MG_DE_AD_Ctrl$is.target <- MG_DE_AD_Ctrl$X %in% mir_146_5p$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG_DE_AD_Ctrl, alt = "two.sided")
c1 <- cumulative_plot(MG_DE_AD_Ctrl, is.target = is.target,"mir_146_5p", "p-value = 0.2561") + ggtitle("AD vs Ctrl in Microglia mir-146-5p")

MG_DE_AD_Ctrl$is.target <- MG_DE_AD_Ctrl$X %in% mir_146_5p[mir_146_5p$`Cumulative weighted context++ score` < -.19, ]$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG_DE_AD_Ctrl, alt = "two.sided")
c2 <- cumulative_plot(MG_DE_AD_Ctrl, is.target = is.target,"mir_146_5p filtered", "p-value = 0.9692") + ggtitle("AD vs Ctrl in Microglia mir-146-5p filtered")

#################
# 2. mir-146a-3p
#################
MG_DE_AD_Ctrl$is.target <- MG_DE_AD_Ctrl$X %in% mir_146a_3p$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG_DE_AD_Ctrl, alt = "two.sided") #0.0002547
c3 <- cumulative_plot(MG_DE_AD_Ctrl, is.target = is.target,"mir_146a_3p", "p-value = 0.0002547") + ggtitle("AD vs Ctrl in Microglia mir-146a-3p")

MG_DE_AD_Ctrl$is.target <- MG_DE_AD_Ctrl$X %in% mir_146a_3p[mir_146a_3p$`Cumulative weighted context++ score` < -.3,]$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG_DE_AD_Ctrl, alt = "two.sided") 0.1037
c4 <- cumulative_plot(MG_DE_AD_Ctrl, is.target = is.target,"mir_146a_3p filtered", "p-value = 0.1037") + ggtitle("AD vs Ctrl in Microglia mir-146a-3p filtered")
grid.arrange(c1, c2, c3, c4, ncol = 2)
```

### 2.  mir-146a AD vs ADp40KO

```{r warning=FALSE, message=FALSE, fig.width= 11, fig.height=10}
#################
# 1. mir-146-5p
#################
MG_DE_AD_ADp40KO$is.target <- MG_DE_AD_ADp40KO$X %in% mir_146_5p$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG_DE_AD_ADp40KO, alt = "two.sided") #0.767
c1 <- cumulative_plot(MG_DE_AD_ADp40KO, is.target = is.target,"mir-146-5p", "p-value = 0.767") + ggtitle("AD vs ADp40KO in Microglia mir-146-5p")

MG_DE_AD_ADp40KO$is.target <- MG_DE_AD_ADp40KO$X %in% mir_146_5p[mir_146_5p$`Cumulative weighted context++ score` < -.19, ]$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG_DE_AD_ADp40KO, alt = "two.sided") #0.5508
c2 <- cumulative_plot(MG_DE_AD_ADp40KO, is.target = is.target,"mir-146-5p filtered", "p-value = 0.5508") + ggtitle("AD vs ADp40KO in Microglia mir-146-5p filtered")

#################
# 2. mir-146a-3p
#################
MG_DE_AD_ADp40KO$is.target <- MG_DE_AD_ADp40KO$X %in% mir_146a_3p$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG_DE_AD_ADp40KO, alt = "two.sided") #0.002224
c3 <- cumulative_plot(MG_DE_AD_ADp40KO, is.target = is.target,"mir_146a_3p", "p-value = 0.002224") + ggtitle("AD vs ADp40KO in Microglia mir-146a-3p")

MG_DE_AD_ADp40KO$is.target <- MG_DE_AD_ADp40KO$X %in% mir_146a_3p[mir_146a_3p$`Cumulative weighted context++ score` < -.3,]$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG_DE_AD_ADp40KO, alt = "two.sided") #0.2492
c4 <- cumulative_plot(MG_DE_AD_ADp40KO, is.target = is.target,"mir_146a_3p filtered", "p-value = 0.2492") + ggtitle("AD vs ADp40KO in Microglia mir-146a-3p filtered")
grid.arrange(c1, c2, c3, c4, ncol = 2)

```

### 3.  mir-146a Cluster 8 vs 3

```{r warning=FALSE, message=FALSE, fig.width= 11, fig.height=10}
#################
# 1. mir-146-5p
#################
MG.marker8vs3$is.target <- MG.marker8vs3$X %in% mir_146_5p$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG.marker8vs3, alt = "two.sided")0.02178
c1 <- cumulative_plot(MG.marker8vs3, is.target = is.target,"mir-146-5p", "p-value = 0.02178") + ggtitle("Cluster 8 vs 3 in Microglia mir-146-5p")

MG.marker8vs3$is.target <- MG.marker8vs3$X %in% mir_146_5p[mir_146_5p$`Cumulative weighted context++ score` < -.19, ]$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG.marker8vs3, alt = "two.sided") #0.5685
c2 <- cumulative_plot(MG.marker8vs3, is.target = is.target,"mir-146-5p filtered", "p-value = 0.5685") + ggtitle("Cluster 8 vs 3 in Microglia mir-146-5p filtered")

#################
# 2. mir-146a-3p
#################
MG.marker8vs3$is.target <- MG.marker8vs3$X %in% mir_146a_3p$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG.marker8vs3, alt = "two.sided") #0.2707
c3 <- cumulative_plot(MG.marker8vs3, is.target = is.target,"mir_146a_3p", "p-value = 0.2707") + ggtitle("Cluster 8 vs 3 in Microglia mir-146a-3p")

MG.marker8vs3$is.target <- MG.marker8vs3$X %in% mir_146a_3p[mir_146a_3p$`Cumulative weighted context++ score` < -.3,]$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG.marker8vs3, alt = "two.sided") #0.4797
c4 <- cumulative_plot(MG_DE_AD_ADp40KO, is.target = is.target,"mir_146a_3p filtered", "p-value = 0.4797") + ggtitle("Cluster 8 vs 3 in Microglia mir-146a-3p filtered")
grid.arrange(c1, c2, c3, c4, ncol = 2)


```

### 4. pri-mir-146a feature plots

```{r warning=FALSE, message=FALSE, fig.width= 15, fig.height=5}
data9set_MG.SO <- subset(data9set.SO, subset = seurat_clusters %in% c(3,8))

data9set_MG_Ctrl.SO <- subset(data9set_MG.SO, subset = sample %in% "Ctrl")
data9set_MG_AD.SO <- subset(data9set_MG.SO, subset = sample %in% "AD")
data9set_MG_ADp40KO.SO <- subset(data9set_MG.SO, subset = sample %in% "ADp40KO")

data9set_3.SO <- subset(data9set.SO, subset = seurat_clusters %in% c(3))
data9set_8.SO <- subset(data9set.SO, subset = seurat_clusters %in% c(8))


f1 <- FeaturePlot(data9set_MG_Ctrl.SO, features = "mir-146a", order = TRUE, pt.size = 1) + xlim(c(-7, 7)) + ylim(c(18,28)) + ggtitle("pri-mir-146a (WT)")
f2 <- FeaturePlot(data9set_MG_AD.SO, features = "mir-146a", order = TRUE, pt.size = 1) + xlim(c(-7, 7)) + ylim(c(18,28)) + ggtitle("pri-mir-146a (AD)")
f3 <- FeaturePlot(data9set_MG_ADp40KO.SO, features = "mir-146a", order = TRUE, pt.size = 1) + xlim(c(-7, 7)) + ylim(c(18,28)) + ggtitle("pri-mir-146a (ADp40KO)")

grid.arrange(f1, f2, f3, ncol = 3)
```

```{r warning=FALSE, message=FALSE, fig.width= 5.5, fig.height=5}
# extract UMI counts and calculate average per each gene
MG_Ctrl_count.df <- data.frame(GetAssayData(data9set_MG_Ctrl.SO, slot = "counts"))
MG_AD_count.df <- data.frame(GetAssayData(data9set_MG_AD.SO, slot = "counts"))
MG_ADp40KO_count.df <- data.frame(GetAssayData(data9set_MG_ADp40KO.SO, slot = "counts"))
MG_3_count.df <- data.frame(GetAssayData(data9set_3.SO, slot = "counts"))
MG_8_count.df <- data.frame(GetAssayData(data9set_8.SO, slot = "counts"))


MG_Ctrl_count_cpm.df <- apply(MG_Ctrl_count.df, 2, function(x) (x/sum(x)) * 1000000)
MG_AD_count_cpm.df <- apply(MG_AD_count.df, 2, function(x) (x/sum(x)) * 1000000)
MG_ADp40KO_count_cpm.df <- apply(MG_ADp40KO_count.df, 2, function(x) (x/sum(x)) * 1000000)
MG_3_count_cpm.df <- apply(MG_3_count.df, 2, function(x) (x/sum(x)) * 1000000)
MG_8_count_cpm.df <- apply(MG_8_count.df, 2, function(x) (x/sum(x)) * 1000000)


MG_Ctrl_count_cpm.df <- data.frame(Matrix::rowMeans(MG_Ctrl_count_cpm.df))
MG_AD_count_cpm.df <- data.frame(Matrix::rowMeans(MG_AD_count_cpm.df))
MG_ADp40KO_count_cpm.df <- data.frame(Matrix::rowMeans(MG_ADp40KO_count_cpm.df))
MG_3_count_cpm.df <- data.frame(Matrix::rowMeans(MG_3_count_cpm.df))
MG_8_count_cpm.df <- data.frame(Matrix::rowMeans(MG_8_count_cpm.df))



MG_count_cpm.df <- cbind(MG_Ctrl_count_cpm.df, MG_AD_count_cpm.df,MG_ADp40KO_count_cpm.df, MG_3_count_cpm.df, MG_8_count_cpm.df)
MG_count_cpm.df <- as.data.frame(MG_count_cpm.df)
MG_count_cpm.df$gene <- rownames(MG_count_cpm.df)
# Remove Ttr
MG_count_cpm.df <- MG_count_cpm.df[MG_count_cpm.df$gene != "Ttr", ]
MG_count_cpm.df$gene <- NULL

# remove genes where all zero in samples
MG_count_cpm.df <- MG_count_cpm.df[apply(MG_count_cpm.df == 0, 1, sum) != 5, ]
colnames(MG_count_cpm.df) <- c("WT", "AD", "ADp40KO", "Cluster3", "Cluster8")
# add pseudocount 0.001 and logarithm 
MG_count_cpm.df <- log(MG_count_cpm.df + 1)
MG_count_cpm.df$gene <- rownames(MG_count_cpm.df)

ggplot(MG_count_cpm.df, aes(x = WT, y = AD)) + geom_point(size = 1, color = "#666666", alpha = 0.2) +
   geom_point(data = MG_count_cpm.df[MG_count_cpm.df$gene %in% "mir-146a", ], aes(x = WT, y = AD), size = 2, color = "red") + xlim(c(0,9)) + ylim(c(0,9)) +
  ggtitle("WT vs AD in Microglia") +
  theme_classic() +
  xlab("log (average CPM + 1) WT") +
  ylab("log (average CPM + 1) AD") +
  theme(plot.title = element_text(size = 15, hjust = 0.5, family = "helvetica"),
        axis.title = element_text(size = 15),
        axis.text = element_text(size = 12, color = "black")) +
  geom_abline(slope = 1, color="red", linetype="dashed", size=.25) + 
  annotate("text", x = 5.5, y = 4.2, label = "pri-mir-146a", size = 4.5)
```

```{r warning=FALSE, message=FALSE, fig.width= 5.5, fig.height=5}
ggplot(MG_count_cpm.df, aes(x = AD, y = ADp40KO)) + geom_point(size = 1, color = "#666666", alpha = 0.2) +
   geom_point(data = MG_count_cpm.df[MG_count_cpm.df$gene %in% "mir-146a", ], aes(x = AD, y = ADp40KO), size = 2, color = "red") + xlim(c(0,9)) + ylim(c(0,9)) + 
  ggtitle("AD vs ADp40KO in Microglia") +
  theme_classic() +
  xlab("log (average CPM + 1) AD") +
  ylab("log (average CPM + 1) ADp40KO") +
  theme(plot.title = element_text(size = 15, hjust = 0.5, family = "helvetica"),
        axis.title = element_text(size = 15),
        axis.text = element_text(size = 12, color = "black")) +
  geom_abline(slope = 1, color="red", linetype="dashed", size=.25) + 
  annotate("text", x = 4.2, y = 5.8, label = "pri-mir-146a", size = 4.5)
```

```{r warning=FALSE, message=FALSE, fig.width= 5.5, fig.height=5}
ggplot(MG_count_cpm.df, aes(x = Cluster3, y = Cluster8)) + geom_point(size = 1, color = "#666666", alpha = 0.2) +
   geom_point(data = MG_count_cpm.df[MG_count_cpm.df$gene %in% "mir-146a", ], aes(x = Cluster3, y = Cluster8), size = 2, color = "red") + xlim(c(0,9)) + ylim(c(0,9)) + 
  ggtitle("Cluster3 vs Cluster8 in Microglia") +
  theme_classic() +
  xlab("log (average CPM + 1) Cluster3") +
  ylab("log (average CPM + 1) Cluster8") +
  theme(plot.title = element_text(size = 15, hjust = 0.5, family = "helvetica"),
        axis.title = element_text(size = 15),
        axis.text = element_text(size = 12, color = "black")) +
  geom_abline(slope = 1, color="red", linetype="dashed", size=.25) + 
  annotate("text", x = 4.2, y = 5.8, label = "pri-mir-146a", size = 4.5)
```


### 5. mir-206

miR-1-3p/206-3p (broadly conserved), miR-206-5p(very poorly conserved)

```{r warning=FALSE, message=FALSE, fig.width= 5.5, fig.height=5}
mir_206_3p <- read_xlsx("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/miRNA_lncRNA/Targetscan/TargetScan7.2__miR-1-3p_206-3p.predicted_targets.xlsx")
mir_206_5p <- read_xlsx("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/miRNA_lncRNA/Targetscan/TargetScan7.2__miR-206-5p.predicted_targets.xlsx")


```

both filter Cumulative weighted context++ score` < -.3 

```{r warning=FALSE, message=FALSE, fig.width= 11, fig.height=10}
#################
# 1. mir-206-3p
#################

MG_DE_AD_Ctrl$is.target <- MG_DE_AD_Ctrl$X %in% mir_206_3p$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG_DE_AD_Ctrl, alt = "two.sided") #0.2163
c1 <- cumulative_plot(MG_DE_AD_Ctrl, is.target = is.target,"mir_206-3p", "p-value = 0.2163") + ggtitle("AD vs Ctrl in Microglia mir_206-3p")

MG_DE_AD_Ctrl$is.target <- MG_DE_AD_Ctrl$X %in% mir_206_3p[mir_206_3p$`Cumulative weighted context++ score` < -.3, ]$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG_DE_AD_Ctrl, alt = "two.sided") #0.009183
c2 <- cumulative_plot(MG_DE_AD_Ctrl, is.target = is.target,"mir_206-3p filtered", "p-value = 0.009183") + ggtitle("AD vs Ctrl in Microglia mir_206-3p filtered")

#################
# 2. mir-206-5p
#################
MG_DE_AD_Ctrl$is.target <- MG_DE_AD_Ctrl$X %in% mir_206_5p$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG_DE_AD_Ctrl, alt = "two.sided") #0.02463
c3 <- cumulative_plot(MG_DE_AD_Ctrl, is.target = is.target,"mir-206-5p", "p-value = 0.02463") + ggtitle("AD vs Ctrl in Microglia mir-206-5p")

MG_DE_AD_Ctrl$is.target <- MG_DE_AD_Ctrl$X %in% mir_206_5p[mir_206_5p$`Cumulative weighted context++ score` < -.3,]$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG_DE_AD_Ctrl, alt = "two.sided") #0.628
c4 <- cumulative_plot(MG_DE_AD_Ctrl, is.target = is.target,"mir-206-5p filtered", "p-value = 0.628") + ggtitle("AD vs Ctrl in Microglia mir-206-5p filtered")
grid.arrange(c1, c2, c3, c4, ncol = 2)
```

```{r warning=FALSE, message=FALSE, fig.width= 11, fig.height=10}
#################
# 1. mir-206-3p
#################
MG_DE_AD_ADp40KO$is.target <- MG_DE_AD_ADp40KO$X %in% mir_206_3p$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG_DE_AD_ADp40KO, alt = "two.sided") #0.008269
c1 <- cumulative_plot(MG_DE_AD_ADp40KO, is.target = is.target,"mir_206-3p", "p-value = 0.008269") + ggtitle("AD vs ADp40KO in Microglia mir_206-3p")

MG_DE_AD_ADp40KO$is.target <- MG_DE_AD_ADp40KO$X %in% mir_206_3p[mir_206_3p$`Cumulative weighted context++ score` < -.3, ]$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG_DE_AD_ADp40KO, alt = "two.sided") #0.94
c2 <- cumulative_plot(MG_DE_AD_ADp40KO, is.target = is.target,"mir_206-3p filtered", "p-value = 0.94") + ggtitle("AD vs ADp40KO in Microglia mir_206-3p filtered")

#################
# 2. mir-206-5p
#################
MG_DE_AD_ADp40KO$is.target <- MG_DE_AD_ADp40KO$X %in% mir_206_5p$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG_DE_AD_ADp40KO, alt = "two.sided") #0.006348
c3 <- cumulative_plot(MG_DE_AD_ADp40KO, is.target = is.target,"mir_206_5p", "p-value = 0.006348") + ggtitle("AD vs ADp40KO in Microglia mir_206_5p")

MG_DE_AD_ADp40KO$is.target <- MG_DE_AD_ADp40KO$X %in% mir_206_5p[mir_206_5p$`Cumulative weighted context++ score` < -.3,]$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG_DE_AD_ADp40KO, alt = "two.sided") #0.5615
c4 <- cumulative_plot(MG_DE_AD_ADp40KO, is.target = is.target,"mir_206_5p filtered", "p-value = 0.5615") + ggtitle("AD vs ADp40KO in Microglia mir_206_5p filtered")
grid.arrange(c1, c2, c3, c4, ncol = 2)

```

```{r warning=FALSE, message=FALSE, fig.width= 11, fig.height=10}
#################
# 1. mir-206-3p
#################
MG.marker8vs3$is.target <- MG.marker8vs3$X %in% mir_206_3p$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG.marker8vs3, alt = "two.sided")#0.2663
c1 <- cumulative_plot(MG.marker8vs3, is.target = is.target,"mir_206_3p", "p-value = 0.2663") + ggtitle("Cluster 8 vs 3 in Microglia mir_206_3p")

MG.marker8vs3$is.target <- MG.marker8vs3$X %in% mir_206_3p[mir_206_3p$`Cumulative weighted context++ score` < -.3, ]$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG.marker8vs3, alt = "two.sided") #0.7879
c2 <- cumulative_plot(MG.marker8vs3, is.target = is.target,"mir_206_3p filtered", "p-value = 0.7879") + ggtitle("Cluster 8 vs 3 in Microglia mir_206_3p filtered")

#################
# 2. mir-206-5p
#################
MG.marker8vs3$is.target <- MG.marker8vs3$X %in% mir_206_5p$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG.marker8vs3, alt = "two.sided") #0.8873
c3 <- cumulative_plot(MG.marker8vs3, is.target = is.target,"mir_206_5p", "p-value = 0.8873") + ggtitle("Cluster 8 vs 3 in Microglia mir_206_5p")

MG.marker8vs3$is.target <- MG.marker8vs3$X %in% mir_206_5p[mir_206_5p$`Cumulative weighted context++ score` < -.3,]$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG.marker8vs3, alt = "two.sided") #0.3428
c4 <- cumulative_plot(MG_DE_AD_ADp40KO, is.target = is.target,"mir_206_5p filtered", "p-value = 0.3428") + ggtitle("Cluster 8 vs 3 in Microglia mir_206_5p filtered")
grid.arrange(c1, c2, c3, c4, ncol = 2)
```

### 6. let-7a

let-7-5p/miR-98-5p (broadly conserved), let-7-3p/miR-98-3p (very poorly conserved), let-7a-2-3p (very poorly conserved)

```{r warning=FALSE, message=FALSE, fig.width= 5.5, fig.height=5}
let_7_5p <- read_xlsx("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/miRNA_lncRNA/Targetscan/TargetScan7.2__let-7-5p_miR-98-5p.predicted_targets.xlsx")
let_7_3p <- read_xlsx("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/miRNA_lncRNA/Targetscan/TargetScan7.2__let-7-3p_miR-98-3p.predicted_targets.xlsx")
let_7a_2_3p <- read_xlsx("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/miRNA_lncRNA/Targetscan/TargetScan7.2__let-7a-2-3p.predicted_targets.xlsx")

```

let_7_5p: filter Cumulative weighted context++ score` < -.5 

let_7_3p: filter Cumulative weighted context++ score` < -.3 

let_7a_2_3p: filter Cumulative weighted context++ score` < -.4 

```{r warning=FALSE, message=FALSE, fig.width= 11, fig.height=15}
#################
# 1. let_7_5p
#################

MG_DE_AD_Ctrl$is.target <- MG_DE_AD_Ctrl$X %in% let_7_5p$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG_DE_AD_Ctrl, alt = "two.sided") #0.3167
c1 <- cumulative_plot(MG_DE_AD_Ctrl, is.target = is.target,"let_7_5p", "p-value = 0.3167") + ggtitle("AD vs Ctrl in Microglia let_7_5p")

MG_DE_AD_Ctrl$is.target <- MG_DE_AD_Ctrl$X %in% let_7_5p[let_7_5p$`Cumulative weighted context++ score` < -.5, ]$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG_DE_AD_Ctrl, alt = "two.sided") #0.4581
c2 <- cumulative_plot(MG_DE_AD_Ctrl, is.target = is.target,"let_7_5p filtered", "p-value = 0.4581") + ggtitle("AD vs Ctrl in Microglia let_7_5p filtered")

#################
# 2. let-7-3p
#################
MG_DE_AD_Ctrl$is.target <- MG_DE_AD_Ctrl$X %in% let_7_3p$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG_DE_AD_Ctrl, alt = "two.sided") #5.93e-06
c3 <- cumulative_plot(MG_DE_AD_Ctrl, is.target = is.target,"let_7_3p", "p-value = 5.93e-06") + ggtitle("AD vs Ctrl in Microglia let_7_3p")

MG_DE_AD_Ctrl$is.target <- MG_DE_AD_Ctrl$X %in% let_7_3p[let_7_3p$`Cumulative weighted context++ score` < -.3,]$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG_DE_AD_Ctrl, alt = "two.sided") #0.7751
c4 <- cumulative_plot(MG_DE_AD_Ctrl, is.target = is.target,"let_7_3p filtered", "p-value = 0.7751") + ggtitle("AD vs Ctrl in Microglia let_7_3p filtered")

#################
# 3. let_7a_2_3p
#################
MG_DE_AD_Ctrl$is.target <- MG_DE_AD_Ctrl$X %in% let_7a_2_3p$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG_DE_AD_Ctrl, alt = "two.sided") #0.01047
c5 <- cumulative_plot(MG_DE_AD_Ctrl, is.target = is.target,"let_7a_2_3p", "p-value = 0.01047") + ggtitle("AD vs Ctrl in Microglia let_7a_2_3p")

MG_DE_AD_Ctrl$is.target <- MG_DE_AD_Ctrl$X %in% let_7a_2_3p[let_7a_2_3p$`Cumulative weighted context++ score` < -.4,]$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG_DE_AD_Ctrl, alt = "two.sided") #0.009217
c6 <- cumulative_plot(MG_DE_AD_Ctrl, is.target = is.target,"let_7a_2_3p filtered", "p-value = 0.009217") + ggtitle("AD vs Ctrl in Microglia let_7a_2_3p filtered")

grid.arrange(c1, c2, c3, c4, c5, c6, ncol = 2)
```

```{r warning=FALSE, message=FALSE, fig.width= 11, fig.height=15}
#################
# 1. let_7_5p
#################

MG_DE_AD_ADp40KO$is.target <- MG_DE_AD_ADp40KO$X %in% let_7_5p$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG_DE_AD_ADp40KO, alt = "two.sided") #0.8359
c1 <- cumulative_plot(MG_DE_AD_ADp40KO, is.target = is.target,"let_7_5p", "p-value = 0.8359") + ggtitle("AD vs ADp40KO in Microglia let_7_5p")

MG_DE_AD_ADp40KO$is.target <- MG_DE_AD_ADp40KO$X %in% let_7_5p[let_7_5p$`Cumulative weighted context++ score` < -.5, ]$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG_DE_AD_ADp40KO, alt = "two.sided") #0.8822
c2 <- cumulative_plot(MG_DE_AD_ADp40KO, is.target = is.target,"let_7_5p filtered", "p-value = 0.8822") + ggtitle("AD vs ADp40KO in Microglia let_7_5p filtered")

#################
# 2. let-7-3p
#################
MG_DE_AD_ADp40KO$is.target <- MG_DE_AD_ADp40KO$X %in% let_7_3p$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG_DE_AD_ADp40KO, alt = "two.sided") #0.001065
c3 <- cumulative_plot(MG_DE_AD_ADp40KO, is.target = is.target,"let_7_3p", "p-value = 0.001065") + ggtitle("AD vs ADp40KO in Microglia let_7_3p")

MG_DE_AD_ADp40KO$is.target <- MG_DE_AD_ADp40KO$X %in% let_7_3p[let_7_3p$`Cumulative weighted context++ score` < -.3,]$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG_DE_AD_ADp40KO, alt = "two.sided") #0.7597
c4 <- cumulative_plot(MG_DE_AD_ADp40KO, is.target = is.target,"let_7_3p filtered", "p-value = 0.7597") + ggtitle("AD vs ADp40KO in Microglia let_7_3p filtered")

#################
# 3. let_7a_2_3p
#################
MG_DE_AD_ADp40KO$is.target <- MG_DE_AD_ADp40KO$X %in% let_7a_2_3p$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG_DE_AD_ADp40KO, alt = "two.sided") #0.002325
c5 <- cumulative_plot(MG_DE_AD_ADp40KO, is.target = is.target,"let_7a_2_3p", "p-value = 0.002325") + ggtitle("AD vs ADp40KO in Microglia let_7a_2_3p")

MG_DE_AD_ADp40KO$is.target <- MG_DE_AD_ADp40KO$X %in% let_7a_2_3p[let_7a_2_3p$`Cumulative weighted context++ score` < -.4,]$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG_DE_AD_ADp40KO, alt = "two.sided") #0.3017
c6 <- cumulative_plot(MG_DE_AD_ADp40KO, is.target = is.target,"let_7a_2_3p filtered", "p-value = 0.3017") + ggtitle("AD vs ADp40KO in Microglia let_7a_2_3p filtered")

grid.arrange(c1, c2, c3, c4, c5, c6, ncol = 2)
```

```{r warning=FALSE, message=FALSE, fig.width= 11, fig.height=15}
#################
# 1. let_7_5p
#################

MG.marker8vs3$is.target <- MG.marker8vs3$X %in% let_7_5p$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG.marker8vs3, alt = "two.sided") #0.1216
c1 <- cumulative_plot(MG.marker8vs3, is.target = is.target,"let_7_5p", "p-value = 0.1216") + ggtitle("Cluster 8 vs 3 in Microglia let_7_5p")

MG.marker8vs3$is.target <- MG.marker8vs3$X %in% let_7_5p[let_7_5p$`Cumulative weighted context++ score` < -.5, ]$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG.marker8vs3, alt = "two.sided") #0.77
c2 <- cumulative_plot(MG.marker8vs3, is.target = is.target,"let_7_5p filtered", "p-value = 0.77") + ggtitle("Cluster 8 vs 3 in Microglia let_7_5p filtered")

#################
# 2. let-7-3p
#################
MG.marker8vs3$is.target <- MG.marker8vs3$X %in% let_7_3p$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG.marker8vs3, alt = "two.sided") #0.006998
c3 <- cumulative_plot(MG.marker8vs3, is.target = is.target,"let_7_3p", "p-value = 0.006998") + ggtitle("Cluster 8 vs 3 in Microglia let_7_3p")

MG.marker8vs3$is.target <- MG.marker8vs3$X %in% let_7_3p[let_7_3p$`Cumulative weighted context++ score` < -.3,]$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG.marker8vs3, alt = "two.sided") #0.9665
c4 <- cumulative_plot(MG.marker8vs3, is.target = is.target,"let_7_3p filtered", "p-value = 0.9665") + ggtitle("Cluster 8 vs 3 in Microglia let_7_3p filtered")

#################
# 3. let_7a_2_3p
#################
MG.marker8vs3$is.target <- MG.marker8vs3$X %in% let_7a_2_3p$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG.marker8vs3, alt = "two.sided") #0.2467
c5 <- cumulative_plot(MG.marker8vs3, is.target = is.target,"let_7a_2_3p", "p-value = 0.2467") + ggtitle("Cluster 8 vs 3 in Microglia let_7a_2_3p")

MG.marker8vs3$is.target <- MG.marker8vs3$X %in% let_7a_2_3p[let_7a_2_3p$`Cumulative weighted context++ score` < -.4,]$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG.marker8vs3, alt = "two.sided") #0.02784
c6 <- cumulative_plot(MG.marker8vs3, is.target = is.target,"let_7a_2_3p filtered", "p-value = 0.02784") + ggtitle("Cluster 8 vs 3 in Microglia let_7a_2_3p filtered")

grid.arrange(c1, c2, c3, c4, c5, c6, ncol = 2)
```

### 7. mir-1970

```{r warning=FALSE, message=FALSE, fig.width= 5.5, fig.height=5}
mir_1970 <- read_xlsx("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/miRNA_lncRNA/Targetscan/TargetScan7.2__miR-1970.predicted_targets.xlsx")

```

filter Cumulative weighted context++ score` < -.4

```{r warning=FALSE, message=FALSE, fig.width= 11, fig.height=15}
#################
# 1. AD vs Ctrl
#################

MG_DE_AD_Ctrl$is.target <- MG_DE_AD_Ctrl$X %in% mir_1970$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG_DE_AD_Ctrl, alt = "two.sided") #0.4299
c1 <- cumulative_plot(MG_DE_AD_Ctrl, is.target = is.target,"mir_1970", "p-value = 0.4299") + ggtitle("AD vs Ctrl in Microglia mir_1970")

MG_DE_AD_Ctrl$is.target <- MG_DE_AD_Ctrl$X %in% mir_1970[mir_1970$`Cumulative weighted context++ score` < -.4, ]$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG_DE_AD_Ctrl, alt = "two.sided") #0.7691
c2 <- cumulative_plot(MG_DE_AD_Ctrl, is.target = is.target,"mir_1970 filtered", "p-value = 0.7691") + ggtitle("AD vs Ctrl in Microglia mir_1970 filtered")

###################
# 2. AD vs ADp40KO
###################

MG_DE_AD_ADp40KO$is.target <- MG_DE_AD_ADp40KO$X %in% mir_1970$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG_DE_AD_ADp40KO, alt = "two.sided") #3.381e-05
c3 <- cumulative_plot(MG_DE_AD_ADp40KO, is.target = is.target,"mir_1970", "p-value = 3.381e-05") + ggtitle("AD vs ADp40KO in Microglia mir_1970")

MG_DE_AD_ADp40KO$is.target <- MG_DE_AD_ADp40KO$X %in% mir_1970[mir_1970$`Cumulative weighted context++ score` < -.4, ]$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG_DE_AD_ADp40KO, alt = "two.sided") #0.04052
c4 <- cumulative_plot(MG_DE_AD_ADp40KO, is.target = is.target,"mir_1970 filtered", "p-value = 0.04052") + ggtitle("AD vs ADp40KO in Microglia mir_1970 filtered")

###################
# 3. Cluster 8 vs 3
###################

MG.marker8vs3$is.target <- MG.marker8vs3$X %in% mir_1970$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG.marker8vs3, alt = "two.sided") #0.5368
c5 <- cumulative_plot(MG.marker8vs3, is.target = is.target,"mir_1970", "p-value = 0.5368") + ggtitle("Cluster 8 vs 3 in Microglia mir_1970")

MG.marker8vs3$is.target <- MG.marker8vs3$X %in% mir_1970[mir_1970$`Cumulative weighted context++ score` < -.4, ]$`Target gene`
#wilcox.test(avg_logFC ~ is.target, data = MG.marker8vs3, alt = "two.sided") #0.4462
c6 <- cumulative_plot(MG.marker8vs3, is.target = is.target,"mir_1970 filtered", "p-value = 0.4462") + ggtitle("Cluster 8 vs 3 in Microglia mir_1970 filtered")

grid.arrange(c1, c2, c3, c4, c5, c6, ncol = 2)
```

```{r warning=FALSE, message=FALSE, fig.width= 11, fig.height=10}
sessionInfo()
```
