---
title: "10X_9sets"
author: "Skim"
date: "12/29/2019"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_depth: 4
    toc_float: false
    code_folding: hide
---

TODO: 
1. clustering / identify marker genes (Shirin)  
2. QC low quality cell (doublets)  
3. reproducibility / cell type composition (organoids paper, gini coefficient, KL divergence)    
4. DE / GO / GSEA (both bulk and scRNA)  
5. cell-cell interaction (cellphoneDB, NicheNet)

```{r warning=FALSE, message=FALSE}
Sys.setenv(PYTHONPATH="/data/local/rajewsky/shared_libs/python")
library(reticulate)
#use_python("/usr/bin/python3", required = TRUE)
library(Seurat)
library(ggplot2)
library(dplyr)
library(DT)
library(gridExtra)
library(nichenetr)
library(tidyverse)
#Rscript -e "library(rmarkdown); rmarkdown::render('/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20191229_pre_analysis_9sets.Rmd')"
```

```{r warning=FALSE, message=FALSE}
data9set <- Read10X("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/SP064_All/outs/filtered_feature_bc_matrix/")
```

### 1. QC plots

The number in here shows the median of each sample

```{r warning=FALSE, message=FALSE, fig.width= 8, fig.height= 5}
data9set.SO <- CreateSeuratObject(counts =  data9set,  min.cells = 3, min.features = 200, project = "ADp40KO")

# add gemgroup to meta data
gemgroup <- sapply(strsplit(rownames(data9set.SO@meta.data), split="-"), "[[", 2)
data9set.SO <- AddMetaData(object=data9set.SO, metadata=data.frame(gemgroup=gemgroup, row.names=rownames(data9set.SO@meta.data)))

# calculate MT genes
data9set.SO[["percent.mt"]] <- PercentageFeatureSet(object = data9set.SO, pattern = "^mt-")
```

```{r warning=FALSE, message=FALSE, fig.width= 8, fig.height= 5}
VlnPlot(data9set.SO, features = c("nFeature_RNA"), group.by = "gemgroup", pt.size = 0.2)

```

```{r warning=FALSE, message=FALSE, fig.width= 8, fig.height= 5}
VlnPlot(data9set.SO, features = c("percent.mt"), group.by = "gemgroup", pt.size = 0.2)

```

### 2. visualization

I filtered out nUMIs **more than 60,000** and **less than 200** and **mitochondria genes more than 10%**. 

parameters: nFeatures = 2000, PCs = 40, clustering resolution = 1, 

```{r warning=FALSE, message=FALSE, fig.width= 12, fig.height= 5}

#FeatureScatter(data9set.SO, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")

data9set.SO <- subset(x = data9set.SO, subset = nCount_RNA < 30000 & nCount_RNA > 200 & percent.mt < 5)  # 3787 -> 2722

data9set.SO <- NormalizeData(object = data9set.SO, normalization.method = "LogNormalize", scale.factor = 1e4, verbose = FALSE)
data9set.SO <- FindVariableFeatures(object = data9set.SO, nfeatures = 2000, selection.method = "vst", verbose = FALSE)
data9set.SO <- ScaleData(object = data9set.SO, vars.to.regress = c("percent.mt", "nCount_RNA", "gemgroup"))

# PCA
data9set.SO <- RunPCA(data9set.SO, features = VariableFeatures(object = data9set.SO), verbose = FALSE, ndims.print = "None")
#ElbowPlot(JK_0123.SO, ndims = 50)

data9set.SO <- FindNeighbors(object = data9set.SO, reduction = "pca", dims = 1:40)
data9set.SO <- FindClusters(object = data9set.SO, resolution = 0.8, verbose = FALSE)

data9set.SO <- RunTSNE(object = data9set.SO, dims = 1:40, nthreads = 8, perplexity = 30)
#d1 <- DimPlot(object = data9set.SO, reduction = "tsne")

data9set.SO <- RunUMAP(object = data9set.SO, dims = 1:40, n.neighbors = 20, min.dist = 0.35, n.epochs = 500, spread = 2)
#d2 <- DimPlot(object = data9set.SO, reduction = "umap")

DimPlot(object = data9set.SO, reduction = "umap")

#grid.arrange(d1, d2, ncol = 2)
```

```{r warning=FALSE, message=FALSE, fig.width= 10, fig.height=5}
FeaturePlot(data9set.SO, features = c("nCount_RNA", "Siglech"), by.col = 2)
```

```{r warning=FALSE, message=FALSE, fig.width= 8, fig.height=6}
DimPlot(object = data9set.SO, group.by = "gemgroup", pt.size = 0.05)
```

```{r warning=FALSE, message=FALSE, fig.width= 8, fig.height=6}
data9set.SO@meta.data$sample <- ifelse((data9set.SO$gemgroup == 1 | data9set.SO$gemgroup == 4 | data9set.SO$gemgroup == 7), "Ctrl", ifelse((data9set.SO$gemgroup == 2 | data9set.SO$gemgroup == 5 | data9set.SO$gemgroup == 8), "AD", "ADp40KO"))

data9set.SO@meta.data$sample <- factor(data9set.SO@meta.data$sample, levels = c("Ctrl", "AD", "ADp40KO"))
#table(data9set.SO$sample)

DimPlot(object = data9set.SO, group.by = "sample", pt.size = 0.05, cols = c("#003333", "#FF6600", "#FFFF00"))
```

```{r warning=FALSE, message=FALSE}
#data9set.markers <- FindAllMarkers(data9set.SO, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
#data9set.markers_top10 <- data9set.markers %>% group_by(cluster) %>% top_n(n = 20, wt = avg_logFC)
#datatable(data9set.markers_top10, options = list(pageLength = 10, dom = 'tip'))
#write.csv(data9set.markers, "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20191229_data9set_marker.csv")
#write.csv(data9set.markers_top10, "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20191229_data9set_marker_top20.csv")
data9set.markers_top10 <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20191229_data9set_marker_top20.csv")
datatable(data9set.markers_top10, options = list(pageLength = 10, dom = 'tip'))
```

```{r warning=FALSE, message=FALSE, fig.width= 12, fig.height=6}
dot_genes <- c("Hexb", "Siglech", "P2ry12",
               "Olig1", "Pdgfra",
               "Mbp", "Mag", "Neat1",
               "Sned1", "Alcam",
               "Cldn5", "Flt1", "Rgs5",
               "Gja1", "Cst3",
               "Snap25", "Cnr1",
               "Gad1", "Cit",
               "Cdh18",
               "Man1a", "Brd9",
               "Tshz2", "Gria4",
               "Adam11", "Rnf112",
               "Trpc6")



DotPlot(data9set.SO, features = dot_genes, cols = c("lightgrey", "blue"),  group.by = "seurat_clusters") + 
  theme(axis.text.y = element_text(size = 10), 
        axis.title.x = element_blank(),
        legend.title = element_text(size =10),
        legend.text = element_text(size = 10)) + 
  coord_flip()

```

```{r warning=FALSE, message=FALSE}
# Save Seurat object
# ref: https://satijalab.org/seurat/seurat_clustering_tutorial_part1.html
save(data9set.SO, file="/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20191229_pre_analysis_9sets_Seurat_object_res0.8.Robj")
# load Seurat object
# ref: https://satijalab.org/seurat/seurat_clustering_tutorial_part2.html
```

## whole DE analysis

```{r warning=FALSE, message=FALSE}
AD_vs_Ctrl <- FindMarkers(data9set.SO, ident.1 =  "AD", ident.2 = "Ctrl", group.by = "sample", test.use = "wilcox", logfc.threshold = 0.1)
AD_vs_Ctrl$gene <- rownames(AD_vs_Ctrl)

ADp40KO_vs_AD <- FindMarkers(data9set.SO, ident.1 =  "ADp40KO", ident.2 = "AD", group.by = "sample", test.use = "wilcox", logfc.threshold = 0.1)
ADp40KO_vs_AD$gene <- rownames(ADp40KO_vs_AD)

ADp40KO_vs_Ctrl <- FindMarkers(data9set.SO, ident.1 =  "ADp40KO", ident.2 = "Ctrl", group.by = "sample", test.use = "wilcox", logfc.threshold = 0.1)
ADp40KO_vs_Ctrl$gene <- rownames(ADp40KO_vs_Ctrl)

AD_vs_Ctrl_1 <- AD_vs_Ctrl[,c(2,6)]
ADp40KO_vs_AD_1 <- ADp40KO_vs_AD[,c(2,6)]
ADp40KO_vs_Ctrl_1 <- ADp40KO_vs_Ctrl[,c(2,6)]

AD_vs_Ctrl_ADp40KO <- Reduce(function(x,y) merge(x,y, by="gene",all=TRUE) ,list(AD_vs_Ctrl_1,ADp40KO_vs_AD_1,ADp40KO_vs_Ctrl_1))
colnames(AD_vs_Ctrl_ADp40KO) <- c("gene", "lfc_AD_vs_Ctrl", "lfc_ADp40KO_vs_AD", "lfc_ADp40KO_vs_Ctrl")

datatable(AD_vs_Ctrl_ADp40KO, options = list(pageLength = 10, dom = 'tip'))
```


## Microglia cluster DE analysis

cluster 5, 8, 30

```{r warning=FALSE, message=FALSE}

data9set.SO_cluster5_8_30 <- subset(data9set.SO, subset = seurat_clusters == c(5,8,30))

#data9set.SO_cluster5_8_30@meta.data

# AD vs Ctrl
AD_Ctrl_5_8_30 <- FindMarkers(data9set.SO_cluster5_8_30, ident.1 =  "AD", ident.2 = "Ctrl", 
                             group.by = "sample", test.use = "wilcox")

ADp40_AD_5_8_30 <- FindMarkers(data9set.SO_cluster5_8_30, ident.1 =  "ADp40KO", ident.2 = "AD", 
                             group.by = "sample", test.use = "wilcox")

ADp40_Ctrl_5_8_30 <- FindMarkers(data9set.SO_cluster5_8_30, ident.1 =  "ADp40KO", ident.2 = "Ctrl", 
                             group.by = "sample", test.use = "wilcox")
datatable(AD_Ctrl_5_8_30, options = list(pageLength = 5, dom = 'tip'))
datatable(ADp40_AD_5_8_30, options = list(pageLength = 5, dom = 'tip'))
datatable(ADp40_Ctrl_5_8_30, options = list(pageLength = 5, dom = 'tip'))
```

[Myo1e ref](https://molecularneurodegeneration.biomedcentral.com/articles/10.1186/s13024-018-0266-4)  
[Myo1e ref2](https://molecularneurodegeneration.biomedcentral.com/articles/10.1186/s13024-018-0254-8)
Ttr upregulated ADp40KO which preventing amyloid fibril formation   
[Ttr ref](https://www.ncbi.nlm.nih.gov/pubmed/8938268)
[Ttr ref2](https://www.jneurosci.org/content/31/35/12483)

## Cell population cluters

```{r warning=FALSE, message=FALSE, fig.width= 8, fig.height=8}
##################
# per experiment
##################
exp <- c("Ctrl", "AD", "ADp40KO")
# Create new data frame
exp_cell.df <- data.frame(ident = as.character(), cluster = as.numeric(), cells = as.numeric(), stringsAsFactors = FALSE)

for(i in exp){
  for(j in c(0:42)){
    if(i == "Ctrl"){
      k <- j
    }
    if(i == "AD"){
      k <- j + 43
    }
    if(i =="ADp40KO"){
      k <- j + 86
    }
    exp_cell.df[k+1, 1] <- i
    exp_cell.df[k+1, 2] <- j
    exp_cell.df[k+1, 3] <- table(data9set.SO@meta.data$sample == i & data9set.SO@meta.data$seurat_clusters == j)[2]
    #exp_cell.df[is.na(exp_cell.df)] <- 0
  }
}

#sum(exp_cell.df[exp_cell.df$ident == "AD", ]$cells)

# calculate percentage
exp_cell.df$total <- ifelse(exp_cell.df$ident == "Ctrl", sum(exp_cell.df[exp_cell.df$ident == "Ctrl", ]$cells), 
                            ifelse(exp_cell.df$ident == "AD", sum(exp_cell.df[exp_cell.df$ident == "AD", ]$cells),
                                   sum(exp_cell.df[exp_cell.df$ident == "AD", ]$cells)))
exp_cell.df$percentage <- round(100*(exp_cell.df$cells/exp_cell.df$total),2)

ggplot(exp_cell.df, aes(x = ident, y = cells, fill = factor(cluster)))+
  geom_bar(position="fill", stat="identity")
```

```{r warning=FALSE, message=FALSE, fig.width= 8, fig.height=8}
#############
# per sample 
#############
#data9set.SO$gemgroup

sample <- c(1, 2, 3, 4, 5, 6, 7, 8 ,9)
exp_cell.df2 <- data.frame(sample = as.character(), cluster = as.numeric(), cells = as.numeric(), stringsAsFactors = FALSE)

for(i in sample){
  for(j in c(0:42)){
    if(i == 1){
      k <- j
    }
    if(i == 2){
      k <- j + 43
    }
    if(i ==3){
      k <- j + 86
    }
    if(i ==4){
      k <- j + 129
    }
    if(i ==5){
      k <- j + 172
    }
    if(i ==6){
      k <- j + 215
    }
    if(i ==7){
      k <- j + 258
    }
    if(i ==8){
      k <- j + 301
    }
    if(i ==9){
      k <- j + 344
    }
    exp_cell.df2[k+1, 1] <- i
    exp_cell.df2[k+1, 2] <- j
    exp_cell.df2[k+1, 3] <- table(data9set.SO@meta.data$gemgroup == i & data9set.SO@meta.data$seurat_clusters == j)[2]
    exp_cell.df2[is.na(exp_cell.df2)] <- 0
  }
}

# calculate percentage
exp_cell.df2$total <- ifelse(exp_cell.df2$sample == 1, sum(exp_cell.df2[exp_cell.df2$sample == 1, ]$cells), 
                            ifelse(exp_cell.df2$sample == 2, sum(exp_cell.df2[exp_cell.df2$sample == 2, ]$cells),
                                   ifelse(exp_cell.df2$sample == 3, sum(exp_cell.df2[exp_cell.df2$sample == 3, ]$cells),
                                          ifelse(exp_cell.df2$sample == 4, sum(exp_cell.df2[exp_cell.df2$sample == 4, ]$cells),
                                                 ifelse(exp_cell.df2$sample == 5, sum(exp_cell.df2[exp_cell.df2$sample == 5, ]$cells),
                                                        ifelse(exp_cell.df2$sample == 6, sum(exp_cell.df2[exp_cell.df2$sample == 6, ]$cells),
                                                               ifelse(exp_cell.df2$sample == 7, sum(exp_cell.df2[exp_cell.df2$sample == 7, ]$cells),
                                                                      ifelse(exp_cell.df2$sample == 8, sum(exp_cell.df2[exp_cell.df2$sample == 8, ]$cells), sum(exp_cell.df2[exp_cell.df2$sample == 9, ]$cells)))))))))

exp_cell.df2$percentage <- round(100*(exp_cell.df2$cells/exp_cell.df2$total),2)

exp_cell.df2$sample <- factor(exp_cell.df2$sample, levels = c(1,4,7,2,5,8,3,6,9))

ggplot(exp_cell.df2, aes(x = sample, y = cells, fill = factor(cluster)))+
  geom_bar(position="fill", stat="identity")

```



## reproducibility (organoids)

**ICC**: To compare variability in cell-type proportion between data sets, intraclass correlation was computed by creating a table of cell-type proportions across each individual organoid, and using the ICC command in the irr R package v0.84, with model = “twoway”, type = “agreement”, and unit = “single”. 

**MI**: investigate the degree of variation whether the reproducibility of cell types generated was similar to that between brains. We calculated mutual information (MI) scores between cluster assignments and sample identity in each dataset. a lower MI score and lower z-score indicate higher reproducibility between individuals.

Mutual information was calculated between cluster assignments and individuals with the mpmi R package. z-scores were calculated by creating background distributions for each dataset, by permuting cluster assignments and re-calculating the mutual information score 1,000 times.

## NicheNet
ref: https://github.com/saeyslab/nichenetr/blob/master/vignettes/seurat_wrapper.md
ref: https://github.com/saeyslab/nichenetr

```{r warning=FALSE, message=FALSE}
# Read in NicheNet’s ligand-target prior model, ligand-receptor network and weighted integrated networks

ligand_target_matrix = readRDS(url("https://zenodo.org/record/3260758/files/ligand_target_matrix.rds"))
lr_network = readRDS(url("https://zenodo.org/record/3260758/files/lr_network.rds"))
weighted_networks = readRDS(url("https://zenodo.org/record/3260758/files/weighted_networks.rds"))
weighted_networks_lr = weighted_networks$lr_sig %>% inner_join(lr_network %>% distinct(from,to), by = c("from","to"))

# Convert to human to mouse data

lr_network = lr_network %>% mutate(from = convert_human_to_mouse_symbols(from), to = convert_human_to_mouse_symbols(to)) %>% drop_na()
colnames(ligand_target_matrix) = ligand_target_matrix %>% colnames() %>% convert_human_to_mouse_symbols()
rownames(ligand_target_matrix) = ligand_target_matrix %>% rownames() %>% convert_human_to_mouse_symbols()

ligand_target_matrix = ligand_target_matrix %>% .[!is.na(rownames(ligand_target_matrix)), !is.na(colnames(ligand_target_matrix))]

weighted_networks_lr = weighted_networks_lr %>% mutate(from = convert_human_to_mouse_symbols(from), to = convert_human_to_mouse_symbols(to)) %>% drop_na()
```

#### Define a “sender/niche” cell population and a “receiver/target” cell

```{r warning=FALSE, message=FALSE}
## receiver
receiver = c(0, 1, 2, 3, 4, 6)
expressed_genes_receiver = get_expressed_genes(receiver, data9set.SO, pct = 0.10)

background_expressed_genes = expressed_genes_receiver %>% .[. %in% rownames(ligand_target_matrix)]

## sender

sender_celltypes = c(5, 8, 30)
list_expressed_genes_sender = sender_celltypes %>% unique() %>% lapply(get_expressed_genes, data9set.SO, 0.10) # lapply to get the expressed genes of every sender cell type separately here
expressed_genes_sender = list_expressed_genes_sender %>% unlist() %>% unique()
```

#### Define a gene set of interest

```{r warning=FALSE, message=FALSE}
seurat_obj_receiver= subset(data9set.SO, idents = receiver)
seurat_obj_receiver = SetIdent(data9set.SO, value = seurat_obj_receiver[["sample"]])

condition_oi = "AD"
condition_reference = "Ctrl" 

DE_table_receiver = FindMarkers(object = seurat_obj_receiver, ident.1 = condition_oi, ident.2 = condition_reference, group.by = "sample", min.pct = 0.10) %>% rownames_to_column("gene")

geneset_oi = DE_table_receiver %>% filter(p_val_adj <= 0.05 & abs(avg_logFC) >= 0.25) %>% pull(gene)
geneset_oi = geneset_oi %>% .[. %in% rownames(ligand_target_matrix)]
```


#### Define a set of potential ligands

```{r warning=FALSE, message=FALSE}
ligands = lr_network %>% pull(from) %>% unique()
receptors = lr_network %>% pull(to) %>% unique()

expressed_ligands = intersect(ligands,expressed_genes_sender)
expressed_receptors = intersect(receptors,expressed_genes_receiver)

potential_ligands = lr_network %>% filter(from %in% expressed_ligands & to %in% expressed_receptors) %>% pull(from) %>% unique()

```

#### Perform NicheNet ligand activity analysis

```{r warning=FALSE, message=FALSE, fig.width= 8, fig.height=5}
ligand_activities = predict_ligand_activities(geneset = geneset_oi, background_expressed_genes = background_expressed_genes, ligand_target_matrix = ligand_target_matrix, potential_ligands = potential_ligands)

ligand_activities = ligand_activities %>% arrange(-pearson) %>% mutate(rank = rank(desc(pearson)))


best_upstream_ligands = ligand_activities %>% top_n(20, pearson) %>% arrange(-pearson) %>% pull(test_ligand) %>% unique()


DotPlot(data9set.SO, features = best_upstream_ligands %>% rev(), cols = "RdYlBu") + RotatedAxis() + coord_flip()
```

#### Infer receptors and top-predicted target genes of ligands

#### Active target gene inference

```{r warning=FALSE, message=FALSE}
active_ligand_target_links_df = best_upstream_ligands %>% lapply(get_weighted_ligand_target_links, geneset = geneset_oi, ligand_target_matrix = ligand_target_matrix, n = 200) %>% bind_rows() %>% drop_na()

#################################################################################
get_weighted_ligand_target_links
top_n_score = ligand_target_matrix[, best_upstream_ligands] %>% sort(decreasing = T) %>% head(500) %>% min()
top_n_score

intersect(colnames(ligand_target_matrix), best_upstream_ligands)


targets = intersect(ligand_target_matrix[, best_upstream_ligands] %>% .[. >= top_n_score] %>% names(), geneset_oi)

targets
##################################################################################
#active_ligand_target_links = prepare_ligand_target_visualization(ligand_target_df = active_ligand_target_links_df, ligand_target_matrix = ligand_target_matrix, cutoff = 0.33)

#order_ligands = intersect(best_upstream_ligands, colnames(active_ligand_target_links)) %>% rev() %>% make.names()
#order_targets = active_ligand_target_links_df$target %>% unique() %>% intersect(rownames(active_ligand_target_links)) %>% make.names()
#rownames(active_ligand_target_links) = rownames(active_ligand_target_links) %>% make.names() # make.names() for heatmap visualization of genes like H2-T23
#colnames(active_ligand_target_links) = colnames(active_ligand_target_links) %>% make.names() # make.names() for heatmap visualization of genes like H2-T23

#vis_ligand_target = active_ligand_target_links[order_targets,order_ligands] %>% t()


#p_ligand_target_network = vis_ligand_target %>% make_heatmap_ggplot("Prioritized ligands","Predicted target genes", color = "purple",legend_position = "top", x_axis_position = "top",legend_title = "Regulatory potential") + 
#  theme(axis.text.x = element_text(face = "italic")) + 
#  scale_fill_gradient2(low = "whitesmoke",  high = "purple", breaks = c(0,0.006,0.012))

```

```{r warning=FALSE, message=FALSE}
# combined heatmap: overlay ligand activities with target genes
ligand_pearson_matrix = ligand_activities %>% tidygraph::select(pearson) %>% as.matrix() %>% magrittr::set_rownames(ligand_activities$test_ligand)


rownames(ligand_pearson_matrix) = rownames(ligand_pearson_matrix) %>% make.names()
colnames(ligand_pearson_matrix) = colnames(ligand_pearson_matrix) %>% make.names()

vis_ligand_pearson = ligand_pearson_matrix[order_ligands, ] %>% as.matrix(ncol = 1) %>% magrittr::set_colnames("Pearson")
p_ligand_pearson = vis_ligand_pearson %>% make_heatmap_ggplot("Prioritized ligands","Ligand activity", color = "darkorange",legend_position = "top", x_axis_position = "top", legend_title = "Pearson correlation coefficient\ntarget gene prediction ability)") + theme(legend.text = element_text(size = 9))

figures_without_legend = cowplot::plot_grid(p_ligand_pearson + theme(legend.position = "none", axis.ticks = element_blank()) + theme(axis.title.x = element_text()),
    p_ligand_target_network + theme(legend.position = "none", axis.ticks = element_blank()) + ylab(""),
    align = "hv",
    nrow = 1,
    rel_widths = c(ncol(vis_ligand_pearson)+10, ncol(vis_ligand_target)))

legends = cowplot::plot_grid(
    ggpubr::as_ggplot(ggpubr::get_legend(p_ligand_pearson)),
    ggpubr::as_ggplot(ggpubr::get_legend(p_ligand_target_network)),
    nrow = 1,
    align = "h")

```

## Bigscale

[ref](https://github.com/iaconogi/bigSCale2)


```{r warning=FALSE, message=FALSE}
sessionInfo()

```
