---
title: "Ttr 2nd analysis"
author: "Skim"
date: '2020 4 29 '
output: 
  html_document:
    df_print: paged
    toc: true
    toc_depth: 4
    toc_float: false
    code_folding: hide
---

This report is for Ttr heatmap in below cell types (Choroid Plexus, Neurons, Microglia, Oligodendrocytes). However, while creating heatmap, I found other cell types except Choroid Plexus, expression of Ttr depends on exist of Choroid Plexus in their data set.  


```{r warning=FALSE, message=FALSE}
.libPaths(c("/home/skim/R/usr_lib", .libPaths()))
library(Seurat)
library(ggplot2)
library(gridExtra)
library(ggrepel)
library(tidyr)
library(dplyr)
```

```{r warning=FALSE, message=FALSE}

load(file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200221_figure1_for_meeting_report_Seurat_object.Robj")
```

```{r warning=FALSE, message=FALSE}
data9set_cleaned.SO@meta.data$sample <- ifelse((data9set_cleaned.SO$gemgroup == 1 | data9set_cleaned.SO$gemgroup == 4 | data9set_cleaned.SO$gemgroup == 7), "Ctrl", ifelse((data9set_cleaned.SO$gemgroup == 2 | data9set_cleaned.SO$gemgroup == 5 | data9set_cleaned.SO$gemgroup == 8), "ADp40KO", "AD"))
```

```{r warning=FALSE, message=FALSE}
data9set_CP.SO <- subset(data9set_cleaned.SO, subset = seurat_clusters %in% c(26))
data9set_OL.SO <- subset(data9set_cleaned.SO, subset = seurat_clusters %in% c(1, 5, 6))
data9set_MG.SO <- subset(data9set_cleaned.SO, subset = seurat_clusters %in% c(3, 8))
```

#### 1. Number of Choroid Plexus in each sample

```{r warning=FALSE, message=FALSE}
cp.df <- table(data9set_CP.SO$gemgroup) %>% as.data.frame
colnames(cp.df) <- c("sample", "Freq")
ggplot(cp.df, aes(x = sample, y = Freq)) + geom_bar(stat="identity") + ggtitle("Number of Choroid Plexus cells in each sample")
```

#### 2. Ttr expression in Choroid Plexus

```{r warning=FALSE, message=FALSE}
VlnPlot(data9set_CP.SO, features = "Ttr", group.by = "gemgroup") + ggtitle("Ttr in Choroid Plexus")
```

#### 2. Ttr expression in Oligodendrocytes

```{r warning=FALSE, message=FALSE}
VlnPlot(data9set_OL.SO, features = "Ttr", group.by = "gemgroup") + ggtitle("Ttr in Oligo")
```

#### 3. Ttr expression in Microglia

```{r warning=FALSE, message=FALSE}
VlnPlot(data9set_MG.SO, features = "Ttr", group.by = "gemgroup") + ggtitle("Ttr in Microglia")
```

#### 4. Ttr expression in total sample

```{r warning=FALSE, message=FALSE}
VlnPlot(data9set_cleaned.SO, features = "Ttr", group.by = "gemgroup", pt.size = 0)
```

#### 5. heatmap of Ttr in Choroid Plexus

```{r warning=FALSE, message=FALSE}
##########################
# z-score for each sample
##########################
data9set_cleaned_OL_1.SO <- subset(data9set_OL.SO, subset = gemgroup %in% "1")
data9set_cleaned_OL_2.SO <- subset(data9set_OL.SO, subset = gemgroup %in% "2")
data9set_cleaned_OL_3.SO <- subset(data9set_OL.SO, subset = gemgroup %in% "3")
data9set_cleaned_OL_4.SO <- subset(data9set_OL.SO, subset = gemgroup %in% "4")
data9set_cleaned_OL_5.SO <- subset(data9set_OL.SO, subset = gemgroup %in% "5")
data9set_cleaned_OL_6.SO <- subset(data9set_OL.SO, subset = gemgroup %in% "6")
data9set_cleaned_OL_7.SO <- subset(data9set_OL.SO, subset = gemgroup %in% "7")
data9set_cleaned_OL_8.SO <- subset(data9set_OL.SO, subset = gemgroup %in% "8")
data9set_cleaned_OL_9.SO <- subset(data9set_OL.SO, subset = gemgroup %in% "9")

```

```{r warning=FALSE, message=FALSE}
# extract log(CPM+1) data from Seurat object
z_score_from_gene_list <- function(Seurat.SO, gene_list){
  
  df <- as.data.frame(Matrix::rowMeans(GetAssayData(Seurat.SO, slot = "data")))
  #df$gene <- rownames(df)
  df <- df[rownames(df) %in% gene_list,]
  return(df)
}

OL1.df <- z_score_from_gene_list(data9set_cleaned_OL_1.SO, "Ttr") %>% as.data.frame
colnames(OL1.df) <- c("Ctrl_1")
rownames(OL1.df) <- c("Ttr")
OL2.df <- z_score_from_gene_list(data9set_cleaned_OL_2.SO, "Ttr") %>% as.data.frame
colnames(OL2.df) <- c("ADp40KO_1")
rownames(OL2.df) <- c("Ttr")
OL3.df <- z_score_from_gene_list(data9set_cleaned_OL_3.SO, "Ttr") %>% as.data.frame
colnames(OL3.df) <- c("AD_1")
rownames(OL3.df) <- c("Ttr")
OL4.df <- z_score_from_gene_list(data9set_cleaned_OL_4.SO, "Ttr") %>% as.data.frame
colnames(OL4.df) <- c("Ctrl_2")
rownames(OL4.df) <- c("Ttr")
OL5.df <- z_score_from_gene_list(data9set_cleaned_OL_5.SO, "Ttr") %>% as.data.frame
colnames(OL5.df) <- c("ADp40KO_2")
rownames(OL5.df) <- c("Ttr")
OL6.df <- z_score_from_gene_list(data9set_cleaned_OL_6.SO, "Ttr") %>% as.data.frame
colnames(OL6.df) <- c("AD_2")
rownames(OL6.df) <- c("Ttr")
OL7.df <- z_score_from_gene_list(data9set_cleaned_OL_7.SO, "Ttr") %>% as.data.frame
colnames(OL7.df) <- c("Ctrl_3")
rownames(OL7.df) <- c("Ttr")
OL8.df <- z_score_from_gene_list(data9set_cleaned_OL_8.SO, "Ttr") %>% as.data.frame
colnames(OL8.df) <- c("ADp40KO_3")
rownames(OL8.df) <- c("Ttr")
OL9.df <- z_score_from_gene_list(data9set_cleaned_OL_9.SO, "Ttr") %>% as.data.frame
colnames(OL9.df) <- c("AD_3")
rownames(OL9.df) <- c("Ttr")

# merge all data frame
OL.df <- cbind(OL1.df, OL2.df, OL3.df, OL4.df, OL5.df, OL6.df, OL7.df, OL8.df, OL9.df)

```

```{r warning=FALSE, message=FALSE, fig.width=8, fig.height=2.5}
OL.df2 <- OL.df %>% gather(sample, z_score, Ctrl_1:AD_3)

OL.df2$sample <- factor(OL.df2$sample, levels = c("Ctrl_1", "Ctrl_2", "Ctrl_3", "AD_1", "AD_2", "AD_3", "ADp40KO_1", "ADp40KO_2", "ADp40KO_3"))

ggplot(data = OL.df2, mapping = aes(x = sample, y = "Ttr", fill = z_score)) +
  geom_tile() +  theme_classic() + theme(axis.title = element_blank(), axis.text.x = element_text(angle = 90))

```
