---
title: "OPC analysis"
author: "Skim"
date: '2020 3 20 '
output: 
  html_document:
    df_print: paged
    toc: true
    toc_depth: 4
    toc_float: false
    code_folding: hide
---

This report is about OPC analysis. From 

From previous literatures, PDGFRa is not required for generation and differentiation of dorsally derived cortical OPCs. [ref](https://www.jneurosci.org/content/jneuro/38/44/9505.full.pdf)


```{r warning=FALSE, message=FALSE}
.libPaths(c("/home/skim/R/usr_lib", .libPaths()))
#Sys.setenv(PYTHONPATH="/data/local/rajewsky/shared_libs/python")
#library(reticulate)
#use_python("/usr/bin/python3", required = TRUE)
library(Seurat)
library(ggplot2)
library(dplyr)
library(DT)
library(gridExtra)
library(nichenetr)
library(tidyverse)
library(irr)
library(fgsea)
library(MAST)
library(ggrepel)
library(biomaRt)
library(topGO)
library(viridis)
source("/data/rajewsky/home/skim/Microglia_Heppner/Codes/DroNc_AD/src/TopGO_scRNA.R")
#Rscript -e "library(rmarkdown); rmarkdown::render('/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20191229_pre_analysis_9sets.Rmd')"
```

```{r warning=FALSE, message=FALSE}
load(file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200221_figure1_for_meeting_report_Seurat_object.Robj")
```

### 1. OPC cluster analysis

This analysis is done by OPC cluster (from res 0.8 and cluster 12 and 38) and I investigate more in here.

```{r warning=FALSE, message=FALSE, fig.width = 6, fig.height=5}
data9set_cleaned_OPC.SO <- subset(data9set_cleaned.SO, subset = seurat_clusters %in% c(12, 38))

data9set_cleaned_OPC_Ctrl.SO <- subset(data9set_cleaned_OPC.SO, subset = sample %in% "Ctrl")
data9set_cleaned_OPC_AD.SO <- subset(data9set_cleaned_OPC.SO, subset = sample %in% "AD")
data9set_cleaned_OPC_ADp40KO.SO <- subset(data9set_cleaned_OPC.SO, subset = sample %in% "ADp40KO")


DimPlot(object = data9set_cleaned_OPC.SO, reduction = "umap", group.by = "seurat_clusters")+ scale_x_continuous(limits = c(15,25)) + scale_y_continuous(limits = c(-20, 0))
```

```{r warning=FALSE, message=FALSE, fig.width = 12, fig.height=5}
d1 <- DimPlot(object = data9set_cleaned_OPC.SO, 
        cells = as.vector(rownames(data9set_cleaned_OPC.SO@meta.data[data9set_cleaned_OPC.SO@meta.data$sample %in% "Ctrl",])), reduction = "umap", cols = viridis(3)[1]) + ggtitle("OPC Ctrl") + scale_x_continuous(limits = c(15,25)) + scale_y_continuous(limits = c(-20, 0)) + theme(legend.position = "none", axis.text = element_blank(), axis.ticks = element_blank(), plot.title = element_text(hjust = 0.5)) + geom_text(aes(x = 24, y = -2), label = paste0(length(rownames(data9set_cleaned_OPC.SO@meta.data[data9set_cleaned_OPC.SO@meta.data$sample %in% "Ctrl",])), " cells"), size = 5)

d2 <- DimPlot(object = data9set_cleaned_OPC.SO, 
        cells = as.vector(rownames(data9set_cleaned_OPC.SO@meta.data[data9set_cleaned_OPC.SO@meta.data$sample %in% "AD",])), reduction = "umap", cols = viridis(3)[2]) + ggtitle("OPC AD") + scale_x_continuous(limits = c(15,25)) + scale_y_continuous(limits = c(-20, 0)) + theme(legend.position = "none", axis.text = element_blank(), axis.ticks = element_blank(), plot.title = element_text(hjust = 0.5)) + geom_text(aes(x = 24, y = -2), label = paste0(length(rownames(data9set_cleaned_OPC.SO@meta.data[data9set_cleaned_OPC.SO@meta.data$sample %in% "AD",])), " cells"), size = 5)

d3 <- DimPlot(object = data9set_cleaned_OPC.SO, 
        cells = as.vector(rownames(data9set_cleaned_OPC.SO@meta.data[data9set_cleaned_OPC.SO@meta.data$sample %in% "ADp40KO",])), reduction = "umap", cols = viridis(3)[3]) + ggtitle("OPC ADp40KO") + scale_x_continuous(limits = c(15,25)) + scale_y_continuous(limits = c(-20, 0)) + theme(legend.position = "none", axis.text = element_blank(), axis.ticks = element_blank(), plot.title = element_text(hjust = 0.5)) + geom_text(aes(x = 24, y = -2), label = paste0(length(rownames(data9set_cleaned_OPC.SO@meta.data[data9set_cleaned_OPC.SO@meta.data$sample %in% "ADp40KO",])), " cells"), size = 5)

grid.arrange(d1, d2, d3, ncol = 3)
```

### 2. cell type proportion

```{r warning=FALSE, message=FALSE, fig.width = 6, fig.height=4.5}
plot.data <- data9set_cleaned_OPC.SO@meta.data %>%
    dplyr::select(sample, gemgroup = gemgroup, cluster = seurat_clusters) %>%
    mutate(cluster = cluster) %>%
    group_by(gemgroup, cluster) %>%
  summarise(count = n()) %>%
    mutate(sample_total = sum(count)) %>%
    mutate(cell_prop = count / sample_total)

plot.data$gemgroup <- factor(plot.data$gemgroup, levels = c("1", "4", "7", "2", "5", "8", "3", "6", "9"))
plot.data <- as.data.frame(plot.data)

p1 <- data.frame(expgroup = c(rep("Ctrl-1", 2), rep("ADp40KO-1", 2), rep("AD-1", 2),
                              rep("Ctrl-2", 2), rep("ADp40KO-2", 2), rep("AD-2", 2),
                              rep("Ctrl-3", 2), rep("ADp40KO-3", 2), rep("AD-3", 2)), stringsAsFactors = FALSE)

plot.data <- cbind(plot.data, p1)

plot.data$expgroup <- factor(plot.data$expgroup, levels = c("Ctrl-1", "Ctrl-2", "Ctrl-3", "AD-1", "AD-2", "AD-3", "ADp40KO-1", "ADp40KO-2", "ADp40KO-3"))

ggplot(plot.data, aes(x = expgroup, y = cell_prop, fill = cluster)) +
    geom_col() + theme_classic() + scale_y_continuous(expand = c(0,0)) + ylab("percentage") + 
  theme(axis.text.x = element_text(size =14, angle = 90, color = "black"), axis.title.x = element_blank(), axis.text.y = element_text(size =14, color = "black"),
        axis.title.y = element_text(size =14), legend.title = element_blank(), legend.text = element_text(size = 12)) +  scale_fill_manual(values=c("#CC0000",  "#33CC33"))

```

### 3. feature markers

```{r warning=FALSE, message=FALSE, fig.width = 12, fig.height=5}
data9set.markers <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200210_FindMarkers_each_cluster_data9set_marker.csv", row.names = 1)
data9set_OPC.markers <- data9set.markers[data9set.markers$cluster %in% c(12, 38), ]


data9set_OPC_DE_AD_Ctrl_markers <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/DE_csv_files/Cell_type/20200221_Cell_type_DE_by_MAST_OPCAD_Ctrl.csv", row.names = 1)

data9set_OPC_DE_AD_ADp40KO_markers <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/DE_csv_files/Cell_type/20200221_Cell_type_DE_by_MAST_OPCAD_ADp40KO.csv", row.names = 1)
```

First I checked known OPC markers from literatures Olig1, Cspg4, Pdgfra [marker ref](https://www.nature.com/articles/s41598-017-14207-0)

```{r warning=FALSE, message=FALSE, fig.width = 15, fig.height=4.5}
f1 <- FeaturePlot(data9set_cleaned_OPC.SO, features = c("Olig1"), cols = c("#FFCC00",'#003366'))+ scale_x_continuous(limits = c(15,25)) + scale_y_continuous(limits = c(-20, 0))
f2 <- FeaturePlot(data9set_cleaned_OPC.SO, features = c("Cspg4"), cols = c("#FFCC00",'#003366'))+ scale_x_continuous(limits = c(15,25)) + scale_y_continuous(limits = c(-20, 0))
f3 <- FeaturePlot(data9set_cleaned_OPC.SO, features = c("Pdgfra"), cols = c("#FFCC00",'#003366'))+ scale_x_continuous(limits = c(15,25)) + scale_y_continuous(limits = c(-20, 0))

grid.arrange(f1, f2, f3,  ncol = 3)
```

#### 3-1. Cluster 12 markers

```{r warning=FALSE, message=FALSE, fig.width = 15, fig.height=4.5}
f1 <- FeaturePlot(data9set_cleaned_OPC.SO, features = c("Lhfpl3"), cols = c("#FFCC00",'#003366'))+ scale_x_continuous(limits = c(15,25)) + scale_y_continuous(limits = c(-20, 0))
f2 <- FeaturePlot(data9set_cleaned_OPC.SO, features = c("Sox6"), cols = c("#FFCC00",'#003366'))+ scale_x_continuous(limits = c(15,25)) + scale_y_continuous(limits = c(-20, 0))
f3 <- FeaturePlot(data9set_cleaned_OPC.SO, features = c("Nxph1"), cols = c("#FFCC00",'#003366'))+ scale_x_continuous(limits = c(15,25)) + scale_y_continuous(limits = c(-20, 0))

grid.arrange(f1, f2, f3,  ncol = 3)
```

#### 3-2. Cluster 38 markers

I previously reported in lncRNAs that 9630013A20Rik is in GO:0048709 (oligodendrocyte differentiation) [ref](http://www.informatics.jax.org/go/term/GO:0022008)

```{r warning=FALSE, message=FALSE, fig.width = 15, fig.height=4.5}
f1 <- FeaturePlot(data9set_cleaned_OPC.SO, features = c("Bcas1"), cols = c("#FFCC00",'#003366'))+ scale_x_continuous(limits = c(15,25)) + scale_y_continuous(limits = c(-20, 0))
f2 <- FeaturePlot(data9set_cleaned_OPC.SO, features = c("9630013A20Rik"), cols = c("#FFCC00",'#003366'))+ scale_x_continuous(limits = c(15,25)) + scale_y_continuous(limits = c(-20, 0))
f3 <- FeaturePlot(data9set_cleaned_OPC.SO, features = c("Fyn"), cols = c("#FFCC00",'#003366'))+ scale_x_continuous(limits = c(15,25)) + scale_y_continuous(limits = c(-20, 0))

grid.arrange(f1, f2, f3,  ncol = 3)
```

#### 3-3. 9630013A20Rik

9630013A20Rik is gene that up-regulated in AD vs Ctrl & ADp40KO

```{r warning=FALSE, message=FALSE, fig.width = 12, fig.height=4.5}
f1 <- FeaturePlot(data9set_cleaned_OPC_Ctrl.SO, features = c("9630013A20Rik"),order = FALSE, cols = c("#FFCC00",'#003366'))+ scale_x_continuous(limits = c(15,25)) + scale_y_continuous(limits = c(-20, 0))
f2 <- FeaturePlot(data9set_cleaned_OPC_AD.SO, features = c("9630013A20Rik"),order = FALSE, cols = c("#FFCC00",'#003366'))+ scale_x_continuous(limits = c(15,25)) + scale_y_continuous(limits = c(-20, 0))
f3 <- FeaturePlot(data9set_cleaned_OPC_ADp40KO.SO, features = c("9630013A20Rik"),order = FALSE, cols = c("#FFCC00",'#003366'))+ scale_x_continuous(limits = c(15,25)) + scale_y_continuous(limits = c(-20, 0))

grid.arrange(f1, f2, f3, ncol = 3)
```

#### 3-3. Ttr

Ttr is up-regulated in AD compare to ADp40KO

```{r warning=FALSE, message=FALSE, fig.width = 12, fig.height=4.5}
f1 <- FeaturePlot(data9set_cleaned_OPC_Ctrl.SO, features = c("Ttr"),order = FALSE, cols = c("#FFCC00",'#003366'))+ scale_x_continuous(limits = c(15,25)) + scale_y_continuous(limits = c(-20, 0))
f2 <- FeaturePlot(data9set_cleaned_OPC_AD.SO, features = c("Ttr"),order = FALSE, cols = c("#FFCC00",'#003366'))+ scale_x_continuous(limits = c(15,25)) + scale_y_continuous(limits = c(-20, 0))
f3 <- FeaturePlot(data9set_cleaned_OPC_ADp40KO.SO, features = c("Ttr"),order = FALSE, cols = c("#FFCC00",'#003366'))+ scale_x_continuous(limits = c(15,25)) + scale_y_continuous(limits = c(-20, 0))

grid.arrange(f1, f2, f3, ncol = 3)
```

### 4. GO analysis

To identify the function of OPC two clusters, I used GO for each cluster. 

In here, genes of interest are padj < 0.01 & logFC > 0.5 from marker genes. Backgrond genes are derived from DE analysis genes with pct > 0.01. 

```{r warning=FALSE, message=FALSE, fig.width = 15, fig.height=5}
cluster_GO <- function(genesOfInterest, geneUniverse){

geneList <- factor(as.integer(geneUniverse %in% genesOfInterest))
names(geneList) <- geneUniverse

onts = c( "MF", "BP", "CC" )
tab <- as.list(onts)
names(tab) <- onts

for(i in 1:3){

GOdata <- new("topGOdata",
                description = "GOanalysis",
                ontology = onts[i],
                allGenes = geneList,
                annot = annFUN.org,
                mapping = "org.Mm.eg.db",
                ID = "SYMBOL",
                nodeSize = 20)


res.result1 <- runTest(GOdata, statistic = "fisher", algorithm = "elim")
res.result2 <- runTest(GOdata, statistic = "fisher", algorithm = "classic")


tab[[i]] <- cbind(data.frame(category = onts[i]), GenTable(GOdata, Fisher.elim = res.result1,
                        Fisher.classic = res.result2,
                        orderBy = "Fisher.elim" , topNodes = 30))


}

topGOResults <- plyr::rbind.fill(tab)
topGOResults.df <- as.data.frame(topGOResults)
topGOResults.df$gene_ratio <- topGOResults.df$Significant / topGOResults.df$Annotated

# modification appropriate for plot
topGOResults.df$Fisher.elim <- as.numeric(topGOResults.df$Fisher.elim)
topGOResults.df$Fisher.classic <- as.numeric(topGOResults.df$Fisher.classic)
topGOResults.df$Term <- factor(topGOResults.df$Term, levels = rev(unique(topGOResults.df$Term)))

return(topGOResults.df)
}
# ggplot function
scRNA_TopGO_plot2 <- function(topGOResults.df){

topGOResults.df$Fisher.elim <- as.numeric(topGOResults.df$Fisher.elim)
topGOResults.df$Fisher.elim <- -log10(topGOResults.df$Fisher.elim)
topGOResults.df$Term <- factor(topGOResults.df$Term, levels = rev(unique(topGOResults.df$Term)))

ggplot(topGOResults.df, aes(x=gene_ratio,
               y=Term,
               colour=Fisher.elim,
               size=Significant)) +
        geom_point() +
        expand_limits(x=0) +
        labs(x="gene ratio", y="GO term", colour="-log10(p-value)", size="Significant") +
  theme_minimal() + theme(axis.text = element_text(size = 10))

}
```

#### 4-1. cluster 12

```{r warning=FALSE, message=FALSE, fig.width = 15, fig.height=5}
# gene of interest
# In here, I extract genes padj < 0.01 & logFC > 0.5
genesOfInterest <- as.character(data9set_OPC.markers[which(data9set_OPC.markers$cluster %in% "12" &data9set_OPC.markers$p_val_adj < 0.01 & data9set_OPC.markers$avg_logFC > 0.5), ]$gene) 

# background genes
geneUniverse <- rownames(data9set_OPC_DE_AD_Ctrl_markers[data9set_OPC_DE_AD_Ctrl_markers$pct.1 > 0.01 & data9set_OPC_DE_AD_Ctrl_markers$pct.2 > 0.01, ])

topGOResults.df <- cluster_GO(genesOfInterest, geneUniverse)

topGOResults_MF.df <- topGOResults.df[which(topGOResults.df$category == "MF" & topGOResults.df$Annotated < 500), ][c(1:7), ]
g1 <- scRNA_TopGO_plot2(topGOResults_MF.df)
g1 <- g1 + ggtitle("Molecular function")

topGOResults_BP.df <- topGOResults.df[which(topGOResults.df$category == "BP" & topGOResults.df$Annotated < 500), ][c(1:7), ]
g2 <- scRNA_TopGO_plot2(topGOResults_BP.df)
g2 <- g2 + ggtitle("Biological process")

grid.arrange(g1, g2, ncol = 2)
```

#### 4-2. cluster 38

```{r warning=FALSE, message=FALSE, fig.width = 15, fig.height=5}
# gene of interest
# In here, I extract genes padj < 0.01 & logFC > 0.5
genesOfInterest <- as.character(data9set_OPC.markers[which(data9set_OPC.markers$cluster %in% "38" &data9set_OPC.markers$p_val_adj < 0.01 & data9set_OPC.markers$avg_logFC > 0.5), ]$gene) 

# background genes
geneUniverse <- rownames(data9set_OPC_DE_AD_Ctrl_markers[data9set_OPC_DE_AD_Ctrl_markers$pct.1 > 0.01 & data9set_OPC_DE_AD_Ctrl_markers$pct.2 > 0.01, ])

topGOResults.df <- cluster_GO(genesOfInterest, geneUniverse)

topGOResults_MF.df <- topGOResults.df[which(topGOResults.df$category == "MF" & topGOResults.df$Annotated < 500), ][c(1:7), ]
g1 <- scRNA_TopGO_plot2(topGOResults_MF.df)
g1 <- g1 + ggtitle("Molecular function")

topGOResults_BP.df <- topGOResults.df[which(topGOResults.df$category == "BP" & topGOResults.df$Annotated < 500), ][c(1:7), ]
g2 <- scRNA_TopGO_plot2(topGOResults_BP.df)
g2 <- g2 + ggtitle("Biological process")

grid.arrange(g1, g2, ncol = 2)
```

### 5. DE analysis

#### 5-1. AD vs Ctrl

```{r warning=FALSE, message=FALSE, fig.height=5, fig.width=6}
data9set_OPC_DE_AD_Ctrl_markers_padj <- data9set_OPC_DE_AD_Ctrl_markers[which(data9set_OPC_DE_AD_Ctrl_markers$p_val_adj < 0.01), ]

ggplot(data9set_OPC_DE_AD_Ctrl_markers) + geom_point(aes(x = avg_logFC, y = -log10(p_val_adj))) + 
  geom_point(data = data9set_OPC_DE_AD_Ctrl_markers_padj, aes(x = avg_logFC, y = -log10(p_val_adj)), color = "red") + theme_classic() +
  ggtitle("OPC DE analysis AD vs Ctrl") + 
   xlab("log2 fold change") + 
  ylab("-log10(adjusted p-value)") +
  theme(legend.position = "none",
        plot.title = element_text(size = rel(1.5), hjust = 0.5),
        axis.title = element_text(size = rel(1.25))) +
  geom_text_repel(aes(x = avg_logFC, y = -log10(p_val_adj), label = ifelse(-log10(data9set_OPC_DE_AD_Ctrl_markers$p_val_adj) > 3 , rownames(data9set_OPC_DE_AD_Ctrl_markers),""))) 

```

#### 5-2. AD vs ADp40KO

```{r warning=FALSE, message=FALSE, fig.height=5, fig.width=6}
data9set_OPC_DE_AD_ADp40KO_markers_padj <- data9set_OPC_DE_AD_ADp40KO_markers[which(data9set_OPC_DE_AD_ADp40KO_markers$p_val_adj < 0.01), ]

ggplot(data9set_OPC_DE_AD_ADp40KO_markers) + geom_point(aes(x = avg_logFC, y = -log10(p_val_adj))) + 
  geom_point(data = data9set_OPC_DE_AD_ADp40KO_markers_padj, aes(x = avg_logFC, y = -log10(p_val_adj)), color = "red") + theme_classic() +
  ggtitle("OPC DE analysis AD vs ADp40KO") + 
   xlab("log2 fold change") + 
  ylab("-log10(adjusted p-value)") +
  theme(legend.position = "none",
        plot.title = element_text(size = rel(1.5), hjust = 0.5),
        axis.title = element_text(size = rel(1.25))) +
  geom_text_repel(aes(x = avg_logFC, y = -log10(p_val_adj), label = ifelse(-log10(data9set_OPC_DE_AD_ADp40KO_markers$p_val_adj) > 3 , rownames(data9set_OPC_DE_AD_ADp40KO_markers),""))) 

```

### 6. GSEA

```{r warning=FALSE, message=FALSE}
GSEA_AD_Ctrl <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/GSEA_csv_files/Cell_type/20200221_Cell_type_DE_by_MAST_OPCAD_Ctrl.csv_GSEA_logFC_reactome.csv", row.names = 1)

GSEA_AD_ADp40KO <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/GSEA_csv_files/Cell_type/20200221_Cell_type_DE_by_MAST_OPCAD_ADp40KO.csv_GSEA_logFC_reactome.csv", row.names = 1)

GSEA_ADp40KO_Ctrl <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/GSEA_csv_files/Cell_type/20200221_Cell_type_DE_by_MAST_OPCADp40KO_Ctrl.csv_GSEA_logFC_reactome.csv", row.names = 1)
```

#### 6-1. AD vs Ctrl

In here, I only showed statistically significant pathways. 

```{r warning=FALSE, message=FALSE, fig.width = 7, fig.height=6}
ggplot(GSEA_AD_Ctrl[GSEA_AD_Ctrl$padj < 0.25, ], aes(reorder(pathway, NES), NES)) +
  geom_col(aes(fill=padj<0.05)) +
  coord_flip() +
  labs(title = "GSEA", y="Normalized Enrichment Score") + 
  theme_minimal() + theme(axis.title.y = element_blank(), axis.text.y = element_text(size = 10), legend.position = "none")

```

#### 6-2. AD vs ADp40KO

In here, I only showed statistically significant pathways. 

```{r warning=FALSE, message=FALSE, fig.width = 9, fig.height=8}
ggplot(GSEA_AD_Ctrl[GSEA_AD_ADp40KO$padj < 0.25, ], aes(reorder(pathway, NES), NES)) +
  geom_col(aes(fill=padj<0.05)) +
  coord_flip() +
  labs(title = "GSEA", y="Normalized Enrichment Score") + 
  theme_minimal() + theme(axis.title.y = element_blank(), axis.text.y = element_text(size = 10), legend.position = "none")

```

### 7. OPC cluster 38

```{r warning=FALSE, message=FALSE, fig.width = 6, fig.height=5}
data9set_cleaned_OPC_c38.SO <- subset(data9set_cleaned_OPC.SO, subset = seurat_clusters %in% c(38))


```

```{r warning=FALSE, message=FALSE, fig.height=5, fig.width=5}
# load subset of GO data (1st and 2nd column)
GO_myelination <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/GO_files/GO_0042552_myelination_sub.txt", sep = "\t")

# remove overlap genes 
myelination_genes <- as.vector(as.character(unique(GO_myelination$Symbol)))

# remove myelination genes not in the list
myelination_genes <- myelination_genes[myelination_genes %in% data9set_cleaned_OPC_c38.SO@assays$RNA@data@Dimnames[[1]]]

GO_myelination_excel <- read.xlsx("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/GO_files/GO_0042552_myelination.xlsx", 1, header=TRUE)

GO_myelination_excel_sub <- GO_myelination_excel[GO_myelination_excel$Symbol %in% myelination_genes, ]

##########################
# z-score for each sample
##########################
data9set_cleaned_OPC_c38_1.SO <- subset(data9set_cleaned_OPC_c38.SO, subset = gemgroup %in% "1")
data9set_cleaned_OPC_c38_2.SO <- subset(data9set_cleaned_OPC_c38.SO, subset = gemgroup %in% "2")
data9set_cleaned_OPC_c38_3.SO <- subset(data9set_cleaned_OPC_c38.SO, subset = gemgroup %in% "3")
data9set_cleaned_OPC_c38_4.SO <- subset(data9set_cleaned_OPC_c38.SO, subset = gemgroup %in% "4")
data9set_cleaned_OPC_c38_5.SO <- subset(data9set_cleaned_OPC_c38.SO, subset = gemgroup %in% "5")
data9set_cleaned_OPC_c38_6.SO <- subset(data9set_cleaned_OPC_c38.SO, subset = gemgroup %in% "6")
data9set_cleaned_OPC_c38_7.SO <- subset(data9set_cleaned_OPC_c38.SO, subset = gemgroup %in% "7")
data9set_cleaned_OPC_c38_8.SO <- subset(data9set_cleaned_OPC_c38.SO, subset = gemgroup %in% "8")
data9set_cleaned_OPC_c38_9.SO <- subset(data9set_cleaned_OPC_c38.SO, subset = gemgroup %in% "9")

# extract log(CPM+1) data from Seurat object
z_score_from_gene_list <- function(Seurat.SO, gene_list){
  
  df <- as.data.frame(Matrix::rowMeans(GetAssayData(Seurat.SO, slot = "data")))
  df$gene <- rownames(df)
  df <- df[rownames(df) %in% gene_list,]
  return(df)
}
```

```{r warning=FALSE, message=FALSE, fig.height=5, fig.width=5}
OPC1.df <- z_score_from_gene_list(data9set_cleaned_OPC_c38_1.SO, myelination_genes)
colnames(OPC1.df) <- c("Ctrl_1", "gene")
OPC2.df <- z_score_from_gene_list(data9set_cleaned_OPC_c38_2.SO, myelination_genes)
colnames(OPC2.df) <- c("ADp40KO_1", "gene")
OPC3.df <- z_score_from_gene_list(data9set_cleaned_OPC_c38_3.SO, myelination_genes)
colnames(OPC3.df) <- c("AD_1", "gene")
OPC4.df <- z_score_from_gene_list(data9set_cleaned_OPC_c38_4.SO, myelination_genes)
colnames(OPC4.df) <- c("Ctrl_2", "gene")
OPC5.df <- z_score_from_gene_list(data9set_cleaned_OPC_c38_5.SO, myelination_genes)
colnames(OPC5.df) <- c("ADp40KO_2", "gene")
OPC6.df <- z_score_from_gene_list(data9set_cleaned_OPC_c38_6.SO, myelination_genes)
colnames(OPC6.df) <- c("AD_2", "gene")
OPC7.df <- z_score_from_gene_list(data9set_cleaned_OPC_c38_7.SO, myelination_genes)
colnames(OPC7.df) <- c("Ctrl_3", "gene")
OPC8.df <- z_score_from_gene_list(data9set_cleaned_OPC_c38_8.SO, myelination_genes)
colnames(OPC8.df) <- c("ADp40KO_3", "gene")
OPC9.df <- z_score_from_gene_list(data9set_cleaned_OPC_c38_9.SO, myelination_genes)
colnames(OPC9.df) <- c("AD_3", "gene")

# merge all data frame
OPC.df <- merge(OPC1.df, c(OPC2.df, OPC3.df, OPC4.df, OPC5.df, OPC6.df, OPC7.df, OPC8.df, OPC9.df), by = "gene")
OPC.df <- OPC.df[,c(1,2,3,4, 6,8, 10, 12,14,16)]
rownames(OPC.df) <- OPC.df$gene
OPC.df$gene <- NULL

# remove if all counts is zero
OPC.df <- OPC.df[Matrix::rowSums(OPC.df) != 0, ]

#hist(rowSums(OPC.df), breaks = 30)

OPC.df2 <- OPC.df[rowSums(OPC.df) > 5, ]


# calculate z-score
OPC_zs.df <- apply(OPC.df2, 1, function(x) (x - mean(x)) / sd(x))
OPC_zs_final.df <- data.frame(t(OPC_zs.df))
```

```{r warning=FALSE, message=FALSE, fig.height=6, fig.width=7}
mypalette <- brewer.pal(11,"RdYlBu")
morecols <- colorRampPalette(mypalette)


heatmap.2(as.matrix(OPC_zs_final.df), 
          trace="none", 
          col=rev(morecols(50)),
          main="GO:0042552 myelination genes",
          scale="row", lhei=c(4, 10), lwid = c(3,10), cexCol=0.8)
```

```{r warning=FALSE, message=FALSE, fig.height=4.5, fig.width=7}
genes <- unique(as.character(GO_myelination_excel_sub[GO_myelination_excel_sub$Annotated.Term %in% "central nervous system myelination",]$Symbol))

heatmap.2(as.matrix(OPC_zs_final.df[rownames(OPC_zs_final.df) %in% genes, ]), 
          trace="none", 
          col=rev(morecols(50)),
          main="central nervous system myelination",
          scale="row", lhei=c(4, 8), lwid = c(2,10), cexCol=0.8, cexRow = 1, cex.main = 1)


```

```{r warning=FALSE, message=FALSE, fig.height=4.5, fig.width=7}
genes <- unique(as.character(GO_myelination_excel_sub[GO_myelination_excel_sub$Annotated.Term %in% "positive regulation of myelination",]$Symbol))

heatmap.2(as.matrix(OPC_zs_final.df[rownames(OPC_zs_final.df) %in% genes, ]), 
          trace="none", 
          col=rev(morecols(50)),
          main="positive regulation of myelination",
          scale="row", lhei=c(4, 8), lwid = c(2,10), cexCol=0.8, cexRow = 1, cex.main = 1)


```

```{r warning=FALSE, message=FALSE, fig.height=4.5, fig.width=7}
genes <- unique(as.character(GO_myelination_excel_sub[GO_myelination_excel_sub$Annotated.Term %in% "regulation of myelination",]$Symbol))

heatmap.2(as.matrix(OPC_zs_final.df[rownames(OPC_zs_final.df) %in% genes, ]), 
          trace="none", 
          col=rev(morecols(50)),
          main="regulation of myelination",
          scale="row", lhei=c(4, 8), lwid = c(2,10), cexCol=0.8, cexRow = 1, cex.main = 1)


```

```{r warning=FALSE, message=FALSE, fig.height=4.5, fig.width=7}
genes <- unique(as.character(GO_myelination_excel_sub[GO_myelination_excel_sub$Annotated.Term %in% "myelination",]$Symbol))

heatmap.2(as.matrix(OPC_zs_final.df[rownames(OPC_zs_final.df) %in% genes, ]), 
          trace="none", 
          col=rev(morecols(50)),
          main="myelination",
          scale="row", lhei=c(4, 8), lwid = c(2,10), cexCol=0.8, cexRow = 1, cex.main = 1)


```

```{r warning=FALSE, message=FALSE, fig.height=4.5, fig.width=7}
DE_C38_AD_Ctrl <- FindMarkers(data9set_cleaned_OPC_c38.SO, ident.1 =  "AD", ident.2 = "Ctrl", group.by = "sample", test.use = "MAST", logfc.threshold = 0.1)
DE_C38_AD_Ctrl$gene <- rownames(DE_C38_AD_Ctrl)


```

```{r warning=FALSE, message=FALSE, fig.height=4.5, fig.width=7}
DE_C38_AD_Ctrl_m <- DE_C38_AD_Ctrl[DE_C38_AD_Ctrl$gene %in% myelination_genes, ]


# log2FC > 1 or padj > 50
#DE_C38_AD_Ctrl_plus.df <- DE_C38_AD_Ctrl[which(DE_C38_AD_Ctrl$avg_logFC > 0 & DE_C38_AD_Ctrl$p_val_adj < 0.01),]
#OL_DE_AD_Ctrl_minus.df <- OL_DE_AD_Ctrl.df[which(OL_DE_AD_Ctrl.df$avg_logFC < 0 & OL_DE_AD_Ctrl.df$p_val_adj < 0.01),]

```

```{r warning=FALSE, message=FALSE, fig.height=4.5, fig.width=7}
ggplot(DE_C38_AD_Ctrl_m) + geom_point(aes(x = avg_logFC, y = -log10(p_val_adj))) + 
  #geom_point(data = OL_DE_AD_ADp40KO_plus.df, aes(x = avg_logFC, y = -log10(p_val_adj)), color = "red") +
  ggtitle("myelination genes in AD versus ADp40KO of Oligo") + 
  theme_classic() + 
  xlab("log2 fold change") + 
  ylab("-log10(adjusted p-value)") +
  scale_y_continuous(expand = c(0,0), limits = c(0,5)) + 
  theme(legend.position = "none",
        plot.title = element_text(size = 15, hjust = 0.5, family = "helvetica"),
        axis.title = element_text(size = 15), 
        axis.text = element_text(size = 12, color = "black"))
```

```{r warning=FALSE, message=FALSE, fig.height=4.5, fig.width=7}
VlnPlot(data9set_cleaned_OPC_c38.SO, features = "Nrg1", group.by = "sample")

```

```{r warning=FALSE, message=FALSE, fig.height=4.5, fig.width=7}
VlnPlot(data9set_cleaned_OPC_c38.SO, features = "Gnb4", group.by = "sample")

```



```{r warning=FALSE, message=FALSE}
sessionInfo()
```
