---
title: "Microglia_sub_analysis"
author: "Skim"
date: "1/9/2020"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_depth: 4
    toc_float: false
    code_folding: hide
---

```{r warning=FALSE, message=FALSE}
Sys.setenv(PYTHONPATH="/data/local/rajewsky/shared_libs/python")
library(reticulate)
#use_python("/usr/bin/python3", required = TRUE)
library(Seurat)
library(ggplot2)
library(dplyr)
library(DT)
library(gridExtra)
library(nichenetr)
library(tidyverse)
library(irr)
library(fgsea)
library(MAST)
library(ggrepel)
library(biomaRt)
library(topGO)
source("/data/rajewsky/home/skim/Microglia_Heppner/Codes/DroNc_AD/src/TopGO_scRNA.R")
#Rscript -e "library(rmarkdown); rmarkdown::render('/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20191229_pre_analysis_9sets.Rmd')"
```

```{r warning=FALSE, message=FALSE}
data9set <- Read10X("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/SP064_All/outs/filtered_feature_bc_matrix/")
load("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200104_pre_analysis_9sets_Seurat_object_res0.8.Robj")
```

## 1. MG cluster analysis

This analysis is done by subset 4 Microglia clusters (from res 0.8) and I investigate more in here. 

```{r warning=FALSE, message=FALSE, fig.width= 12, fig.height=5}
# subset Microglia population
# resoltuion 0.6
# data9set.SO_MG <- subset(data9set.SO, subset = seurat_clusters == c(3, 29, 32))

# resolution 0.8
# 3, 8, 34, 37
data9set.SO_MG <- subset(data9set.SO, subset = seurat_clusters == c(3, 8, 34, 37))

d1 <- DimPlot(object = data9set.SO_MG, reduction = "umap")

data9set.SO_MG@meta.data$sample <- ifelse((data9set.SO_MG$gemgroup == 1 | data9set.SO_MG$gemgroup == 4 | data9set.SO_MG$gemgroup == 7), "Ctrl", ifelse((data9set.SO_MG$gemgroup == 2 | data9set.SO_MG$gemgroup == 5 | data9set.SO_MG$gemgroup == 8), "AD", "ADp40KO"))

data9set.SO_MG@meta.data$sample <- factor(data9set.SO_MG@meta.data$sample, levels = c("Ctrl", "AD", "ADp40KO"))
d2 <- DimPlot(object = data9set.SO_MG, group.by = "sample", pt.size = 0.05, cols = c("#003333", "#FF6600", "#FFFF00"))
grid.arrange(d1, d2, ncol = 2)
```

```{r warning=FALSE, message=FALSE}
data9set.MG_markers <- FindAllMarkers(data9set.SO_MG, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
data9set.MG_markers_top10 <- data9set.MG_markers %>% group_by(cluster) %>% top_n(n = 20, wt = avg_logFC)
datatable(data9set.MG_markers_top10, options = list(pageLength = 10, dom = 'tip'))
```

```{r warning=FALSE, message=FALSE, fig.width= 7, fig.height=8}
FeaturePlot(data9set.SO_MG, features = c("Hexb", "Zfhx3", "Ank", "Myo1e", "St18", "F13a1"), pt.size = 0.1)
```

#### 1-2. DE analysis

In here, I compare DE of cluster 8 vs 3 and also I implemented GSEA and GO analysis next.  

red dots are padj < 0.01

```{r warning=FALSE, message=FALSE, fig.width= 7, fig.height=5}
#data9set.SO_MG.marker8vs3 <- FindMarkers(data9set.SO_MG, ident.1 =  "8", ident.2 = "3", group.by = "seurat_clusters", test.use = "MAST", logfc.threshold = 0, min.pct = 0.01)
#write.csv(data9set.SO_MG.marker8vs3, "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200109_MG_sub_analysis.MG_marker.csv")
data9set.SO_MG.marker8vs3 <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200109_MG_sub_analysis.MG_marker.csv", row.names = 1)
data9set.SO_MG.marker8vs3_padj <- data9set.SO_MG.marker8vs3[which(data9set.SO_MG.marker8vs3$p_val_adj < 0.01), ]

ggplot(data9set.SO_MG.marker8vs3) + geom_point(aes(x = avg_logFC, y = -log10(p_val_adj))) + 
  geom_point(data = data9set.SO_MG.marker8vs3_padj, aes(x = avg_logFC, y = -log10(p_val_adj)), color = "red") +
  ggtitle("DE analysis cluster 8 vs 3") + 
   xlab("log2 fold change") + 
  ylab("-log10 adjusted p-value") +
  theme(legend.position = "none",
        plot.title = element_text(size = rel(1.5), hjust = 0.5),
        axis.title = element_text(size = rel(1.25))) +
  geom_text_repel(aes(x = avg_logFC, y = -log10(p_val_adj), label = ifelse(-log10(data9set.SO_MG.marker8vs3$p_val_adj) > 30 , rownames(data9set.SO_MG.marker8vs3),"")))
 
```

#### 1-3. GSEA analysis

Rank based on **logFC**

```{r warning=FALSE, message=FALSE}
# loading pathways
pathways.hallmark <- gmtPathways("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/ext_files/ReactomePathways.gmt")
pathways.hallmark2 <- gmtPathways("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/ext_files/h.all.v7.0.symbols.gmt")
```

```{r warning=FALSE, message=FALSE}
data9set.SO_MG.marker8vs3$Gene <- rownames(data9set.SO_MG.marker8vs3)

# HUMAN and MOUSE
mart <- useDataset("mmusculus_gene_ensembl", mart=useMart("ensembl"))

# extract human associated genes (cannot extract both mgi_symbol and hsapiens_homolog_associated_gene_name)
bm <- getBM(attributes=c("ensembl_gene_id", "hsapiens_homolog_associated_gene_name"),  
            filters = 'mgi_symbol',
            values = rownames(data9set.SO_MG.marker8vs3),
            mart=mart)

bm <- bm %>% distinct() %>% as_tibble() %>% na_if("") %>% na.omit() %>% as.data.frame()
bm.df <- data.frame(ensembl_gene_id= bm$ensembl_gene_id, hsapiens_homolog_associated_gene_name = bm$hsapiens_homolog_associated_gene_name, stringsAsFactors = FALSE)

# extract mouse genes
bm2 <- getBM(attributes=c("ensembl_gene_id", "mgi_symbol"),  
            filters = 'mgi_symbol',
            values = rownames(data9set.SO_MG.marker8vs3),
            mart=mart)

bm2.df <- as.data.frame(bm2)

#colnames(bm2.df) %in% colnames(bm.df)
#names(bm.df)
#names(bm.df) <- gsub("\\s"," ",names(bm.df))
#names(bm.df) <- names(bm.df) %>% stringr::str_replace_all("\\s","_")

total <- merge(bm.df,bm2.df, by = "ensembl_gene_id")
#total$mgi_symbol <- toupper(total$mgi_symbol) 
#total$sanity <- ifelse(total$hsapiens_homolog_associated_gene_name == total$mgi_symbol, "same", "diff")
total <- total[, c(2,3)]
colnames(total) <- c("HGene", "Gene")

data9set.SO_MG.marker8vs3_all <- inner_join(data9set.SO_MG.marker8vs3, total, by = "Gene")
```

Only Hallmark pathway showed significant pathways while reactomes didn't show any pathways significant. 

```{r warning=FALSE, message=FALSE, fig.width= 8, fig.height=2}
# Creating rnk file
data9set.SO_MG.marker8vs3_all$metric <-  data9set.SO_MG.marker8vs3_all$avg_logFC

MG_8_3_metric <- data9set.SO_MG.marker8vs3_all[,c("HGene", "metric")]
ranks <- tibble::deframe(MG_8_3_metric)
fgseaRes <- fgsea(pathways.hallmark, ranks, minSize=10, maxSize=500, nperm=100000)
fgseaRes2 <- fgsea(pathways.hallmark2, ranks, minSize=15, maxSize=500, nperm=100000)

fgseaRes2 <- fgseaRes2[which(fgseaRes2$padj < 0.05), ]

ggplot(fgseaRes2, aes(reorder(pathway, NES), NES)) +
  geom_col(aes(fill=padj<0.05)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score") + 
  theme_minimal() + theme(axis.text.y = element_text(size = 10))
```

Rank based on sign(log2FC) x -log10(p_val_adj)

```{r warning=FALSE, message=FALSE, fig.width= 10, fig.height=10}
# Creating rnk file
data9set.SO_MG.marker8vs3_all$fcsign <- sign(data9set.SO_MG.marker8vs3_all$avg_logFC)
data9set.SO_MG.marker8vs3_all$logP = -log10(data9set.SO_MG.marker8vs3_all$p_val_adj)
data9set.SO_MG.marker8vs3_all$metric= data9set.SO_MG.marker8vs3_all$logP/data9set.SO_MG.marker8vs3_all$fcsign

MG_8_3_metric2 <- data9set.SO_MG.marker8vs3_all[,c("HGene", "metric")]
ranks2 <- tibble::deframe(MG_8_3_metric2)

fgseaRes3 <- fgsea(pathways.hallmark, ranks2, minSize=10, maxSize=500, nperm=100000)
fgseaRes4 <- fgsea(pathways.hallmark2, ranks2, minSize=15, maxSize=500, nperm=100000)

#library(future)
#plan(strategy = "multicore", workers = 6)

#topUp <- fgseaRes %>% 
#    tidygraph::filter(ES > 0) %>% 
#    top_n(10, wt=-padj)
#topDown <- fgseaRes %>% 
#    filter(ES < 0) %>% 
#    top_n(10, wt=-padj)
#topPathways <- bind_rows(topUp, topDown) %>% 
#    arrange(-ES)
#datatable(topPathways, options = list(pageLength = 5, dom = 'tip'))
#barplot(sort(ranks, decreasing = T))

ggplot(fgseaRes4, aes(reorder(pathway, NES), NES)) +
  geom_col(aes(fill=padj<0.05)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score") + 
  theme_minimal() + theme(axis.text.y = element_text(size = 10))
```

## 1-4. GO analysis

genes: avg_log_FC > 1 & padj < 0.01

22 upregulated genes and 3 down regulated genes, algorithm = elim, plot here shows BP

```{r warning=FALSE, message=FALSE, fig.width= 12, fig.height=8}
# Background genes
GO_bg_genes <- rownames(data9set.SO_MG.marker8vs3)

# log2FC up > 1 & padj < 0.01
GO_up_genes <- rownames(data9set.SO_MG.marker8vs3[which(data9set.SO_MG.marker8vs3$avg_logFC > 1 & data9set.SO_MG.marker8vs3$p_val_adj < 0.01), ])

# log2FC up < -1 & padj < 0.01
GO_down_genes <- rownames(data9set.SO_MG.marker8vs3[which(data9set.SO_MG.marker8vs3$avg_logFC < -1 & data9set.SO_MG.marker8vs3$p_val_adj < 0.01), ])


# GO Analysis 
# I create scRNA_topGO function
up_GO <- scRNA_topGO(fg.genes = GO_up_genes, bg.genes = GO_bg_genes)
down_GO <- scRNA_topGO(fg.genes = GO_down_genes, bg.genes = GO_bg_genes)



# GO plot function
BP_GO_plot_table <- function(df){
  df <- df[c(21:40), ]
  df$pval <- as.numeric(df$pval)
  df$PValue <- -log10(df$pval)
  df$Term <- factor(df$Term, levels = rev(unique(df$Term)))
  return(df)
}
up_GO.df <- BP_GO_plot_table(up_GO)
down_GO.df <- BP_GO_plot_table(down_GO)


g1 <- ggplot(up_GO.df, aes(x = Term, y = PValue)) + geom_bar(stat="identity", color = "chocolate2", fill = "chocolate2") + coord_flip() +
  ylab("-log10(p-value)") + ggtitle("upregulate 8 vs 3") + 
    theme(axis.title.y = element_blank(), axis.text = element_text(color = "black", size = 8), 
        axis.title.x = element_text(size = 8, color = "black"), title = element_text(size = 8),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), 
        axis.line = element_line(colour = "black"), 
        axis.ticks.y = element_blank())
g2 <- ggplot(down_GO.df, aes(x = Term, y = PValue)) + geom_bar(stat="identity", color = "steelblue", fill = "steelblue") + coord_flip() +
  ylab("-log10(p-value)") + ggtitle("downregulated 8 vs 3") + 
    theme(axis.title.y = element_blank(), axis.text = element_text(color = "black", size = 8), 
        axis.title.x = element_text(size = 8, color = "black"), title = element_text(size = 8),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), 
        axis.line = element_line(colour = "black"), 
        axis.ticks.y = element_blank())

grid.arrange(g1, g2, ncol = 2)

```

Annotated : number of genes which are annotated with the GO-term.  
Significant : number of genes belonging to your **input** which are annotated with the GO-term.  
Expected : show an estimate of the number of genes a node of size Annotated would have if the significant genes were to be randomly selected from the gene universe.  
pvalues : pvalue obtained after the test  


```{r warning=FALSE, message=FALSE, fig.width= 12, fig.height=8}
g1 <- ggplot(up_GO.df, aes(x=Significant, 
               y=Term, 
               colour=pval, 
               size=Annotated)) +
        geom_point() +
        expand_limits(x=0) +
        labs(x="Significant", y="GO term", colour="p value", size="Annotated")
g2 <- ggplot(down_GO.df, aes(x=Significant, 
               y=Term, 
               colour=pval, 
               size=Annotated)) +
        geom_point() +
        expand_limits(x=0) +
        labs(x="Significant", y="GO term", colour="p value", size="Annotated")
grid.arrange(g1, g2, ncol = 2)

```
## 1-5. Cell population cluters

```{r warning=FALSE, message=FALSE, fig.width= 8, fig.height=8}
##################
# per experiment
##################
exp <- c("Ctrl", "AD", "ADp40KO")
# Create new data frame
exp_cell.df <- data.frame(ident = as.character(), cluster = as.numeric(), cells = as.numeric(), stringsAsFactors = FALSE)

index <- c(3, 8, 34, 37)
for(i in exp){
  for(j in index){
    if(i == "Ctrl"){
      k <- which(index == j)
    }
    if(i == "AD"){
      k <- which(index == j) + 4
    }
    if(i =="ADp40KO"){
      k <- which(index == j) + 8
    }
    exp_cell.df[k, 1] <- i
    exp_cell.df[k, 2] <- j
    exp_cell.df[k, 3] <- table(data9set.SO_MG@meta.data$sample == i & data9set.SO_MG@meta.data$seurat_clusters == j)[2]
    #exp_cell.df[is.na(exp_cell.df)] <- 0
  }
}

#sum(exp_cell.df[exp_cell.df$ident == "AD", ]$cells)

# calculate percentage
exp_cell.df$total <- ifelse(exp_cell.df$ident == "Ctrl", sum(exp_cell.df[exp_cell.df$ident == "Ctrl", ]$cells), 
                            ifelse(exp_cell.df$ident == "AD", sum(exp_cell.df[exp_cell.df$ident == "AD", ]$cells),
                                   sum(exp_cell.df[exp_cell.df$ident == "AD", ]$cells)))
exp_cell.df$percentage <- round(100*(exp_cell.df$cells/exp_cell.df$total),2)

ggplot(exp_cell.df, aes(x = ident, y = cells, fill = factor(cluster)))+
  geom_bar(position="fill", stat="identity")
```

## 2. Subclustering

HVGs = 100  
PCs 1:5  
res = 0.4

```{r warning=FALSE, message=FALSE, fig.width= 6, fig.height=5}
# 
data9set_MG <- data9set[, data9set@Dimnames[[2]] %in% names(data9set.SO_MG@active.ident)]
data9set_MG.SO <- CreateSeuratObject(counts =  data9set_MG,  min.cells = 3, min.features = 200, project = "ADp40KO")
# add gemgroup to meta data
gemgroup <- sapply(strsplit(rownames(data9set_MG.SO@meta.data), split="-"), "[[", 2)
data9set_MG.SO <- AddMetaData(object=data9set_MG.SO, metadata=data.frame(gemgroup=gemgroup, row.names=rownames(data9set_MG.SO@meta.data)))

# calculate MT genes
data9set_MG.SO[["percent.mt"]] <- PercentageFeatureSet(object = data9set_MG.SO, pattern = "^mt-")
data9set_MG.SO <- subset(x = data9set_MG.SO, subset = nCount_RNA < 30000 & nCount_RNA > 200 & percent.mt < 5)  # 3787 -> 2722

data9set_MG.SO <- NormalizeData(object = data9set_MG.SO, normalization.method = "LogNormalize", scale.factor = 1e4, verbose = FALSE)

data9set_MG.SO <- FindVariableFeatures(object = data9set_MG.SO, nfeatures = 100, selection.method = "vst", verbose = FALSE)
data9set_MG.SO <- ScaleData(object = data9set_MG.SO, vars.to.regress = c("percent.mt", "nCount_RNA", "gemgroup"))
data9set_MG.SO <- RunPCA(data9set_MG.SO, features = VariableFeatures(object = data9set_MG.SO), verbose = FALSE, ndims.print = "None")
#ElbowPlot(data9set_MG.SO, ndims = 20)
data9set_MG.SO <- FindNeighbors(object = data9set_MG.SO, reduction = "pca", dims = 1:5)
data9set_MG.SO <- FindClusters(object = data9set_MG.SO, resolution = 0.4, verbose = FALSE)

#data9set_MG.SO <- RunTSNE(object = data9set_MG.SO, dims = 1:3, nthreads = 8, perplexity = 30)
data9set_MG.SO <- RunUMAP(object = data9set_MG.SO, dims = 1:5, n.neighbors = 10, min.dist = 0.35, n.epochs = 500, spread = 1)
DimPlot(object = data9set_MG.SO, reduction = "umap")
#DimPlot(object = data9set_MG.SO, reduction = "tsne")
```

```{r warning=FALSE, message=FALSE, fig.width= 6, fig.height=5}
#DimPlot(object = data9set_MG.SO, group.by = "gemgroup", pt.size = 0.05)

data9set_MG.SO@meta.data$sample <- ifelse((data9set_MG.SO$gemgroup == 1 | data9set_MG.SO$gemgroup == 4 | data9set_MG.SO$gemgroup == 7), "Ctrl", ifelse((data9set_MG.SO$gemgroup == 2 | data9set_MG.SO$gemgroup == 5 | data9set_MG.SO$gemgroup == 8), "AD", "ADp40KO"))

data9set_MG.SO@meta.data$sample <- factor(data9set_MG.SO@meta.data$sample, levels = c("Ctrl", "AD", "ADp40KO"))
#table(data9set.SO$sample)
DimPlot(object = data9set_MG.SO, group.by = "sample", pt.size = 0.5, cols = c("#003333", "#FF6600", "#FFFF00"))

```

```{r warning=FALSE, message=FALSE, fig.width= 10, fig.height=4}
DimPlot(data9set_MG.SO, reduction = "umap", split.by = "sample", no.legend = TRUE)
```

```{r warning=FALSE, message=FALSE, fig.width= 7, fig.height=8}
FeaturePlot(data9set_MG.SO, features = c("Hexb", "Zfhx3", "Ank", "Myo1e", "St18", "F13a1"), pt.size = 0.1)
```

```{r warning=FALSE, message=FALSE}
#data9set_MG.markers <- FindAllMarkers(data9set_MG.SO, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0)
#data9set_MG.markers_top10 <- data9set_MG.markers %>% group_by(cluster) %>% top_n(n = 20, wt = avg_logFC)
#datatable(data9set.markers_top10, options = list(pageLength = 10, dom = 'tip'))
#write.csv(data9set.markers, "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20191229_data9set_marker.csv")
#write.csv(data9set.markers_top10, "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20191229_data9set_marker_top20.csv")
#data9set.markers_top10 <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20191229_data9set_marker_top20.csv")
#datatable(data9set.markers_top10, options = list(pageLength = 10, dom = 'tip'))
```

```{r warning=FALSE, message=FALSE}
sessionInfo()

```

