---
title: "CCI_test_MG_Oligo_check"
author: "Skim"
date: "1/31/2020"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_depth: 4
    toc_float: false
    code_folding: hide
---

In here, I consider swapped beteen AD and ADp40KO (Before changing)

```{r warning=FALSE, message=FALSE}
library(ggrepel)
library(ggplot2)
```

## 1. Cluster 1 (Oligo) ADp40KO vs Ctrl

```{r warning=FALSE, message=FALSE}
# DE
DE_C1_AD_Ctrl <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/DE_csv_files/20200117_Clustering_DE_by_MAST_each_Cluster_1AD_Ctrl.csv", row.names = 1)

# GO
GO_C1_AD_Ctrl <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/GO_csv_files/20200117_Clustering_DE_by_MAST_each_Cluster_1AD_Ctrl.csv_GO_BP_only.csv", row.names = 1)

# GSEA
GSEA_C1_AD_Ctrl_logFC_hall <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/GSEA_csv_files/20200117_Clustering_DE_by_MAST_each_Cluster_1AD_Ctrl.csv_GSEA_logFC_hall.csv", row.names = 1)
GSEA_C1_AD_Ctrl_logFC_reactome <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/GSEA_csv_files/20200117_Clustering_DE_by_MAST_each_Cluster_1AD_Ctrl.csv_GSEA_logFC_reactome.csv", row.names = 1)
GSEA_C1_AD_Ctrl_padj_hall <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/GSEA_csv_files/20200117_Clustering_DE_by_MAST_each_Cluster_1AD_Ctrl.csv_GSEA_padj_hall.csv", row.names = 1)
GSEA_C1_AD_Ctrl_padj_reactome <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/GSEA_csv_files/20200117_Clustering_DE_by_MAST_each_Cluster_1AD_Ctrl.csv_GSEA_padj_reactome.csv", row.names = 1)
```

```{r warning=FALSE, message=FALSE, fig.width= 6, fig.height= 5}
DE_figure <- function(DE.df, text = "something"){
  DE_padj <- DE.df[which(DE.df$p_val_adj < 0.01), ]
  
  ggplot(DE.df) + geom_point(aes(x = avg_logFC, y = -log10(p_val_adj))) + 
  geom_point(data = DE_padj, aes(x = avg_logFC, y = -log10(p_val_adj)), color = "red") +
  ggtitle(text) + 
   xlab("log2 fold change") + 
  ylab("-log10 adjusted p-value") +
  theme(legend.position = "none",
        plot.title = element_text(size = rel(1.5), hjust = 0.5),
        axis.title = element_text(size = rel(1.25))) +
  geom_text_repel(aes(x = avg_logFC, y = -log10(p_val_adj), label = ifelse(-log10(DE.df$p_val_adj) > 20 , rownames(DE.df),"")))

}

GO_figure <- function(GO.df){
  GO.df <- GO.df[c(1:10), ]
  GO.df$Term <- factor(GO.df$Term, levels = GO.df$Term[order(GO.df$Fisher.elim, decreasing = TRUE)])

  ggplot(GO.df, aes(x=gene_ratio, 
               y=Term, 
               colour=Fisher.elim, 
               size=Significant)) +
        geom_point() +
        expand_limits(x=0) +
        labs(x="gene ratio", y="GO term", colour="p value", size="Significant")
}

GSEA_Hall_figure <- function(GSEA.df){
  ggplot(GSEA.df, aes(reorder(pathway, NES), NES)) +
  geom_col(aes(fill=padj<0.05)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score") + 
  theme_minimal() + theme(axis.text.y = element_text(size = 7))
}

GSEA_reactome_figure <- function(GSEA.df){
  GSEA.df <- GSEA.df[which(GSEA.df$padj < 0.1), ]

  ggplot(GSEA.df, aes(reorder(pathway, NES), NES)) +
  geom_col(aes(fill=padj<0.05)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score") + 
  theme_minimal() + theme(axis.text.y = element_text(size = 7))
}

```

#### 1-2. Cluster 1 (Oligo) ADp40KO vs Ctrl DE results

```{r warning=FALSE, message=FALSE, fig.width= 6, fig.height= 5}
DE_figure(DE_C1_AD_Ctrl, text = "DE analysis cluster 1 ADp40KO vs Ctrl")
```

#### 1-3. Cluster 1 (Oligo) ADp40KO vs Ctrl GO results

```{r warning=FALSE, message=FALSE, fig.width= 7, fig.height= 4}
GO_figure(GO_C1_AD_Ctrl)
```

#### 1-4. Cluster 1 (Oligo) ADp40KO vs Ctrl GSEA results logFC

```{r warning=FALSE, message=FALSE, fig.width= 8, fig.height= 6}
GSEA_Hall_figure(GSEA_C1_AD_Ctrl_logFC_hall)
```

#### 1-5. Cluster 1 (Oligo) ADp40KO vs Ctrl GSEA results logFC

```{r warning=FALSE, message=FALSE, fig.width= 8, fig.height= 4}
GSEA_reactome_figure(GSEA_C1_AD_Ctrl_logFC_reactome)
```

GSEA with padj didn't give any significant results. 

## 2. Cluster 1 (Oligo) AD vs Ctrl

```{r warning=FALSE, message=FALSE}
# DE
DE_C1_ADp40KO_Ctrl <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/DE_csv_files/20200117_Clustering_DE_by_MAST_each_Cluster_1ADp40KO_Ctrl.csv", row.names = 1)

# GO
GO_C1_ADp40KO_Ctrl <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/GO_csv_files/20200117_Clustering_DE_by_MAST_each_Cluster_1ADp40KO_Ctrl.csv_GO_BP_only.csv", row.names = 1)

# GSEA
GSEA_C1_ADp40KO_Ctrl_logFC_hall <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/GSEA_csv_files/20200117_Clustering_DE_by_MAST_each_Cluster_1ADp40KO_Ctrl.csv_GSEA_logFC_hall.csv", row.names = 1)
GSEA_C1_ADp40KO_Ctrl_logFC_reactome <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/GSEA_csv_files/20200117_Clustering_DE_by_MAST_each_Cluster_1ADp40KO_Ctrl.csv_GSEA_logFC_reactome.csv", row.names = 1)
GSEA_C1_ADp40KO_Ctrl_padj_hall <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/GSEA_csv_files/20200117_Clustering_DE_by_MAST_each_Cluster_1ADp40KO_Ctrl.csv_GSEA_padj_hall.csv", row.names = 1)
GSEA_C1_ADp40KO_Ctrl_padj_reactome <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/GSEA_csv_files/20200117_Clustering_DE_by_MAST_each_Cluster_1ADp40KO_Ctrl.csv_GSEA_padj_reactome.csv", row.names = 1)
```

#### 2-2. Cluster 1 (Oligo) AD vs Ctrl DE results

```{r warning=FALSE, message=FALSE, fig.width= 6, fig.height= 5}
DE_figure(DE_C1_ADp40KO_Ctrl, text = "DE analysis cluster 1 AD vs Ctrl")
```

#### 2-3. Cluster 1 (Oligo) AD vs Ctrl GO results

```{r warning=FALSE, message=FALSE, fig.width= 7, fig.height= 4}
GO_figure(GO_C1_ADp40KO_Ctrl)
```

#### 2-4. Cluster 1 (Oligo) AD vs Ctrl GSEA results logFC

```{r warning=FALSE, message=FALSE, fig.width= 8, fig.height= 6}
GSEA_Hall_figure(GSEA_C1_ADp40KO_Ctrl_logFC_hall)
```

#### 2-5. Cluster 1 (Oligo) AD vs Ctrl GSEA results logFC

```{r warning=FALSE, message=FALSE, fig.width= 8, fig.height= 4}
GSEA_reactome_figure(GSEA_C1_ADp40KO_Ctrl_logFC_reactome)
```

## 3. Cluster 1 (Oligo) ADp40KO vs AD

```{r warning=FALSE, message=FALSE}
# DE
DE_C1_ADp40KO_AD <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/DE_csv_files/20200117_Clustering_DE_by_MAST_each_Cluster_1AD_ADp40KO.csv", row.names = 1)

# GO
GO_C1_ADp40KO_AD <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/GO_csv_files/20200117_Clustering_DE_by_MAST_each_Cluster_1AD_ADp40KO.csv_GO_BP_only.csv", row.names = 1)

# GSEA
GSEA_C1_ADp40KO_AD_logFC_hall <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/GSEA_csv_files/20200117_Clustering_DE_by_MAST_each_Cluster_1AD_ADp40KO.csv_GSEA_logFC_hall.csv", row.names = 1)
GSEA_C1_ADp40KO_AD_logFC_reactome <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/GSEA_csv_files/20200117_Clustering_DE_by_MAST_each_Cluster_1AD_ADp40KO.csv_GSEA_logFC_reactome.csv", row.names = 1)
```

#### 3-2. Cluster 1 (Oligo) ADp40KO vs AD DE results

```{r warning=FALSE, message=FALSE, fig.width= 6, fig.height= 5}
DE_figure(DE_C1_ADp40KO_AD, text = "DE analysis cluster 1 ADp40KO vs AD")
```

#### 3-3. Cluster 1 (Oligo) ADp40KO vs AD GO results

```{r warning=FALSE, message=FALSE, fig.width= 7, fig.height= 4}
GO_figure(GO_C1_ADp40KO_AD)
```

#### 3-4. Cluster 1 (Oligo) ADp40KO vs AD GSEA results logFC

```{r warning=FALSE, message=FALSE, fig.width= 8, fig.height= 6}
GSEA_Hall_figure(GSEA_C1_ADp40KO_AD_logFC_hall)
```

#### 3-5. Cluster 1 (Oligo) ADp40KO vs AD GSEA results logFC

```{r warning=FALSE, message=FALSE, fig.width= 8, fig.height= 6}
GSEA_reactome_figure(GSEA_C1_ADp40KO_AD_logFC_reactome)
```

## 4. Cluster 3 (Microglia) ADp40KO vs Ctrl

```{r warning=FALSE, message=FALSE}
# DE
DE_C3_AD_Ctrl <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/DE_csv_files/20200117_Clustering_DE_by_MAST_each_Cluster_3AD_Ctrl.csv", row.names = 1)

# GO
GO_C3_AD_Ctrl <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/GO_csv_files/20200117_Clustering_DE_by_MAST_each_Cluster_3AD_Ctrl.csv_GO_BP_only.csv", row.names = 1)

# GSEA
GSEA_C3_AD_Ctrl_logFC_hall <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/GSEA_csv_files/20200117_Clustering_DE_by_MAST_each_Cluster_3AD_Ctrl.csv_GSEA_logFC_hall.csv", row.names = 1)
GSEA_C3_AD_Ctrl_logFC_reactome <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/GSEA_csv_files/20200117_Clustering_DE_by_MAST_each_Cluster_3AD_Ctrl.csv_GSEA_logFC_reactome.csv", row.names = 1)
```

#### 4-2. Cluster 3 (MG) ADp40KO vs Ctrl DE results

```{r warning=FALSE, message=FALSE, fig.width= 6, fig.height= 5}
DE_figure(DE_C3_AD_Ctrl, text = "DE analysis cluster 3 ADp40KO vs Ctrl")
```

#### 4-3. Cluster 3 (MG) ADp40KO vs Ctrl GO results

```{r warning=FALSE, message=FALSE, fig.width= 7, fig.height= 4}
GO_figure(GO_C3_AD_Ctrl)
```

#### 4-4. Cluster 3 (MG) ADp40KO vs Ctrl GSEA results logFC

```{r warning=FALSE, message=FALSE, fig.width= 8, fig.height= 6}
GSEA_Hall_figure(GSEA_C3_AD_Ctrl_logFC_hall)
```

#### 4-5. Cluster 3 (MG) ADp40KO vs Ctrl GSEA results logFC

```{r warning=FALSE, message=FALSE, fig.width= 8, fig.height= 4}
GSEA_reactome_figure(GSEA_C3_AD_Ctrl_logFC_reactome)
```

## 5. Cluster 3 (MG) AD vs Ctrl

```{r warning=FALSE, message=FALSE}
# DE
DE_C3_ADp40KO_Ctrl <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/DE_csv_files/20200117_Clustering_DE_by_MAST_each_Cluster_3ADp40KO_Ctrl.csv", row.names = 1)

# GO
GO_C3_ADp40KO_Ctrl <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/GO_csv_files/20200117_Clustering_DE_by_MAST_each_Cluster_3ADp40KO_Ctrl.csv_GO_BP_only.csv", row.names = 1)

# GSEA
GSEA_C3_ADp40KO_Ctrl_logFC_hall <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/GSEA_csv_files/20200117_Clustering_DE_by_MAST_each_Cluster_3ADp40KO_Ctrl.csv_GSEA_logFC_hall.csv", row.names = 1)
GSEA_C3_ADp40KO_Ctrl_logFC_reactome <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/GSEA_csv_files/20200117_Clustering_DE_by_MAST_each_Cluster_3ADp40KO_Ctrl.csv_GSEA_logFC_reactome.csv", row.names = 1)
```

#### 5-2. Cluster 3 (MG) AD vs Ctrl DE results

```{r warning=FALSE, message=FALSE, fig.width= 6, fig.height= 5}
DE_figure(DE_C3_ADp40KO_Ctrl, text = "DE analysis cluster 1 AD vs Ctrl")
```

#### 5-3. Cluster 3 (MG) AD vs Ctrl GO results

```{r warning=FALSE, message=FALSE, fig.width= 7, fig.height= 4}
GO_figure(GO_C3_ADp40KO_Ctrl)
```

#### 5-4. Cluster 3 (MG) AD vs Ctrl GSEA results logFC

```{r warning=FALSE, message=FALSE, fig.width= 8, fig.height= 6}
GSEA_Hall_figure(GSEA_C3_ADp40KO_Ctrl_logFC_hall)
```

#### 5-5. Cluster 3 (MG) AD vs Ctrl GSEA results logFC

In here, Nothing was less than padj < 0.1, therefore no results. 

## 6. Cluster 3 (MG) ADp40KO vs AD

```{r warning=FALSE, message=FALSE}
# DE
DE_C3_ADp40KO_AD <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/DE_csv_files/20200117_Clustering_DE_by_MAST_each_Cluster_3AD_ADp40KO.csv", row.names = 1)

# GO
GO_C3_ADp40KO_AD <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/GO_csv_files/20200117_Clustering_DE_by_MAST_each_Cluster_3AD_ADp40KO.csv_GO_BP_only.csv", row.names = 1)

# GSEA
GSEA_C3_ADp40KO_AD_logFC_hall <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/GSEA_csv_files/20200117_Clustering_DE_by_MAST_each_Cluster_3AD_ADp40KO.csv_GSEA_logFC_hall.csv", row.names = 1)
GSEA_C3_ADp40KO_AD_logFC_reactome <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/GSEA_csv_files/20200117_Clustering_DE_by_MAST_each_Cluster_3AD_ADp40KO.csv_GSEA_logFC_reactome.csv", row.names = 1)
```

#### 3-2. Cluster 3 (MG) ADp40KO vs AD DE results

```{r warning=FALSE, message=FALSE, fig.width= 6, fig.height= 5}
DE_figure(DE_C3_ADp40KO_AD, text = "DE analysis cluster 3 ADp40KO vs AD")
```

#### 3-3. Cluster 3 (MG) ADp40KO vs AD GO results

```{r warning=FALSE, message=FALSE, fig.width= 7, fig.height= 4}
GO_figure(GO_C3_ADp40KO_AD)
```

#### 3-4. Cluster 3 (MG) ADp40KO vs AD GSEA results logFC

```{r warning=FALSE, message=FALSE, fig.width= 8, fig.height= 6}
GSEA_Hall_figure(GSEA_C3_ADp40KO_AD_logFC_hall)
```

#### 3-5. Cluster 3 (MG) ADp40KO vs AD GSEA results logFC

```{r warning=FALSE, message=FALSE, fig.width= 8, fig.height= 6}
GSEA_reactome_figure(GSEA_C3_ADp40KO_AD_logFC_reactome)
```

```{r warning=FALSE, message=FALSE, fig.width= 8, fig.height= 6}
sessionInfo()
```

