---
title: "CA2/3 analysis"
author: "Skim"
date: '2020 6 17 '
output: 
  html_document:
    df_print: paged
    toc: true
    toc_depth: 4
    toc_float: false
    code_folding: hide
---

```{r warning=FALSE, message=FALSE}
.libPaths(c("/home/skim/R/usr_lib", .libPaths()))
library(Seurat)
library(ggplot2)
library(gridExtra)
library(grid)
library(ggrepel)
library(DT)
library(biomaRt)
library(dplyr)
library(viridis)
library(tidyr)
source("/data/rajewsky/home/skim/Microglia_Heppner/Codes/DroNc_AD/src/TopGO_scRNA.R")
```

Below is each cluster in Neuronal cell types

Dentate Gyrus = 0, 10  
CA1 = 2, 13, 18, 40  
CA2/3 = 9, 21, 27  
Subiculum = 11, 14, 15, 17, 19, 20  
Inhibitory = 7, 16, 22, 30  
Neurons = 24, 29, 32

```{r warning=FALSE, message=FALSE}
load(file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200221_figure1_for_meeting_report_Seurat_object.Robj")
```

```{r warning=FALSE, message=FALSE}
cluster_GO <- function(genesOfInterest, geneUniverse){

geneList <- factor(as.integer(geneUniverse %in% genesOfInterest))
names(geneList) <- geneUniverse

onts = c( "MF", "BP", "CC" )
tab <- as.list(onts)
names(tab) <- onts

for(i in 1:3){

GOdata <- new("topGOdata",
                description = "GOanalysis",
                ontology = onts[i],
                allGenes = geneList,
                annot = annFUN.org,
                mapping = "org.Mm.eg.db",
                ID = "SYMBOL",
                nodeSize = 20)

# export GO ID from BP
if(i == 1){
  allGO_MF <- genesInTerm(GOdata)
}
if(i == 2){
  allGO_BP <- genesInTerm(GOdata)
}

if(i == 3){
  allGO_CC <- genesInTerm(GOdata)
}

res.result1 <- runTest(GOdata, statistic = "fisher", algorithm = "elim")
res.result2 <- runTest(GOdata, statistic = "fisher", algorithm = "classic")


tab[[i]] <- cbind(data.frame(category = onts[i]), GenTable(GOdata, Fisher.elim = res.result1,
                        Fisher.classic = res.result2,
                        orderBy = "Fisher.elim" , topNodes = 30))


}

topGOResults <- plyr::rbind.fill(tab)
topGOResults.df <- as.data.frame(topGOResults)
topGOResults.df$gene_ratio <- topGOResults.df$Significant / topGOResults.df$Annotated

# modification appropriate for plot
topGOResults.df$Fisher.elim <- as.numeric(topGOResults.df$Fisher.elim)
topGOResults.df$Fisher.classic <- as.numeric(topGOResults.df$Fisher.classic)
topGOResults.df$Term <- factor(topGOResults.df$Term, levels = rev(unique(topGOResults.df$Term)))

for(i in 1:30){
 ID <- topGOResults.df$GO.ID[i]
 #print(ID)
 #print(paste(genesOfInterest[genesOfInterest %in% allGO[ID][[1]]], collapse=', ' ))
 topGOResults.df[i, "genes"] <- paste(genesOfInterest[genesOfInterest %in% allGO_MF[ID][[1]]], collapse=', ' )
}
for(i in 31:60){
 ID <- topGOResults.df$GO.ID[i]
 #print(ID)
 #print(paste(genesOfInterest[genesOfInterest %in% allGO[ID][[1]]], collapse=', ' ))
 topGOResults.df[i, "genes"] <- paste(genesOfInterest[genesOfInterest %in% allGO_BP[ID][[1]]], collapse=', ' )
}
for(i in 61:90){
 ID <- topGOResults.df$GO.ID[i]
 #print(ID)
 #print(paste(genesOfInterest[genesOfInterest %in% allGO[ID][[1]]], collapse=', ' ))
 topGOResults.df[i, "genes"] <- paste(genesOfInterest[genesOfInterest %in% allGO_CC[ID][[1]]], collapse=', ' )
}

return(topGOResults.df)
}

# ggplot function
scRNA_TopGO_plot2 <- function(topGOResults.df){

topGOResults.df$Fisher.elim <- as.numeric(topGOResults.df$Fisher.elim)
topGOResults.df$Fisher.elim <- -log10(topGOResults.df$Fisher.elim)
topGOResults.df$Term <- factor(topGOResults.df$Term, levels = rev(unique(topGOResults.df$Term)))

ggplot(topGOResults.df, aes(x=gene_ratio,
               y=Term,
               colour=Fisher.elim,
               size=Significant)) +
        geom_point() +
        expand_limits(x=0) +
        labs(x="gene ratio", y="GO term", colour="-log10(p-value)", size="Significant") +
  theme_minimal() + theme(axis.text = element_text(size = 10))

}
```

### 1. cluster 32

Cluster 32 marker genes

4930467D21Rik & Cpne9is marker gene of (TEGLU23	Excitatory neurons, hippocampus CA3)

Cluster 32 will be included in CA2/3 cluster

```{r warning=FALSE, message=FALSE, fig.width=13, fig.height=12}
# markers <- list.files("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/DropViz_cluster_Markers/")
# genes <- read_csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/DropViz_cluster_Markers/cluster-markers_Neuron_CA2CA3_Pvrl3.csv")

Cluster_markers <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200210_FindMarkers_each_cluster_data9set_marker.csv")

f1 <- FeaturePlot(data9set_cleaned.SO, features = "Gm32647")
f2 <- FeaturePlot(data9set_cleaned.SO, features = "Cpne9")
f3 <- FeaturePlot(data9set_cleaned.SO, features = "4930467D21Rik")
f4 <- FeaturePlot(data9set_cleaned.SO, features = "Crispld1")

grid.arrange(f1, f2, f3, f4, ncol = 2)
```

CA2/3 marker genes

```{r warning=FALSE, message=FALSE, fig.width=13, fig.height=6}
f1 <- FeaturePlot(data9set_cleaned.SO, features = "Lsm11")
f2 <- FeaturePlot(data9set_cleaned.SO, features = "Homer3")

grid.arrange(f1, f2, ncol = 2)
```


### 2. CA2/3 overview

```{r warning=FALSE, message=FALSE, fig.width=8.5, fig.height=5}
data9set_CA23.SO <- subset(data9set_cleaned.SO, subset = seurat_clusters %in% c(9, 21, 27,32))
```

```{r warning=FALSE, message=FALSE, fig.width=8.5, fig.height=5}
plot.data <- data9set_cleaned.SO@meta.data %>%
    dplyr::select(sample, gemgroup, cluster = seurat_clusters) %>%
    mutate(cluster = cluster) %>%
    group_by(gemgroup, cluster, sample) %>%
  summarise(count = n()) %>%
  group_by(gemgroup) %>%
    mutate(gemgroup_total = sum(count)) %>%
    mutate(cell_prop = count / gemgroup_total)

ggplot(plot.data[plot.data$cluster %in% c(9, 21, 27, 32),], aes(x=cluster, y=cell_prop, fill=sample)) +
  geom_boxplot(position=position_dodge(1)) + ylim(c(0, 0.05))
```

```{r  warning=FALSE, message=FALSE, fig.width=8.5, fig.height=5}
Idents(data9set_CA23.SO) <- data9set_CA23.SO$seurat_clusters

# c9.markers <- FindMarkers(data9set_CA23.SO, ident.1 = "9", ident.2 = c("21", "27", "32"), min.pct = 0.05)
# write.csv(c9.markers, file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200617_CA2_3_neurons_analysis_CA2_3_Neurons_analysis_cluster9_markers.csv")
c9.markers <- read.csv(file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200617_CA2_3_neurons_analysis_CA2_3_Neurons_analysis_cluster9_markers.csv", row.names = 1)

# c21.markers <- FindMarkers(data9set_CA23.SO, ident.1 = "21", ident.2 = c("9", "27", "32"), min.pct = 0.05)
# write.csv(c21.markers, file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200617_CA2_3_neurons_analysis_CA2_3_Neurons_analysis_cluster21_markers.csv")
c21.markers <- read.csv(file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200617_CA2_3_neurons_analysis_CA2_3_Neurons_analysis_cluster21_markers.csv", row.names = 1)

# c27.markers <- FindMarkers(data9set_CA23.SO, ident.1 = "27", ident.2 = c("9", "21", "32"), min.pct = 0.05)
# write.csv(c27.markers, file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200617_CA2_3_neurons_analysis_CA2_3_Neurons_analysis_cluster27_markers.csv")
c27.markers <- read.csv(file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200617_CA2_3_neurons_analysis_CA2_3_Neurons_analysis_cluster27_markers.csv", row.names = 1)

# c32.markers <- FindMarkers(data9set_CA23.SO, ident.1 = "32", ident.2 = c("9", "21", "27"), min.pct = 0.05)
# write.csv(c32.markers, file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200617_CA2_3_neurons_analysis_CA2_3_Neurons_analysis_cluster32_markers.csv")
c32.markers <- read.csv(file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200617_CA2_3_neurons_analysis_CA2_3_Neurons_analysis_cluster32_markers.csv", row.names = 1)
```

```{r  warning=FALSE, message=FALSE, fig.width=15, fig.height = 5}
DimPlot(data9set_CA23.SO, split.by = "sample", group.by = "seurat_clusters") + xlim(c(0, 10)) + ylim(c(-16, -6))
```

#### 2-1. Cluster 9

```{r warning=FALSE, message=FALSE, fig.width= 8, fig.height= 7.5}
f1 <- FeaturePlot(data9set_CA23.SO, features = "Epha3") + xlim(c(0, 10)) + ylim(c(-16, -6))
f2 <- FeaturePlot(data9set_CA23.SO, features = "Cntnap5c") + xlim(c(0, 10)) + ylim(c(-16, -6))
f3 <- FeaturePlot(data9set_CA23.SO, features = "Fstl4") + xlim(c(0, 10)) + ylim(c(-16, -6))
f4 <- FeaturePlot(data9set_CA23.SO, features = "Etl4") + xlim(c(0, 10)) + ylim(c(-16, -6))

grid.arrange(f1, f2, f3, f4, ncol = 2)
```

```{r warning=FALSE, message=FALSE, fig.width= 15, fig.height= 5}

# geneUniverse <- rownames(data9set_CA23.SO@assays$RNA@counts[rowSums(data9set_CA23.SO@assays$RNA@counts) > 0, ]) %>% as.character
# 
# genesOfInterest <- as.character(rownames(c9.markers[which(c9.markers$p_val_adj < 0.01 & c9.markers$avg_logFC > 0.5 & c9.markers$pct.2 < 0.5), ]))
# 
# topGOResults.df <- cluster_GO(genesOfInterest, geneUniverse)
# #
# write.csv(topGOResults.df, file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200617_CA2_3_neurons_analysis_GO_cluster9.csv")

topGOResults.df <- read.csv(file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200617_CA2_3_neurons_analysis_GO_cluster9.csv", row.names = 1)

datatable(topGOResults.df, options = list(pageLength = 5, dom = 'tip'))
```

```{r warning=FALSE, message=FALSE, fig.width= 15, fig.height= 7}
topGOResults_MF.df <- topGOResults.df[which(topGOResults.df$category == "MF" & topGOResults.df$Annotated < 1000 & topGOResults.df$Annotated > 30), ][c(1:15), ]
g1 <- scRNA_TopGO_plot2(topGOResults_MF.df)
g1 <- g1 + ggtitle("Molecular function")

topGOResults_BP.df <- topGOResults.df[which(topGOResults.df$category == "BP" & topGOResults.df$Annotated < 1000 & topGOResults.df$Annotated > 30), ][c(1:15), ]
g2 <- scRNA_TopGO_plot2(topGOResults_BP.df)
g2 <- g2 + ggtitle("Biological process")

grid.arrange(g1, g2, ncol = 2)
```

#### 2-2. Cluster 21

```{r warning=FALSE, message=FALSE, fig.width= 8, fig.height= 7.5}
f1 <- FeaturePlot(data9set_CA23.SO, features = "Gabrg3") + xlim(c(0, 10)) + ylim(c(-16, -6))
f2 <- FeaturePlot(data9set_CA23.SO, features = "St18") + xlim(c(0, 10)) + ylim(c(-16, -6))
f3 <- FeaturePlot(data9set_CA23.SO, features = "Nrg1") + xlim(c(0, 10)) + ylim(c(-16, -6))
f4 <- FeaturePlot(data9set_CA23.SO, features = "Htr2c") + xlim(c(0, 10)) + ylim(c(-16, -6))

grid.arrange(f1, f2, f3, f4, ncol = 2)
```

```{r warning=FALSE, message=FALSE, fig.width= 15, fig.height= 5}

# genesOfInterest <- as.character(rownames(c21.markers[which(c21.markers$p_val_adj < 0.01 & c21.markers$avg_logFC > 0.5 & c21.markers$pct.2 < 0.5), ]))
# 
# topGOResults.df <- cluster_GO(genesOfInterest, geneUniverse)
# #
# write.csv(topGOResults.df, file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200617_CA2_3_neurons_analysis_GO_cluster21.csv")

topGOResults.df <- read.csv(file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200617_CA2_3_neurons_analysis_GO_cluster21.csv", row.names = 1)

datatable(topGOResults.df, options = list(pageLength = 5, dom = 'tip'))
```

```{r warning=FALSE, message=FALSE, fig.width= 15, fig.height= 7}
topGOResults_MF.df <- topGOResults.df[which(topGOResults.df$category == "MF" & topGOResults.df$Annotated < 1000 & topGOResults.df$Annotated > 30), ][c(1:15), ]
g1 <- scRNA_TopGO_plot2(topGOResults_MF.df)
g1 <- g1 + ggtitle("Molecular function")

topGOResults_BP.df <- topGOResults.df[which(topGOResults.df$category == "BP" & topGOResults.df$Annotated < 1000 & topGOResults.df$Annotated > 30), ][c(1:15), ]
g2 <- scRNA_TopGO_plot2(topGOResults_BP.df)
g2 <- g2 + ggtitle("Biological process")

grid.arrange(g1, g2, ncol = 2)
```

#### 2-3. Cluster 27

```{r warning=FALSE, message=FALSE, fig.width= 8, fig.height= 7.5}
f1 <- FeaturePlot(data9set_CA23.SO, features = "Asic2") + xlim(c(0, 10)) + ylim(c(-16, -6))
f2 <- FeaturePlot(data9set_CA23.SO, features = "Egfem1") + xlim(c(0, 10)) + ylim(c(-16, -6))
f3 <- FeaturePlot(data9set_CA23.SO, features = "Kcnk2") + xlim(c(0, 10)) + ylim(c(-16, -6))
f4 <- FeaturePlot(data9set_CA23.SO, features = "Adcy1") + xlim(c(0, 10)) + ylim(c(-16, -6))

grid.arrange(f1, f2, f3, f4, ncol = 2)
```

```{r warning=FALSE, message=FALSE, fig.width= 15, fig.height= 5}

# genesOfInterest <- as.character(rownames(c27.markers[which(c27.markers$p_val_adj < 0.01 & c27.markers$avg_logFC > 0.5 & c27.markers$pct.2 < 0.5), ]))
# 
# topGOResults.df <- cluster_GO(genesOfInterest, geneUniverse)
# #
# write.csv(topGOResults.df, file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200617_CA2_3_neurons_analysis_GO_cluster27.csv")

topGOResults.df <- read.csv(file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200617_CA2_3_neurons_analysis_GO_cluster27.csv", row.names = 1)

datatable(topGOResults.df, options = list(pageLength = 5, dom = 'tip'))
```

```{r warning=FALSE, message=FALSE, fig.width= 15, fig.height= 7}
topGOResults_MF.df <- topGOResults.df[which(topGOResults.df$category == "MF" & topGOResults.df$Annotated < 1000 & topGOResults.df$Annotated > 30), ][c(1:15), ]
g1 <- scRNA_TopGO_plot2(topGOResults_MF.df)
g1 <- g1 + ggtitle("Molecular function")

topGOResults_BP.df <- topGOResults.df[which(topGOResults.df$category == "BP" & topGOResults.df$Annotated < 1000 & topGOResults.df$Annotated > 30), ][c(1:15), ]
g2 <- scRNA_TopGO_plot2(topGOResults_BP.df)
g2 <- g2 + ggtitle("Biological process")

grid.arrange(g1, g2, ncol = 2)
```

#### 2-4. Cluster 32

CNTN4/BIG-2, an neuronal glycoprotein, which belongs to a subfamily of the immunoglobulin superfamily of cell adhesion molecules is known to guide axons during development. This function has been studied in the development of the olfactory and visual neural circuit However, the precise role of CNTN4 in establishing neural networks and synaptic contacts remains to be clarified. [ref](https://www.frontiersin.org/articles/10.3389/fnmol.2018.00212/full)

```{r warning=FALSE, message=FALSE, fig.width= 8, fig.height= 7.5}
f1 <- FeaturePlot(data9set_CA23.SO, features = "Cntn5") + xlim(c(0, 10)) + ylim(c(-16, -6))
f2 <- FeaturePlot(data9set_CA23.SO, features = "Cntn4") + xlim(c(0, 10)) + ylim(c(-16, -6))
f3 <- FeaturePlot(data9set_CA23.SO, features = "Grm8") + xlim(c(0, 10)) + ylim(c(-16, -6))
f4 <- FeaturePlot(data9set_CA23.SO, features = "Slc35f1") + xlim(c(0, 10)) + ylim(c(-16, -6))

grid.arrange(f1, f2, f3, f4, ncol = 2)
```

```{r warning=FALSE, message=FALSE, fig.width= 15, fig.height= 5}

# genesOfInterest <- as.character(rownames(c32.markers[which(c32.markers$p_val_adj < 0.01 & c32.markers$avg_logFC > 0.5 & c32.markers$pct.2 < 0.5), ]))
# 
# topGOResults.df <- cluster_GO(genesOfInterest, geneUniverse)
# #
# write.csv(topGOResults.df, file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200617_CA2_3_neurons_analysis_GO_cluster32.csv")

topGOResults.df <- read.csv(file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200617_CA2_3_neurons_analysis_GO_cluster32.csv", row.names = 1)

datatable(topGOResults.df, options = list(pageLength = 5, dom = 'tip'))
```

```{r warning=FALSE, message=FALSE, fig.width= 15, fig.height= 7}
topGOResults_MF.df <- topGOResults.df[which(topGOResults.df$category == "MF" & topGOResults.df$Annotated < 1000 & topGOResults.df$Annotated > 30), ][c(1:15), ]
g1 <- scRNA_TopGO_plot2(topGOResults_MF.df)
g1 <- g1 + ggtitle("Molecular function")

topGOResults_BP.df <- topGOResults.df[which(topGOResults.df$category == "BP" & topGOResults.df$Annotated < 1000 & topGOResults.df$Annotated > 30), ][c(1:15), ]
g2 <- scRNA_TopGO_plot2(topGOResults_BP.df)
g2 <- g2 + ggtitle("Biological process")

grid.arrange(g1, g2, ncol = 2)
```

### 3. DE in CA2/3 neurons

```{r warning=FALSE, message=FALSE, fig.width= 15, fig.height= 7}

# CA2_DE_AD_ADp40KO <- FindMarkers(data9set_CA23.SO, ident.1 = "AD", ident.2 = "ADp40KO", group.by = "sample", min.pct = 0.01, logfc.threshold = 0.01)
# 
# write.csv(CA2_DE_AD_ADp40KO, file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200617_CA2_3_neurons_analysis_CA2_3_Neurons_DE_AD_ADp40KO.csv")

CA2_DE_AD_ADp40KO <- read.csv(file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200617_CA2_3_neurons_analysis_CA2_3_Neurons_DE_AD_ADp40KO.csv")


CA2_DE_AD_ADp40KO2 <- CA2_DE_AD_ADp40KO[abs(CA2_DE_AD_ADp40KO$avg_logFC) > 0.1, ]
# remove Ttr
CA2_DE_AD_ADp40KO2 <- CA2_DE_AD_ADp40KO2[CA2_DE_AD_ADp40KO2$X != "Ttr", ]

# only logFC > 0.1
CA2_DE_AD_ADp40KO_padj <- CA2_DE_AD_ADp40KO[abs(CA2_DE_AD_ADp40KO$avg_logFC) > 0.1, ]
# remove Ttr
CA2_DE_AD_ADp40KO_padj <- CA2_DE_AD_ADp40KO_padj[CA2_DE_AD_ADp40KO_padj$X != "Ttr", ]

CA2_DE_AD_ADp40KO_padj <- CA2_DE_AD_ADp40KO_padj[which(CA2_DE_AD_ADp40KO_padj$p_val_adj < 0.05), ]

datatable(CA2_DE_AD_ADp40KO_padj, options = list(pageLength = 5, dom = 'tip'))
```

```{r warning=FALSE, message=FALSE, fig.width= 5.5, fig.height= 5}

ggplot(CA2_DE_AD_ADp40KO2) + geom_point(aes(x = avg_logFC, y = -log10(p_val_adj))) + 
  geom_point(data = CA2_DE_AD_ADp40KO_padj, aes(x = avg_logFC, y = -log10(p_val_adj)), color = "red") + theme_classic() +
  ggtitle("CA2/3 Neurons DE analysis AD vs ADp40KO") + 
   xlab("log2 fold change") + 
  ylab("-log10(adjusted p-value)") +
  theme(legend.position = "none",
        plot.title = element_text(size = 15, hjust = 0.5, family = "helvetica"),
        axis.title = element_text(size = 15), 
        axis.text = element_text(size = 12, color = "black")) +
  geom_text_repel(data = CA2_DE_AD_ADp40KO_padj[CA2_DE_AD_ADp40KO_padj$p_val_adj < 1e-10, ], aes(x = avg_logFC, y = -log10(p_val_adj), label = CA2_DE_AD_ADp40KO_padj[CA2_DE_AD_ADp40KO_padj$p_val_adj < 1e-10, ]$X)) 
```

GO of AD vs ADp40KO up-regulated genes

```{r warning=FALSE, message=FALSE, fig.width= 15, fig.height= 5}
# genesOfInterest <- CA2_DE_AD_ADp40KO_padj[CA2_DE_AD_ADp40KO_padj$avg_logFC > 0, ]$X %>% as.character
# 
# topGOResults.df <- cluster_GO(genesOfInterest, geneUniverse)
# 
# write.csv(topGOResults.df, file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200617_CA2_3_neurons_analysis_GO_AD_ADp40KO_UP.csv")

topGOResults.df <- read.csv(file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200617_CA2_3_neurons_analysis_GO_AD_ADp40KO_UP.csv", row.names = 1)

datatable(topGOResults.df, options = list(pageLength = 5, dom = 'tip'))
```

```{r warning=FALSE, message=FALSE, fig.width= 15, fig.height= 7}
topGOResults_MF.df <- topGOResults.df[which(topGOResults.df$category == "MF" & topGOResults.df$Annotated < 1000 & topGOResults.df$Annotated > 30), ][c(1:15), ]
g1 <- scRNA_TopGO_plot2(topGOResults_MF.df)
g1 <- g1 + ggtitle("Molecular function")

topGOResults_BP.df <- topGOResults.df[which(topGOResults.df$category == "BP" & topGOResults.df$Annotated < 1000 & topGOResults.df$Annotated > 30), ][c(1:15), ]
g2 <- scRNA_TopGO_plot2(topGOResults_BP.df)
g2 <- g2 + ggtitle("Biological process")

grid.arrange(g1, g2, ncol = 2)
```

GO of AD vs ADp40KO down-regulated genes

```{r warning=FALSE, message=FALSE, fig.width= 15, fig.height= 5}
# genesOfInterest <- CA2_DE_AD_ADp40KO_padj[CA2_DE_AD_ADp40KO_padj$avg_logFC < 0, ]$X %>% as.character
# 
# topGOResults.df <- cluster_GO(genesOfInterest, geneUniverse)
# 
# write.csv(topGOResults.df, file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200617_CA2_3_neurons_analysis_GO_AD_ADp40KO_DOWN.csv")

topGOResults.df <- read.csv(file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200617_CA2_3_neurons_analysis_GO_AD_ADp40KO_DOWN.csv", row.names = 1)

datatable(topGOResults.df, options = list(pageLength = 5, dom = 'tip'))
```

```{r warning=FALSE, message=FALSE, fig.width= 15, fig.height= 7}
topGOResults_MF.df <- topGOResults.df[which(topGOResults.df$category == "MF" & topGOResults.df$Annotated < 1000 & topGOResults.df$Annotated > 30), ][c(1:12), ]
g1 <- scRNA_TopGO_plot2(topGOResults_MF.df)
g1 <- g1 + ggtitle("Molecular function")

topGOResults_BP.df <- topGOResults.df[which(topGOResults.df$category == "BP" & topGOResults.df$Annotated < 1000 & topGOResults.df$Annotated > 30), ][c(1:15), ]
g2 <- scRNA_TopGO_plot2(topGOResults_BP.df)
g2 <- g2 + ggtitle("Biological process")

grid.arrange(g1, g2, ncol = 2)
```

### 4. DE by clustter

#### 4-1. DE cluster 9

```{r warning=FALSE, message=FALSE}
c9_DE_AD_ADp40KO <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/DE_csv_files/20200117_Clustering_DE_by_MAST_each_Cluster_9ADp40KO_AD_corrected.csv")

# reverse logFC2
c9_DE_AD_ADp40KO$avg_logFC <- c9_DE_AD_ADp40KO$avg_logFC * -1

c9_DE_AD_ADp40KO2 <- c9_DE_AD_ADp40KO[abs(c9_DE_AD_ADp40KO$avg_logFC) > 0.1, ]
# remove Ttr
c9_DE_AD_ADp40KO2 <- c9_DE_AD_ADp40KO2[c9_DE_AD_ADp40KO2$X != "Ttr", ]

# only logFC > 0.1
c9_DE_AD_ADp40KO_padj <- c9_DE_AD_ADp40KO[abs(c9_DE_AD_ADp40KO$avg_logFC) > 0.1, ]
# remove Ttr
c9_DE_AD_ADp40KO_padj <- c9_DE_AD_ADp40KO_padj[c9_DE_AD_ADp40KO_padj$X != "Ttr", ]

c9_DE_AD_ADp40KO_padj <- c9_DE_AD_ADp40KO_padj[which(c9_DE_AD_ADp40KO_padj$p_val_adj < 0.05), ]

datatable(c9_DE_AD_ADp40KO_padj, options = list(pageLength = 5, dom = 'tip'))

```

#### 4-2. DE cluster 21

```{r warning=FALSE, message=FALSE}
c21_DE_AD_ADp40KO <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/DE_csv_files/20200117_Clustering_DE_by_MAST_each_Cluster_21ADp40KO_AD_corrected.csv")

# reverse logFC2
c21_DE_AD_ADp40KO$avg_logFC <- c21_DE_AD_ADp40KO$avg_logFC * -1

c21_DE_AD_ADp40KO2 <- c21_DE_AD_ADp40KO[abs(c21_DE_AD_ADp40KO$avg_logFC) > 0.1, ]
# remove Ttr
c21_DE_AD_ADp40KO2 <- c21_DE_AD_ADp40KO2[c21_DE_AD_ADp40KO2$X != "Ttr", ]

# only logFC > 0.1
c21_DE_AD_ADp40KO_padj <- c21_DE_AD_ADp40KO[abs(c21_DE_AD_ADp40KO$avg_logFC) > 0.1, ]
# remove Ttr
c21_DE_AD_ADp40KO_padj <- c21_DE_AD_ADp40KO_padj[c21_DE_AD_ADp40KO_padj$X != "Ttr", ]

c21_DE_AD_ADp40KO_padj <- c21_DE_AD_ADp40KO_padj[which(c21_DE_AD_ADp40KO_padj$p_val_adj < 0.05), ]

datatable(c21_DE_AD_ADp40KO_padj, options = list(pageLength = 5, dom = 'tip'))

```


#### 4-3. DE cluster 27

```{r warning=FALSE, message=FALSE}
c27_DE_AD_ADp40KO <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/DE_csv_files/20200117_Clustering_DE_by_MAST_each_Cluster_27ADp40KO_AD_corrected.csv")

# reverse logFC2
c27_DE_AD_ADp40KO$avg_logFC <- c27_DE_AD_ADp40KO$avg_logFC * -1

c27_DE_AD_ADp40KO2 <- c27_DE_AD_ADp40KO[abs(c27_DE_AD_ADp40KO$avg_logFC) > 0.1, ]
# remove Ttr
c27_DE_AD_ADp40KO2 <- c27_DE_AD_ADp40KO2[c27_DE_AD_ADp40KO2$X != "Ttr", ]

# only logFC > 0.1
c27_DE_AD_ADp40KO_padj <- c27_DE_AD_ADp40KO[abs(c27_DE_AD_ADp40KO$avg_logFC) > 0.1, ]
# remove Ttr
c27_DE_AD_ADp40KO_padj <- c27_DE_AD_ADp40KO_padj[c27_DE_AD_ADp40KO_padj$X != "Ttr", ]

c27_DE_AD_ADp40KO_padj <- c27_DE_AD_ADp40KO_padj[which(c27_DE_AD_ADp40KO_padj$p_val_adj < 0.05), ]

datatable(c27_DE_AD_ADp40KO_padj, options = list(pageLength = 5, dom = 'tip'))

```

#### 4-4. DE cluster 32

```{r warning=FALSE, message=FALSE}
c32_DE_AD_ADp40KO <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/DE_csv_files/20200117_Clustering_DE_by_MAST_each_Cluster_32ADp40KO_AD_corrected.csv")

# reverse logFC2
c32_DE_AD_ADp40KO$avg_logFC <- c32_DE_AD_ADp40KO$avg_logFC * -1

c32_DE_AD_ADp40KO2 <- c32_DE_AD_ADp40KO[abs(c32_DE_AD_ADp40KO$avg_logFC) > 0.1, ]
# remove Ttr
c32_DE_AD_ADp40KO2 <- c32_DE_AD_ADp40KO2[c32_DE_AD_ADp40KO2$X != "Ttr", ]

# only logFC > 0.1
c32_DE_AD_ADp40KO_padj <- c32_DE_AD_ADp40KO[abs(c32_DE_AD_ADp40KO$avg_logFC) > 0.1, ]
# remove Ttr
c32_DE_AD_ADp40KO_padj <- c32_DE_AD_ADp40KO_padj[c32_DE_AD_ADp40KO_padj$X != "Ttr", ]

c32_DE_AD_ADp40KO_padj <- c32_DE_AD_ADp40KO_padj[which(c32_DE_AD_ADp40KO_padj$p_val_adj < 0.05), ]

datatable(c32_DE_AD_ADp40KO_padj, options = list(pageLength = 5, dom = 'tip'))

```

```{r warning=FALSE, message=FALSE}
sessionInfo()
```

