---
title: "subclustering_MG_GSEA analysis"
author: "Skim"
date: "2/26/2020"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_depth: 4
    toc_float: false
    code_folding: hide
---

This report is for Microglia GSEA analysis and shows its corresponding genes average logFC

```{r warning=FALSE, message=FALSE}
library(Seurat)
library(ggplot2)
library(dplyr)
library(DT)
library(gridExtra)
library(nichenetr)
library(tidyverse)
library(irr)
library(viridis)
```

```{r warning=FALSE, message=FALSE}
load("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200221_figure1_for_meeting_report_Seurat_object.Robj")
```

## 1. MG cluster analysis

This analysis is done by subset 2 Microglia clusters (from res 0.8 and cluster 34 was removed and 37 is macrophages) and I investigate more in here.

```{r warning=FALSE, message=FALSE, fig.width = 6, fig.height=5}
data9set_cleaned_MG.SO <- subset(data9set_cleaned.SO, subset = seurat_clusters %in% c(3, 8))

d1 <- DimPlot(object = data9set_cleaned_MG.SO, reduction = "umap", group.by = "seurat_clusters") + scale_x_continuous(limits = c(-10,10)) + scale_y_continuous(limits = c(10, 30)) 
d1
```

## 2. AD vs Ctrl (GSEA Hallmark)

GSEA padj < 0.05 & DE padj < 0.01. There is no statistical significant of GSEA for negative NES. 

```{r warning=FALSE, message=FALSE, fig.width= 15, fig.height=5.5}
DE_AD_Ctrl <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/DE_csv_files/Cell_type/20200221_Cell_type_DE_by_MAST_MicrogliaAD_Ctrl.csv", row.names = 1)
GSEA_AD_Ctrl_Hall <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/GSEA_csv_files/Cell_type/20200221_Cell_type_DE_by_MAST_MicrogliaAD_Ctrl.csv_GSEA_logFC_hall.csv", row.names = 1)

GSEA_AD_Ctrl_Hall <- GSEA_AD_Ctrl_Hall[GSEA_AD_Ctrl_Hall$padj < .05, ]
GSEA_AD_Ctrl_Hall <- GSEA_AD_Ctrl_Hall[rev(order(GSEA_AD_Ctrl_Hall$NES)), ]
GSEA_AD_Ctrl_Hall <- GSEA_AD_Ctrl_Hall[c(1:6), ]

g1 <- ggplot(GSEA_AD_Ctrl_Hall, aes(reorder(pathway, NES), NES)) +
  geom_col(aes(fill=padj<0.05)) +
  coord_flip() +
  labs(title = "GSEA", y="Normalized Enrichment Score") + 
  theme_minimal() + theme(axis.title.y = element_blank(), axis.text.y = element_text(size = 10), legend.position = "none")

firstup <- function(x) {
  substr(x, 1, 1) <- toupper(substr(x, 1, 1))
  x
}


DE_AD_Ctrl_gene <- data.frame(pathway = as.character(), avg_logFC = as.numeric(), gene = as.character(), stringsAsFactors = FALSE)

for(i in c(1:6)){
  gene1 <- tolower(unlist(strsplit(as.character(GSEA_AD_Ctrl_Hall[i,]$leadingEdge), ",")))
  gene1 <- as.vector(firstup(str_trim(gene1, side = c("both", "left", "right"))))

  DE_AD_Ctrl_gene1 <- DE_AD_Ctrl[which(rownames(DE_AD_Ctrl) %in% gene1 & DE_AD_Ctrl$p_val_adj < .01), ]
  DE_AD_Ctrl_gene1 <- data.frame(pathway = GSEA_AD_Ctrl_Hall[i, ]$pathway, avg_logFC = DE_AD_Ctrl_gene1[, 2], gene = rownames(DE_AD_Ctrl_gene1))
  DE_AD_Ctrl_gene <- rbind(DE_AD_Ctrl_gene, DE_AD_Ctrl_gene1)

}


DE_AD_Ctrl_gene$pathway <- factor(DE_AD_Ctrl_gene$pathway, levels = GSEA_AD_Ctrl_Hall$pathway[order(GSEA_AD_Ctrl_Hall$NES)])

# 14, 15, 31, 33
DE_AD_Ctrl_gene_text <- DE_AD_Ctrl_gene[-c(14,15,30, 31,33, 53, 54, 72), ]

g2 <- ggplot(DE_AD_Ctrl_gene, aes(x=pathway, y=avg_logFC)) + 
  geom_violin(trim=FALSE, fill="#CCFFFF")+
  labs(title="statistically significant pathway genes", y="avg logFC")+
  theme_classic() + geom_text(data = DE_AD_Ctrl_gene_text[DE_AD_Ctrl_gene_text$avg_logFC > 0.2, ], aes(label= gene), size = 3.5, angle = 45) + coord_flip() + theme(axis.title.y = element_blank(), axis.text.y = element_blank())

grid.arrange(g1, g2, ncol = 2, widths = c(0.8, 1.2))

```

### 2-2. AD vs Ctrl (GSEA Reactome)

None of GSEA pathways were statistically significant. 

```{r warning=FALSE, message=FALSE, fig.width= 15, fig.height=5}
GSEA_AD_Ctrl_Reac <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/GSEA_csv_files/Cell_type/20200221_Cell_type_DE_by_MAST_MicrogliaAD_Ctrl.csv_GSEA_logFC_reactome.csv", row.names = 1)
```

## 3. AD vs ADp40KO (GSEA Hallmark)

None of GSEA pathways were statistically significant. 

```{r warning=FALSE, message=FALSE, fig.width= 15, fig.height=5}
DE_AD_ADp40KO <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/DE_csv_files/Cell_type/20200221_Cell_type_DE_by_MAST_MicrogliaAD_ADp40KO.csv", row.names = 1)
GSEA_AD_ADp40KO_Hall <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/GSEA_csv_files/Cell_type/20200221_Cell_type_DE_by_MAST_MicrogliaAD_ADp40KO.csv_GSEA_logFC_hall.csv", row.names = 1)

GSEA_AD_ADp40KO_Hall <- GSEA_AD_ADp40KO_Hall[GSEA_AD_ADp40KO_Hall$padj < .05, ]
GSEA_AD_ADp40KO_Hall <- GSEA_AD_ADp40KO_Hall[rev(order(GSEA_AD_ADp40KO_Hall$NES)), ]
GSEA_AD_ADp40KO_Hall <- GSEA_AD_ADp40KO_Hall[c(1:6), ]
```

### 3-2. AD vs ADp40KO (GSEA Reactome)

GSEA padj < 0.05 & DE padj < 0.01. For the negative GSEA, there were no statistically significant genes

```{r warning=FALSE, message=FALSE, fig.width= 15, fig.height=5.5}
GSEA_AD_ADp40KO_Reac <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/GSEA_csv_files/Cell_type/20200221_Cell_type_DE_by_MAST_MicrogliaAD_ADp40KO.csv_GSEA_logFC_reactome.csv", row.names = 1)
GSEA_AD_ADp40KO_Reac <- GSEA_AD_ADp40KO_Reac[GSEA_AD_ADp40KO_Reac$padj < .05, ]
GSEA_AD_ADp40KO_Reac <- GSEA_AD_ADp40KO_Reac[rev(order(GSEA_AD_ADp40KO_Reac$NES)), ]

g1 <- ggplot(GSEA_AD_ADp40KO_Reac, aes(reorder(pathway, NES), NES)) +
  geom_col(aes(fill=padj<0.05)) +
  coord_flip() +
  labs(title = "GSEA", y="Normalized Enrichment Score") + 
  theme_minimal() + theme(axis.title.y = element_blank(), axis.text.y = element_text(size = 10), legend.position = "none")

DE_AD_ADp40KO_gene <- data.frame(pathway = as.character(), avg_logFC = as.numeric(), gene = as.character(), stringsAsFactors = FALSE)

for(i in c(1:6)){
  #print(i)
  gene1 <- tolower(unlist(strsplit(as.character(GSEA_AD_ADp40KO_Reac[i,]$leadingEdge), ",")))
  gene1 <- as.vector(firstup(str_trim(gene1, side = c("both", "left", "right"))))
  #print(gene1)
  DE_AD_ADp40KO_gene1 <- DE_AD_ADp40KO[which(rownames(DE_AD_ADp40KO) %in% gene1 & DE_AD_ADp40KO$p_val_adj < 0.01), ]
  if(nrow(DE_AD_ADp40KO_gene1) > 0){
    DE_AD_ADp40KO_gene1 <- data.frame(pathway = GSEA_AD_ADp40KO_Reac[i, ]$pathway, avg_logFC = DE_AD_ADp40KO_gene1[, 2], gene = rownames(DE_AD_ADp40KO_gene1))
  }else{
    DE_AD_ADp40KO_gene1 <- data.frame(pathway = GSEA_AD_ADp40KO_Reac[i, ]$pathway, avg_logFC = NA, gene = NA)
  }
  DE_AD_ADp40KO_gene <- rbind(DE_AD_ADp40KO_gene, DE_AD_ADp40KO_gene1)

}

DE_AD_ADp40KO_gene$pathway <- factor(DE_AD_ADp40KO_gene$pathway, levels = GSEA_AD_ADp40KO_Reac$pathway[order(GSEA_AD_ADp40KO_Reac$NES)])


g2 <- ggplot(DE_AD_ADp40KO_gene, aes(x=pathway, y=avg_logFC)) + 
  geom_violin(trim=FALSE, fill="#CCFFFF")+
  labs(title="statistically significant pathway genes", y="avg logFC")+
  theme_classic() + geom_text(data = DE_AD_ADp40KO_gene, aes(label= gene), size = 3.5, angle = 45) + coord_flip() + theme(axis.title.y = element_blank(), axis.text.y = element_blank())
grid.arrange(g1, g2, ncol = 2, widths = c(0.8, 1.2))

```

## 4. ADp40KO vs Ctrl (GSEA Hallmark)

GSEA padj < 0.05 & DE padj < 0.01. There is no statistical significant of GSEA for negative NES. 

```{r warning=FALSE, message=FALSE, fig.width= 15, fig.height=5.5}
DE_ADp40KO_Ctrl <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/DE_csv_files/Cell_type/20200221_Cell_type_DE_by_MAST_MicrogliaADp40KO_Ctrl.csv", row.names = 1)
GSEA_ADp40KO_Ctrl_Hall <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/GSEA_csv_files/Cell_type/20200221_Cell_type_DE_by_MAST_MicrogliaADp40KO_Ctrl.csv_GSEA_logFC_hall.csv", row.names = 1)

GSEA_ADp40KO_Ctrl_Hall <- GSEA_ADp40KO_Ctrl_Hall[GSEA_ADp40KO_Ctrl_Hall$padj < .05, ]
GSEA_ADp40KO_Ctrl_Hall <- GSEA_ADp40KO_Ctrl_Hall[rev(order(GSEA_ADp40KO_Ctrl_Hall$NES)), ]
GSEA_ADp40KO_Ctrl_Hall <- GSEA_ADp40KO_Ctrl_Hall[c(1:6), ]

g1 <- ggplot(GSEA_ADp40KO_Ctrl_Hall, aes(reorder(pathway, NES), NES)) +
  geom_col(aes(fill=padj<0.05)) +
  coord_flip() +
  labs(title = "GSEA", y="Normalized Enrichment Score") + 
  theme_minimal() + theme(axis.title.y = element_blank(), axis.text.y = element_text(size = 10), legend.position = "none")

DE_ADp40KO_Ctrl_gene <- data.frame(pathway = as.character(), avg_logFC = as.numeric(), gene = as.character(), stringsAsFactors = FALSE)

for(i in c(1:6)){
  gene1 <- tolower(unlist(strsplit(as.character(GSEA_ADp40KO_Ctrl_Hall[i,]$leadingEdge), ",")))
  gene1 <- as.vector(firstup(str_trim(gene1, side = c("both", "left", "right"))))
  DE_ADp40KO_Ctrl_gene1 <- DE_ADp40KO_Ctrl[which(rownames(DE_ADp40KO_Ctrl) %in% gene1 & DE_ADp40KO_Ctrl$p_val_adj < 0.01), ]
  DE_ADp40KO_Ctrl_gene1 <- data.frame(pathway = GSEA_ADp40KO_Ctrl_Hall[i, ]$pathway, avg_logFC = DE_ADp40KO_Ctrl_gene1[, 2], gene = rownames(DE_ADp40KO_Ctrl_gene1))
  DE_ADp40KO_Ctrl_gene <- rbind(DE_ADp40KO_Ctrl_gene, DE_ADp40KO_Ctrl_gene1)

}
DE_ADp40KO_Ctrl_gene$pathway <- factor(DE_ADp40KO_Ctrl_gene$pathway, levels = GSEA_ADp40KO_Ctrl_Hall$pathway[order(GSEA_ADp40KO_Ctrl_Hall$NES)])

# 11, 12, 29, 30
DE_ADp40KO_Ctrl_gene <- DE_ADp40KO_Ctrl_gene[-c(2, 6, 11,12,22, 29,30), ]

g2 <- ggplot(DE_ADp40KO_Ctrl_gene, aes(x=pathway, y=avg_logFC)) + 
  geom_violin(trim=FALSE, fill="#CCFFFF")+
  labs(title="statistically significant pathway genes", y="avg logFC")+
  theme_classic() + geom_text(data = DE_ADp40KO_Ctrl_gene[DE_ADp40KO_Ctrl_gene$avg_logFC > 0.2, ], aes(label= gene), size = 3.5, angle = 45) + coord_flip() + theme(axis.title.y = element_blank(), axis.text.y = element_blank())

grid.arrange(g1, g2, ncol = 2, widths = c(0.8, 1.2))

```

### 4-2. AD vs ADp40KO (GSEA Reactome)

None of GSEA pathways were statistically significant. 

```{r warning=FALSE, message=FALSE, fig.width= 15, fig.height=5}
GSEA_AD_ADp40KO_Reac <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/GSEA_csv_files/Cell_type/20200221_Cell_type_DE_by_MAST_MicrogliaADp40KO_Ctrl.csv_GSEA_logFC_reactome.csv", row.names = 1)
```

```{r warning=FALSE, message=FALSE, fig.width= 15, fig.height=5}
sessionInfo()
```

