---
title: "DAA Habib and Trem2_AD_published data analysis"
author: "Skim"
date: '2020 8 24 '
output: 
  html_document:
    df_print: paged
    toc: true
    toc_depth: 4
    toc_float: false
    code_folding: hide
---

This report is for analysis published data to look at Il12b in other cell types. I analyzed Zhou et al., Nature medicine, 2020
[Files download](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE140399) In here, I check 15 month old hippocampus / Cortex and WT and FAD to see Il12b expression in other cell type. I also tested Naomi Habib et al., DAA Nature Neuroscience paper. 

```{r warning=FALSE, message=FALSE}
.libPaths(c("/data/rajewsky/home/skim/R/usr_lib_Seurat/", "/data/rajewsky/home/skim/R/usr_lib_v4/"))
.libPaths()
# unloadNamespace("mgcv")
# unloadNamespace("Matrix")

library(Seurat)
library(ggplot2)
library(dplyr)
library(DT)
library(gridExtra)
library(openxlsx)
```

### Data load

```{r warning=FALSE, message=FALSE}
excel.df <- read.xlsx("/data/rajewsky/home/skim/Microglia_Heppner/published_Trem2_AD_Zhou/41591_2019_695_MOESM2_ESM.xlsx", sheet = "Cluster 7")

excel.df1 <- excel.df[c(3:16954),c(1:7)]
colnames(excel.df1) <- excel.df1[1, ]
excel.df1 <- excel.df1[-1, ]
```


```{r warning=FALSE, message=FALSE}
# Hippo
WT <- Read10X("/data/rajewsky/home/skim/Microglia_Heppner/published_Trem2_AD_Zhou/GSM4160647_WT_Hip/")

FAD <- Read10X("/data/rajewsky/home/skim/Microglia_Heppner/published_Trem2_AD_Zhou/GSM4160649_WT_5XFAD_Hip/")

WT.SO <- CreateSeuratObject(counts =  WT,  min.cells = 3, min.features = 200, project = "Trem2")

FAD.SO <- CreateSeuratObject(counts =  FAD,  min.cells = 3, min.features = 200, project = "Trem2")

# Cortex
Cor_WT <- Read10X("/data/rajewsky/home/skim/Microglia_Heppner/published_Trem2_AD_Zhou/GSM4160643_WT_Cor/")

Cor_FAD <- Read10X("/data/rajewsky/home/skim/Microglia_Heppner/published_Trem2_AD_Zhou/GSM4160645_WT_5XFAD_Cor/")

Cor_WT.SO <- CreateSeuratObject(counts =  Cor_WT,  min.cells = 3, min.features = 200, project = "Trem2")

Cor_FAD.SO <- CreateSeuratObject(counts =  Cor_FAD,  min.cells = 3, min.features = 200, project = "Trem2")

```

## Hippo

```{r warning=FALSE, message=FALSE}
# calculate MT genes
WT.SO[["percent.mt"]] <- PercentageFeatureSet(object = WT.SO, pattern = "^mt-")

FAD.SO[["percent.mt"]] <- PercentageFeatureSet(object = FAD.SO, pattern = "^mt-")

```

### Hippo QC

```{r warning=FALSE, message=FALSE, fig.width=15, fig.height=6}
plot1 <- FeatureScatter(WT.SO, feature1 = "nCount_RNA", feature2 = "nFeature_RNA") + ggtitle("Hip WT")
plot2 <- FeatureScatter(FAD.SO, feature1 = "nCount_RNA", feature2 = "nFeature_RNA") + ggtitle("Hip FAD")
plot1 + plot2

```

### Hippo mitochondrial

```{r warning=FALSE, message=FALSE, fig.width=15, fig.height=6}
v1 <- VlnPlot(WT.SO, features = c("percent.mt"), pt.size = 0.2) + ggtitle("Hip WT percent.mt")
v2 <- VlnPlot(FAD.SO, features = c("percent.mt"), pt.size = 0.2) + ggtitle("Hip FAD percent.mt")
grid.arrange(v1, v2, ncol = 2)
```


```{r warning=FALSE, message=FALSE}
#####
# WT
#####
WT.SO <- subset(x = WT.SO, subset = nCount_RNA < 5000 & nCount_RNA > 200 & percent.mt < 5)  # 3787 -> 2722

WT.SO <- NormalizeData(object = WT.SO, normalization.method = "LogNormalize", scale.factor = 1e4, verbose = FALSE)
WT.SO <- FindVariableFeatures(object = WT.SO, nfeatures = 2000, selection.method = "vst", verbose = FALSE)
WT.SO <- ScaleData(object = WT.SO, vars.to.regress = c("percent.mt", "nCount_RNA"))

# PCA
WT.SO <- RunPCA(WT.SO, features = VariableFeatures(object = WT.SO), verbose = FALSE, ndims.print = "None")
#ElbowPlot(WT.SO, ndims = 50)

WT.SO <- FindNeighbors(object = WT.SO, reduction = "pca", dims = 1:20)
WT.SO <- FindClusters(object = WT.SO, resolution = 1, verbose = FALSE)

WT.SO <- RunUMAP(object = WT.SO, dims = 1:20, n.neighbors = 20, min.dist = 0.35, n.epochs = 500, spread = 2)
d1 <- DimPlot(object = WT.SO, reduction = "umap")

#####
# FAD
#####
FAD.SO <- subset(x = FAD.SO, subset = nCount_RNA < 6000 & nCount_RNA > 200 & percent.mt < 5)  # 3787 -> 2722

FAD.SO <- NormalizeData(object = FAD.SO, normalization.method = "LogNormalize", scale.factor = 1e4, verbose = FALSE)
FAD.SO <- FindVariableFeatures(object = FAD.SO, nfeatures = 2000, selection.method = "vst", verbose = FALSE)
FAD.SO <- ScaleData(object = FAD.SO, vars.to.regress = c("percent.mt", "nCount_RNA"))

# PCA
FAD.SO <- RunPCA(FAD.SO, features = VariableFeatures(object = FAD.SO), verbose = FALSE, ndims.print = "None")
#ElbowPlot(FAD.SO, ndims = 50)

FAD.SO <- FindNeighbors(object = FAD.SO, reduction = "pca", dims = 1:20)
FAD.SO <- FindClusters(object = FAD.SO, resolution = 1, verbose = FALSE)

FAD.SO <- RunUMAP(object = FAD.SO, dims = 1:20, n.neighbors = 20, min.dist = 0.35, n.epochs = 500, spread = 2)
d2 <- DimPlot(object = FAD.SO, reduction = "umap")

```

### Hippo UMAP

left is WT and right is FAD

```{r warning=FALSE, message=FALSE, fig.width=15, fig.height=6}
grid.arrange(d1, d2, ncol = 2)
```

### Hippo Hexb (MG)

```{r warning=FALSE, message=FALSE, fig.width=15, fig.height=6}
f1 <- FeaturePlot(WT.SO, features = c("Hexb"), by.col = 2)
f2 <- FeaturePlot(FAD.SO, features = c("Hexb"), by.col = 2)
grid.arrange(f1, f2, ncol = 2)
```

### Hippo Il1- genes

WT expression of Il1- genes

```{r warning=FALSE, message=FALSE, fig.width=15, fig.height=6}
WT.SO@assays$RNA@counts@Dimnames[[1]][grepl("^Il1", WT.SO@assays$RNA@counts@Dimnames[[1]])]

```

```{r warning=FALSE, message=FALSE, fig.width=15, fig.height=6}

WT@Dimnames[[1]][grepl("^Il12", WT@Dimnames[[1]])]



```

FAD expression of Il1- genes

```{r warning=FALSE, message=FALSE, fig.width=15, fig.height=6}
FAD.SO@assays$RNA@counts@Dimnames[[1]][grepl("^Il1", FAD.SO@assays$RNA@counts@Dimnames[[1]])]

```

```{r warning=FALSE, message=FALSE, fig.width=15, fig.height=6}

FAD@Dimnames[[1]][grepl("^Il12", FAD@Dimnames[[1]])]



```

### Hippo Il12a 

```{r warning=FALSE, message=FALSE, fig.width=15, fig.height=6}
f1 <- FeaturePlot(WT.SO, features = c("Il12a"), by.col = 2)
f2 <- FeaturePlot(FAD.SO, features = c("Il12a"), by.col = 2)
grid.arrange(f1, f2, ncol = 2)
```

## Cortex

```{r warning=FALSE, message=FALSE}
# calculate MT genes
Cor_WT.SO[["percent.mt"]] <- PercentageFeatureSet(object = Cor_WT.SO, pattern = "^mt-")

Cor_FAD.SO[["percent.mt"]] <- PercentageFeatureSet(object = Cor_FAD.SO, pattern = "^mt-")

```


### Cortex QC

```{r warning=FALSE, message=FALSE, fig.width=15, fig.height=6}
plot1 <- FeatureScatter(Cor_WT.SO, feature1 = "nCount_RNA", feature2 = "nFeature_RNA") + ggtitle("Cor WT")
plot2 <- FeatureScatter(Cor_FAD.SO, feature1 = "nCount_RNA", feature2 = "nFeature_RNA") + ggtitle("Cor FAD")
plot1 + plot2

```

### Cortex mitochondrial

```{r warning=FALSE, message=FALSE, fig.width=15, fig.height=6}
v1 <- VlnPlot(Cor_WT.SO, features = c("percent.mt"), pt.size = 0.2) + ggtitle("Cor WT percent.mt")
v2 <- VlnPlot(Cor_FAD.SO, features = c("percent.mt"), pt.size = 0.2) + ggtitle("Cor FAD percent.mt")
grid.arrange(v1, v2, ncol = 2)
```

```{r warning=FALSE, message=FALSE}
#####
# WT
#####
Cor_WT.SO <- subset(x = Cor_WT.SO, subset = nCount_RNA < 6000 & nCount_RNA > 200 & percent.mt < 5)  

Cor_WT.SO <- NormalizeData(object = Cor_WT.SO, normalization.method = "LogNormalize", scale.factor = 1e4, verbose = FALSE)
Cor_WT.SO <- FindVariableFeatures(object = Cor_WT.SO, nfeatures = 2000, selection.method = "vst", verbose = FALSE)
Cor_WT.SO <- ScaleData(object = Cor_WT.SO, vars.to.regress = c("percent.mt", "nCount_RNA"))

# PCA
Cor_WT.SO <- RunPCA(Cor_WT.SO, features = VariableFeatures(object = Cor_WT.SO), verbose = FALSE, ndims.print = "None")
#ElbowPlot(Cor_WT.SO, ndims = 50)

Cor_WT.SO <- FindNeighbors(object = Cor_WT.SO, reduction = "pca", dims = 1:20)
Cor_WT.SO <- FindClusters(object = Cor_WT.SO, resolution = 1, verbose = FALSE)

Cor_WT.SO <- RunUMAP(object = Cor_WT.SO, dims = 1:20, n.neighbors = 20, min.dist = 0.35, n.epochs = 500, spread = 2)
d3 <- DimPlot(object = Cor_WT.SO, reduction = "umap")

#####
# FAD
#####
Cor_FAD.SO <- subset(x = Cor_FAD.SO, subset = nCount_RNA < 6000 & nCount_RNA > 200 & percent.mt < 5)  # 3787 -> 2722

Cor_FAD.SO <- NormalizeData(object = Cor_FAD.SO, normalization.method = "LogNormalize", scale.factor = 1e4, verbose = FALSE)
Cor_FAD.SO <- FindVariableFeatures(object = Cor_FAD.SO, nfeatures = 2000, selection.method = "vst", verbose = FALSE)
Cor_FAD.SO <- ScaleData(object = Cor_FAD.SO, vars.to.regress = c("percent.mt", "nCount_RNA"))

# PCA
Cor_FAD.SO <- RunPCA(Cor_FAD.SO, features = VariableFeatures(object = Cor_FAD.SO), verbose = FALSE, ndims.print = "None")
#ElbowPlot(Cor_FAD.SO, ndims = 50)

Cor_FAD.SO <- FindNeighbors(object = Cor_FAD.SO, reduction = "pca", dims = 1:20)
Cor_FAD.SO <- FindClusters(object = Cor_FAD.SO, resolution = 1, verbose = FALSE)

Cor_FAD.SO <- RunUMAP(object = Cor_FAD.SO, dims = 1:20, n.neighbors = 20, min.dist = 0.35, n.epochs = 500, spread = 2)
d4 <- DimPlot(object = Cor_FAD.SO, reduction = "umap")

```

### Cortex UMAP

left is WT and right is FAD

```{r warning=FALSE, message=FALSE, fig.width=15, fig.height=6}
grid.arrange(d3, d4, ncol = 2)
```

### Cortex Hexb (MG)

```{r warning=FALSE, message=FALSE, fig.width=15, fig.height=6}
f1 <- FeaturePlot(Cor_WT.SO, features = c("Hexb"), by.col = 2)
f2 <- FeaturePlot(Cor_FAD.SO, features = c("Hexb"), by.col = 2)
grid.arrange(f1, f2, ncol = 2)
```

### Cortex Il1- genes

WT expression of Il1- genes

```{r warning=FALSE, message=FALSE, fig.width=15, fig.height=6}
Cor_WT.SO@assays$RNA@counts@Dimnames[[1]][grepl("^Il1", Cor_WT.SO@assays$RNA@counts@Dimnames[[1]])]

```

WT expression of Il2- genes

```{r warning=FALSE, message=FALSE, fig.width=15, fig.height=6}
Cor_WT.SO@assays$RNA@counts@Dimnames[[1]][grepl("^Il2", Cor_WT.SO@assays$RNA@counts@Dimnames[[1]])]

```

FAD expression of Il1- genes

```{r warning=FALSE, message=FALSE, fig.width=15, fig.height=6}
Cor_FAD.SO@assays$RNA@counts@Dimnames[[1]][grepl("^Il1", Cor_FAD.SO@assays$RNA@counts@Dimnames[[1]])]

```

FAD expression of Il2- genes

```{r warning=FALSE, message=FALSE, fig.width=15, fig.height=6}
Cor_FAD.SO@assays$RNA@counts@Dimnames[[1]][grepl("^Il2", Cor_FAD.SO@assays$RNA@counts@Dimnames[[1]])]

```

## Naomi Hippo

In here, Ad mouse_Hippocampus_7m_All Nuclei UMIcounts [Naomi paper](https://www.nature.com/articles/s41593-020-0624-8?proof=trueIn%EF%BB%BF), [Naomi GEO](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE143758)

single nucleus RNA-seq shows in 5xFAD mouse model of AD and WT (batches = EZ Lysis Buffer + NP40 Lysis Buffer)

```{r warning=FALSE, message=FALSE, fig.width=15, fig.height=6}
Naomi <- read.table(file = "/data/rajewsky/home/skim/Microglia_Heppner/published_DAA_Habib/GSE143758_Admouse_Hippocampus_7m_AllNuclei_UMIcounts.txt.gz", row.names = 1, header = TRUE)


Naomi.SO <- CreateSeuratObject(counts =  Naomi,  min.cells = 3, min.features = 200, project = "DAA")
```


```{r warning=FALSE, message=FALSE}
# calculate MT genes
Naomi.SO[["percent.mt"]] <- PercentageFeatureSet(object = Naomi.SO, pattern = "^mt-")

```

### Naomi Hippo QC

```{r warning=FALSE, message=FALSE, fig.width=7, fig.height=6}
FeatureScatter(Naomi.SO, feature1 = "nCount_RNA", feature2 = "nFeature_RNA") + ggtitle("Naomi DAA")
```

```{r warning=FALSE, message=FALSE, fig.width=7, fig.height=6}
VlnPlot(Naomi.SO, features = c("percent.mt"), pt.size = 0.2) + ggtitle("Naomi Hippo percent.mt")
```

```{r warning=FALSE, message=FALSE, fig.width=15, fig.height=6}
Naomi.SO <- subset(x = Naomi.SO, subset = nCount_RNA < 6000 & nCount_RNA > 200 & percent.mt < 5)  

Naomi.df <- Naomi.SO@meta.data

CB_WT <- rownames(Naomi.df)[grepl("WT|Wt", rownames(Naomi.df))]
CB_AD <- rownames(Naomi.df)[grepl("AD|Ad", rownames(Naomi.df))]
CB_Unknwon <- rownames(Naomi.df)[grepl("Untreated", rownames(Naomi.df))]

Naomi.df$ID <- rownames(Naomi.df)

Naomi.df <- Naomi.df %>% mutate(exp = case_when(ID %in% CB_WT ~ "WT",
                                    ID %in% CB_AD ~ "AD",
                                    ID %in% CB_Unknwon ~ "Untreated"))

Naomi.SO$exp <- Naomi.df$exp

Naomi.df2 <- Naomi.SO@meta.data
```

```{r warning=FALSE, message=FALSE, fig.width=15, fig.height=6}

Naomi.SO <- NormalizeData(object = Naomi.SO, normalization.method = "LogNormalize", scale.factor = 1e4, verbose = FALSE)
Naomi.SO <- FindVariableFeatures(object = Naomi.SO, nfeatures = 2000, selection.method = "vst", verbose = FALSE)
Naomi.SO <- ScaleData(object = Naomi.SO, vars.to.regress = c("percent.mt", "nCount_RNA"))

# PCA
Naomi.SO <- RunPCA(Naomi.SO, features = VariableFeatures(object = Naomi.SO), verbose = FALSE, ndims.print = "None")
#ElbowPlot(Naomi.SO, ndims = 50)

Naomi.SO <- FindNeighbors(object = Naomi.SO, reduction = "pca", dims = 1:20)
Naomi.SO <- FindClusters(object = Naomi.SO, resolution = 1, verbose = FALSE)

Naomi.SO <- RunUMAP(object = Naomi.SO, dims = 1:20, n.neighbors = 20, min.dist = 0.35, n.epochs = 500, spread = 2)
d5 <- DimPlot(object = Naomi.SO, reduction = "umap")

d6 <- DimPlot(object = Naomi.SO, reduction = "umap", group.by = "exp")
```

```{r warning=FALSE, message=FALSE, fig.width=15, fig.height=6}
grid.arrange(d5, d6, ncol = 2)

```


### Naomi Hippo Hexb (MG)

```{r warning=FALSE, message=FALSE, fig.width=7, fig.height=6}
FeaturePlot(Naomi.SO, features = c("Hexb"))
```

### Naomi Hippo Il12b 

Il12b didn't exist while Il12rb2, Il12rb1 only exist

```{r warning=FALSE, message=FALSE, fig.width=15, fig.height=6}
f1 <- FeaturePlot(Naomi.SO, features = c("Il12rb1"), pt.size = 1, order = TRUE)
f2 <- FeaturePlot(Naomi.SO, features = c("Il12rb2"), pt.size = 1, order = TRUE)
grid.arrange(f1, f2, ncol = 2)
```

Naomi expression of Il12- genes in Seurat

```{r warning=FALSE, message=FALSE, fig.width=15, fig.height=6}
Naomi.SO@assays$RNA@counts@Dimnames[[1]][grepl("^Il12", Naomi.SO@assays$RNA@counts@Dimnames[[1]])]

```

Naomi expression of Il12- genes original matrix

```{r warning=FALSE, message=FALSE, fig.width=15, fig.height=6}
rownames(Naomi)[grepl("^Il12", rownames(Naomi))]
```

### Naomi Hippo Oligo 

```{r warning=FALSE, message=FALSE, fig.width=15, fig.height=12}
f1 <- FeaturePlot(Naomi.SO, features = c("Apod"), pt.size = 1, order = FALSE)
f2 <- FeaturePlot(Naomi.SO, features = c("C4b"), pt.size = 1, order = FALSE)
f3 <- FeaturePlot(Naomi.SO, features = c("Mbp"), pt.size = 1, order = FALSE)
f4 <- FeaturePlot(Naomi.SO, features = c("Mog"), pt.size = 1, order = FALSE)

grid.arrange(f1, f2, f3, f4,  ncol = 2)

```

### Naomi Hippo Neurons 

Grip1 = IN, Kcnip4 = EX

```{r warning=FALSE, message=FALSE, fig.width=15, fig.height=6}
f1 <- FeaturePlot(Naomi.SO, features = c("Grip1"), pt.size = 1, order = FALSE)
f2 <- FeaturePlot(Naomi.SO, features = c("Kcnip4"), pt.size = 1, order = FALSE)
grid.arrange(f1, f2, ncol = 2)

```

```{r warning=FALSE, message=FALSE, fig.width=15, fig.height=6}
sessionInfo()

```
