---
title: "NicheNet test"
author: "Skim"
date: "1/27/2020"
output: html_document
---
```{r warning=FALSE, message=FALSE}
library(FNN)
library(nichenetr)
library(Seurat)
library(tidyverse)
library(biomaRt)
library(dplyr)
library(stringr)
```

```{r warning=FALSE, message=FALSE}
# old resolution 0.6
#load("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200104_pre_analysis_9sets_Seurat_object.Robj")
# new resolutio 0.8 by Shirin
load("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200117_Clustering_DE_by_MAST_Seurat_object.Robj")
# data was swapped between AD and ADp40KO
#data9set.SO@meta.data$sample <- ifelse((data9set.SO$gemgroup == 1 | data9set.SO$gemgroup == 4 | data9set.SO$gemgroup == 7), "Ctrl", ifelse((data9set.SO$gemgroup == 2 | data9set.SO$gemgroup == 5 | data9set.SO$gemgroup == 8), "AD", "ADp40KO"))
# new swapped assignment
data9set.SO@meta.data$sample <- ifelse((data9set.SO$gemgroup == 1 | data9set.SO$gemgroup == 4 | data9set.SO$gemgroup == 7), "Ctrl", ifelse((data9set.SO$gemgroup == 2 | data9set.SO$gemgroup == 5 | data9set.SO$gemgroup == 8), "ADp40KO", "AD"))


data9set.SO@meta.data$sample <- factor(data9set.SO@meta.data$sample, levels = c("Ctrl", "AD", "ADp40KO"))

# sample for nichenet
data9set.SO$sample_nichenet <- as.character(data9set.SO$sample)


data9set.SO <- RenameIdents(data9set.SO, '0' = "Dentate Gyrus", '1' = "Oligo", '2' = "Neurons", '3' = "Microglia", 
                            '4' = "Astrocytes", '5' = "Oligo", '6' = "Oligo", '7' = "Inhibitory Interneurons", '8' = "Microglia", 
                            '9' = "Neuron CA2/CA3", '10' = "Dentate Gyrus", '11' = "Subiculum", '12' = "OPC", 
                            '13' = "Subiculum", '14' = "Subiculum", '15' = "Neurons", 
                            '16' = "Interneurons", '17' = "Neuron (Interneuron/subiculum)", '18' = "Neurons",
                            '19' = "Neurons", '20' = "Inhibitory Interneurons", '21' = "Inhibitory Interneurons", '22' = "Inhibitory Interneurons",
                            '23' = "Fibroblast", '24' = "Neurons", '25' = "Dentate Gyrus", '26' = "Choroid Plexus", 
                            '27' = "Neurons (Subiculum/Interneurons/ExCA)", '28' = "Cajal Rutz", '29' = "Microglia/Neuron/Fibroblast", '30' = "Inhibitory Interneurons",
                            '31' = "Dentate Gyrus", '32' = "Neurons", '33' = "Astrocytes", '34' = "Microglia",
                            '35' = "Pericytes", '36' = "Vascular Endothelial",'37' = "Microglia", '38' = "OPC",
                            '39' = "VLMC", '40' = "Neuron (subiculum/Cajal Rutz)", '41' = "OPC", '42' = "OPC", 
                            '43' = "Astrocytes", '44' = "OPC")
```


## 1. NicheNet

NicheNet can predict:  
1) which **ligands** from one cell population(**sender**) are most likely to affect **target gene** expression in an interacting cell population (**receiver**)   
2) which **specific target genes** are affected by which of these **predicted ligands**

NicheNet analysis of the following steps:  
1) Define a sender cell population and a receiver cell population / determine which genes are expressed in both populations  
2) Define a gene set of interest, these are the genes in the **receiver** cell population (e.g. genes differentially expressed upon cell-cell interaction)  
3) Define a set of potential ligands (ligands that are expressed by the **sender** cell population)  
4) Perform NicheNet ligand activity analysis (rank the potential ligands)  
5) Infer top-predicted target genes of ligands that are top-ranked in the ligand activity analysis  

I followed below link [NicheNet Seurat](https://github.com/saeyslab/nichenetr/blob/master/vignettes/seurat_steps.md)

#### 1-1. NicheNet’s ligand-target prior model, ligand-receptor network and weighted integrated networks

#### 1-2. Define a sender cell population and a receiver cell population 

also determine which genes are expressed in both populations

```{r NicheNet_dataload, warning=FALSE, message=FALSE}
# ligand-target prior model
ligand_target_matrix = readRDS(url("https://zenodo.org/record/3260758/files/ligand_target_matrix.rds"))
# ligand-receptor network
lr_network = readRDS(url("https://zenodo.org/record/3260758/files/lr_network.rds"))
# weighted integrated networks
weighted_networks = readRDS(url("https://zenodo.org/record/3260758/files/weighted_networks.rds"))
weighted_networks_lr = weighted_networks$lr_sig %>% inner_join(lr_network %>% distinct(from,to), by = c("from","to"))

# convert human to mouse symbol
#lr_network = lr_network %>% mutate(from = convert_human_to_mouse_symbols(from), to = convert_human_to_mouse_symbols(to)) %>% drop_na()
#colnames(ligand_target_matrix) = ligand_target_matrix %>% colnames() %>% convert_human_to_mouse_symbols()
#rownames(ligand_target_matrix) = ligand_target_matrix %>% rownames() %>% convert_human_to_mouse_symbols()

#ligand_target_matrix = ligand_target_matrix %>% .[!is.na(rownames(ligand_target_matrix)), !is.na(colnames(ligand_target_matrix))]

#weighted_networks_lr = weighted_networks_lr %>% mutate(from = convert_human_to_mouse_symbols(from), to = convert_human_to_mouse_symbols(to)) %>% drop_na()
#head(seuratObj@assays$RNA@data@Dimnames[[1]])
```

```{r NicheNet, warning=FALSE, message=FALSE}
# cellphonedb data create for Oligodendrocytes (1,5,6) and microglia (3, 8, 29, 34, 37)
data9set_CP.SO <- subset(data9set.SO, subset = seurat_clusters == c(1,5,6,3,8,29,34,37))

nichenet_output2 = nichenet_seuratobj_aggregate(seurat_obj = data9set_CP.SO, 
                                               receiver = "Oligo", 
                                               condition_colname = "sample_nichenet", 
                                               condition_oi = "AD", 
                                               condition_reference = "Ctrl", 
                                               sender = "Microglia", 
                                               ligand_target_matrix = ligand_target_matrix, lr_network = lr_network, weighted_networks = weighted_networks, organism = "mouse")

nichenet_output$ligand_activities
nichenet_output$ligand_activity_target_heatmap
```
```{r NicheNet, warning=FALSE, message=FALSE}
nichenet_output2$ligand_activity_target_heatmap

nichenet_output2$ligand_receptor_heatmap
nichenet_output2$ligand_target_heatmap
list(nichenet_output2)
nichenet_output2$ligand_target_heatmap
#FeaturePlot(data9set.SO, features = c("Vim", "Actb", "App", "Ccl5"), pt.size = 0.1)
```
```{r NicheNet, warning=FALSE, message=FALSE}
##################
# receiver e.g. 3
#################
receiver = "3"
expressed_genes_receiver = get_expressed_genes(receiver, data9set.SO, pct = 0.10)

background_expressed_genes = expressed_genes_receiver %>% .[. %in% rownames(ligand_target_matrix)]

#####################
# sender e.g. 2, 12
#####################
sender_celltypes = c("2","12")

list_expressed_genes_sender = sender_celltypes %>% unique() %>% lapply(get_expressed_genes, data9set.SO, 0.10) # lapply to get the expressed genes of every sender cell type separately here
expressed_genes_sender = list_expressed_genes_sender %>% unlist() %>% unique()


??topGO::genTable
```

#### 1-3. Define a gene set of interest

these are the genes in the receiver cell population that are potentially affected by ligands expressed by interacting cells (e.g. genes differentially expressed upon cell-cell interaction)

```{r warning=FALSE, message=FALSE}
seurat_obj_receiver= subset(data9set.SO, idents = receiver)

condition_oi = "AD"
condition_reference = "Ctrl" 
  
DE_table_receiver = FindMarkers(object = data9set.SO, ident.1 = condition_oi, ident.2 = condition_reference, group.by = "sample", min.pct = 0.10) %>% rownames_to_column("gene")

geneset_oi = DE_table_receiver %>% filter(p_val_adj <= 0.05 & abs(avg_logFC) >= 0.25) %>% pull(gene)
geneset_oi = geneset_oi %>% .[. %in% rownames(ligand_target_matrix)]
```

#### 1-4. Define a set of potential ligands

these are ligands that are expressed by the sender cell population and bind a (putative) receptor expressed by the receiver population. pooling all ligands from all cell types together

```{r warning=FALSE, message=FALSE}
ligands = lr_network %>% pull(from) %>% unique()
receptors = lr_network %>% pull(to) %>% unique()

expressed_ligands = intersect(ligands,expressed_genes_sender)
expressed_receptors = intersect(receptors,expressed_genes_receiver)

potential_ligands = lr_network %>% filter(from %in% expressed_ligands & to %in% expressed_receptors)
```

#### 1-5. NicheNet ligand activity analysis

rank the potential ligands based on the presence of their target genes in the gene set of interest

```{r warning=FALSE, message=FALSE}
ligand_activities = predict_ligand_activities(geneset = geneset_oi, 
                                              background_expressed_genes = background_expressed_genes,
                                              ligand_target_matrix = ligand_target_matrix, 
                                              potential_ligands = potential_ligands)

ligand_activities = ligand_activities %>% arrange(-pearson) %>% mutate(rank = rank(desc(pearson)))
ligand_activities
```

```{r warning=FALSE, message=FALSE}
nichenet_output = nichenet_seuratobj_aggregate(
  seurat_obj = data9set.SO, 
  receiver = "3", 
  condition_colname = "aggregate", condition_oi = "AD", condition_reference = "Ctrl", 
  sender = c("2","12"), 
  ligand_target_matrix = ligand_target_matrix, lr_network = lr_network, weighted_networks = weighted_networks, organism = "mouse")

```
