---
title: "pie chart vs bar plot"
author: "Skim"
date: "7/9/2020"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_depth: 4
    toc_float: false
    code_folding: hide
---

```{r warning=FALSE, message=FALSE}
library(readr)
library(dplyr)
library(tibble)
library(ggplot2)
library(ggrepel)
library(forcats)
```



```{r warning=FALSE, message=FALSE}
cell_counts <- data.frame(dataset = c("reference", "reference", "reference", "reference", "reference", "reference", "data", "data", "data", "data", "data", "data"),
                        cell_type = c("Neurons", "Astrocytes", "Microglia","Oligodendrocytes", "OPC","rest", "Neurons", "Astrocytes", "Microglia","Oligodendrocytes", "OPC","rest"), 
                        percent = c(68, 7.5, 11.99, 12.01, 0, 0.4, 61, 4, 21, 9, 3, 2))

```

```{r warning=FALSE, message=FALSE}
# Order cell types by total cell count.
cell_type_order <- 
  cell_counts %>%
  group_by(cell_type) %>%
  summarize(percent = sum(percent)) %>%
  arrange(-percent) %>%
  pull(cell_type) %>%
  as.character

```



```{r warning=FALSE, message=FALSE}

# Prepare data for plotting.
plot.data <-
  cell_counts %>%
  mutate(cell_type = factor(as.character(cell_type), cell_type_order)) %>%
  group_by(dataset) %>%
  arrange(-as.integer(cell_type)) %>%
  mutate(cell_fraction = percent / sum(percent),
         label_pos =
           cumsum(lag(cell_fraction, default = 0)) + cell_fraction / 2) %>%
  ungroup %>%
  mutate(cell_fraction_label = paste(round(100 * cell_fraction), "%"),
         n_cells_label = format(percent, big.mark = " "))

```

```{r warning=FALSE, message=FALSE}
# Plot cell fractions colored by cell type.
base_plot <-
  plot.data %>%
  ggplot(aes(y = cell_fraction)) %>%
  + ylab("cell fraction") %>%
  + scale_fill_brewer("cell type", palette = "Dark2") %>%
  + theme_minimal()

```

Questions from Marcel

1. In which dataset does each cell type account for a larger proportion?  
2. Clearly, there are more fibroblast(other cell type?) in your data compared to the reference. But what is the fold-change?  
3. Are there any neuronal progenitor cells in the reference data? (No for now, but maybe?)    
4. Is the proportion of fibroblasts(other cell type ?) in your data larger than that of oligodendrocyte in the reference data?  

```{r warning=FALSE, message=FALSE}
# Version 1: Clean pie-chart.
pie_chart <-
  base_plot %>%
  + facet_wrap(~dataset) %>%
  + geom_bar(aes(x = 1, fill = cell_type), width = 1, stat = "identity") %>%
  + coord_polar("y", direction = -1) %>%
  + xlab(NULL) %>%
  + theme(axis.text = element_blank(), line = element_blank(),
      strip.text.x = element_text(size = 12, face = "bold"))
pie_chart


```


Marcel: Do you still like this messier? 

```{r warning=FALSE, message=FALSE}

# Version 2: Annotated pie-chart.
pie_chart %>%
  + geom_label_repel(aes(x = 1, y = label_pos, label = n_cells_label, fill = cell_type),
                    direction = "y", segment.size = 0, box.padding = 0, show.legend = FALSE) %>%
  + geom_label_repel(aes(x = 1.5, y = label_pos, label = cell_fraction_label,
                         fill = cell_type),
                     nudge_x = .5, direction = "x", show.legend = FALSE)

```

```{r warning=FALSE, message=FALSE}
# Version 3: Clean bar chart:
bar_chart <-
  base_plot %>%
  + geom_bar(aes(x = cell_type, fill = cell_type, alpha = fct_rev(dataset)),
             stat = "identity", position = "dodge") %>%
  + scale_alpha_discrete("dataset", range = c(.6, 1),
                         guide = guide_legend(reverse = TRUE)) %>%
  + guides(fill = FALSE) %>%
  + scale_x_discrete("cell type", limits = rev(cell_type_order)) %>%
  + coord_flip()
bar_chart


```

Marcel: Note how much cleaner this is, even if you add all the extra

```{r warning=FALSE, message=FALSE}

# Version 4: Annotated bar chart:
bar_chart %>%
  + geom_text(aes(x = cell_type,
                  label = paste0( n_cells_label),
                  alpha = fct_rev(dataset)),
              position = position_dodge(width = .9), hjust = 1.1,
              show.legend = FALSE) %>%
  + geom_text(aes(x = cell_type,
                  label = paste0("[", cell_fraction_label, "]"),
                  alpha = fct_rev(dataset)),
              position = position_dodge(width = .9), hjust = -.1,
              show.legend = FALSE)

```