---
title: "Subiculum"
author: "Skim"
date: '2020 6 18 '
output: 
  html_document:
    df_print: paged
    toc: true
    toc_depth: 4
    toc_float: false
    code_folding: hide
---

```{r warning=FALSE, message=FALSE}
.libPaths(c("/home/skim/R/usr_lib", .libPaths()))
library(Seurat)
library(ggplot2)
library(gridExtra)
library(grid)
library(ggrepel)
library(DT)
library(biomaRt)
library(dplyr)
library(viridis)
library(tidyr)
source("/data/rajewsky/home/skim/Microglia_Heppner/Codes/DroNc_AD/src/TopGO_scRNA.R")
```

Below is each cluster in Neuronal cell types

Dentate Gyrus = 0, 10  
CA1 = 2, 13, 18, 40  
CA2/3 = 9, 21, 27  
Subiculum = 11, 14, 15, 17, 19, 20  
Inhibitory = 7, 16, 22, 30  
Neurons = 24, 29, 32


```{r warning=FALSE, message=FALSE}
load(file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200221_figure1_for_meeting_report_Seurat_object.Robj")
```

```{r warning=FALSE, message=FALSE}
cluster_GO <- function(genesOfInterest, geneUniverse){

geneList <- factor(as.integer(geneUniverse %in% genesOfInterest))
names(geneList) <- geneUniverse

onts = c( "MF", "BP", "CC" )
tab <- as.list(onts)
names(tab) <- onts

for(i in 1:3){

GOdata <- new("topGOdata",
                description = "GOanalysis",
                ontology = onts[i],
                allGenes = geneList,
                annot = annFUN.org,
                mapping = "org.Mm.eg.db",
                ID = "SYMBOL",
                nodeSize = 20)

# export GO ID from BP
if(i == 1){
  allGO_MF <- genesInTerm(GOdata)
}
if(i == 2){
  allGO_BP <- genesInTerm(GOdata)
}

if(i == 3){
  allGO_CC <- genesInTerm(GOdata)
}

res.result1 <- runTest(GOdata, statistic = "fisher", algorithm = "elim")
res.result2 <- runTest(GOdata, statistic = "fisher", algorithm = "classic")


tab[[i]] <- cbind(data.frame(category = onts[i]), GenTable(GOdata, Fisher.elim = res.result1,
                        Fisher.classic = res.result2,
                        orderBy = "Fisher.elim" , topNodes = 30))


}

topGOResults <- plyr::rbind.fill(tab)
topGOResults.df <- as.data.frame(topGOResults)
topGOResults.df$gene_ratio <- topGOResults.df$Significant / topGOResults.df$Annotated

# modification appropriate for plot
topGOResults.df$Fisher.elim <- as.numeric(topGOResults.df$Fisher.elim)
topGOResults.df$Fisher.classic <- as.numeric(topGOResults.df$Fisher.classic)
topGOResults.df$Term <- factor(topGOResults.df$Term, levels = rev(unique(topGOResults.df$Term)))

for(i in 1:30){
 ID <- topGOResults.df$GO.ID[i]
 #print(ID)
 #print(paste(genesOfInterest[genesOfInterest %in% allGO[ID][[1]]], collapse=', ' ))
 topGOResults.df[i, "genes"] <- paste(genesOfInterest[genesOfInterest %in% allGO_MF[ID][[1]]], collapse=', ' )
}
for(i in 31:60){
 ID <- topGOResults.df$GO.ID[i]
 #print(ID)
 #print(paste(genesOfInterest[genesOfInterest %in% allGO[ID][[1]]], collapse=', ' ))
 topGOResults.df[i, "genes"] <- paste(genesOfInterest[genesOfInterest %in% allGO_BP[ID][[1]]], collapse=', ' )
}
for(i in 61:90){
 ID <- topGOResults.df$GO.ID[i]
 #print(ID)
 #print(paste(genesOfInterest[genesOfInterest %in% allGO[ID][[1]]], collapse=', ' ))
 topGOResults.df[i, "genes"] <- paste(genesOfInterest[genesOfInterest %in% allGO_CC[ID][[1]]], collapse=', ' )
}

return(topGOResults.df)
}

# ggplot function
scRNA_TopGO_plot2 <- function(topGOResults.df){

topGOResults.df$Fisher.elim <- as.numeric(topGOResults.df$Fisher.elim)
topGOResults.df$Fisher.elim <- -log10(topGOResults.df$Fisher.elim)
topGOResults.df$Term <- factor(topGOResults.df$Term, levels = rev(unique(topGOResults.df$Term)))

ggplot(topGOResults.df, aes(x=gene_ratio,
               y=Term,
               colour=Fisher.elim,
               size=Significant)) +
        geom_point() +
        expand_limits(x=0) +
        labs(x="gene ratio", y="GO term", colour="-log10(p-value)", size="Significant") +
  theme_minimal() + theme(axis.text = element_text(size = 10))

}
```

### 1. Subiculum overview

```{r warning=FALSE, message=FALSE}
data9set_subi.SO <- subset(data9set_cleaned.SO, subset = seurat_clusters %in% c(11, 14, 15, 17, 19, 20))

```

```{r warning=FALSE, message=FALSE, fig.width=6.5, fig.height=6}
data9set_cleaned.meta.data <- data9set_cleaned.SO@meta.data
data9set_cleaned_umap.df <- as.data.frame(data9set_cleaned.SO@reductions$umap@cell.embeddings)
data9set_cleaned_umap.df <- cbind(data9set_cleaned.meta.data, data9set_cleaned_umap.df)

# Mark CA1 cluster
data9set_cleaned_umap.df$ct <- ifelse(data9set_cleaned_umap.df$seurat_clusters %in% c(11, 14, 15, 17, 19, 20), "subi", "others")


# Microglia whole UMAP
ggplot(data9set_cleaned_umap.df, aes(x = UMAP_1, y = UMAP_2)) + geom_point(color = "gray", size = 0.1) + 
  geom_point(data = data9set_cleaned_umap.df[data9set_cleaned_umap.df$ct %in% "subi", ], aes(x = UMAP_1, y = UMAP_2), color = "#CC0000", size = 0.1) + 
  theme_classic() + 
  theme(axis.title = element_text(color = "black", size =15),axis.text = element_text(color = "black", size =12))
```

```{r  warning=FALSE, message=FALSE, fig.width=15, fig.height = 5}
DimPlot(data9set_subi.SO, split.by = "sample", group.by = "seurat_clusters") + xlim(c(-15, 10)) + ylim(c(-5, 18))
```

cell type proportion

```{r warning=FALSE, message=FALSE, fig.width=8.5, fig.height=5}
plot.data <- data9set_cleaned.SO@meta.data %>%
    dplyr::select(sample, gemgroup, cluster = seurat_clusters) %>%
    mutate(cluster = cluster) %>%
    group_by(gemgroup, cluster, sample) %>%
  summarise(count = n()) %>%
  group_by(gemgroup) %>%
    mutate(gemgroup_total = sum(count)) %>%
    mutate(cell_prop = count / gemgroup_total)

ggplot(plot.data[plot.data$cluster %in% c(11, 14, 15, 17, 19, 20),], aes(x=cluster, y=cell_prop, fill=sample)) +
  geom_boxplot(position=position_dodge(1)) + ylim(c(0, 0.08))
```

```{r  warning=FALSE, message=FALSE, fig.width=8.5, fig.height=5}
Idents(data9set_subi.SO) <- data9set_subi.SO$seurat_clusters

# c11.markers <- FindMarkers(data9set_subi.SO, ident.1 = 11, ident.2 = c(14, 15, 17, 19, 20), min.pct = 0.05)
# write.csv(c11.markers, file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200618_subiculum_neurons_analysis_cluster11_markers.csv")
c11.markers <- read.csv(file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200618_subiculum_neurons_analysis_cluster11_markers.csv", row.names = 1)

# c14.markers <- FindMarkers(data9set_subi.SO, ident.1 = 14, ident.2 = c(11, 15, 17, 19, 20), min.pct = 0.05)
# write.csv(c14.markers, file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200618_subiculum_neurons_analysis_cluster14_markers.csv")
c14.markers <- read.csv(file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200618_subiculum_neurons_analysis_cluster14_markers.csv", row.names = 1)

# c15.markers <- FindMarkers(data9set_subi.SO, ident.1 = 15, ident.2 = c(11, 14, 17, 19, 20), min.pct = 0.05)
# write.csv(c15.markers, file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200618_subiculum_neurons_analysis_cluster15_markers.csv")
c15.markers <- read.csv(file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200618_subiculum_neurons_analysis_cluster15_markers.csv", row.names = 1)

# c17.markers <- FindMarkers(data9set_subi.SO, ident.1 = 17, ident.2 = c(11, 14, 15, 19, 20), min.pct = 0.05)
# write.csv(c17.markers, file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200618_subiculum_neurons_analysis_cluster17_markers.csv")
c17.markers <- read.csv(file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200618_subiculum_neurons_analysis_cluster17_markers.csv", row.names = 1)

# c19.markers <- FindMarkers(data9set_subi.SO, ident.1 = 19, ident.2 = c(11, 14, 15, 17, 20), min.pct = 0.05)
# write.csv(c19.markers, file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200618_subiculum_neurons_analysis_cluster19_markers.csv")
c19.markers <- read.csv(file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200618_subiculum_neurons_analysis_cluster19_markers.csv", row.names = 1)

# c20.markers <- FindMarkers(data9set_subi.SO, ident.1 = 20, ident.2 = c(11, 14, 15, 17, 19), min.pct = 0.05)
# write.csv(c20.markers, file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200618_subiculum_neurons_analysis_cluster20_markers.csv")
c20.markers <- read.csv(file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200618_subiculum_neurons_analysis_cluster20_markers.csv", row.names = 1)

```

### 2. each cluster

#### 2-1. Cluster 11 specific genes

```{r  warning=FALSE, message=FALSE, fig.width=13, fig.height = 12}
f1 <- FeaturePlot(data9set_subi.SO, features = "Dlc1") + xlim(c(-15, 10)) + ylim(c(-5, 18))
f2 <- FeaturePlot(data9set_subi.SO, features = "Hs3st4") + xlim(c(-15, 10)) + ylim(c(-5, 18))
f3 <- FeaturePlot(data9set_subi.SO, features = "Cdh18") + xlim(c(-15, 10)) + ylim(c(-5, 18))
f4 <- FeaturePlot(data9set_subi.SO, features = "Zfpm2") + xlim(c(-15, 10)) + ylim(c(-5, 18))

grid.arrange(f1, f2, f3, f4, ncol = 2)
```

```{r warning=FALSE, message=FALSE, fig.width= 15, fig.height= 5}
# geneUniverse <- rownames(data9set_subi.SO@assays$RNA@counts[rowSums(data9set_subi.SO@assays$RNA@counts) > 0, ]) %>% as.character
# 
# genesOfInterest <- as.character(rownames(c11.markers[which(c11.markers$p_val_adj < 0.01 & c11.markers$avg_logFC > 0.5 & c11.markers$pct.2 < 0.5), ]))
# 
# topGOResults.df <- cluster_GO(genesOfInterest, geneUniverse)
# 
# write.csv(topGOResults.df, file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200618_subiculum_neurons_analysis_GO_cluster11.csv")

topGOResults.df <- read.csv(file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200618_subiculum_neurons_analysis_GO_cluster11.csv", row.names = 1)

datatable(topGOResults.df, options = list(pageLength = 5, dom = 'tip'))
```

```{r warning=FALSE, message=FALSE, fig.width= 15, fig.height= 7}
topGOResults_MF.df <- topGOResults.df[which(topGOResults.df$category == "MF" & topGOResults.df$Annotated < 1000 & topGOResults.df$Annotated > 30), ][c(1:15), ]
g1 <- scRNA_TopGO_plot2(topGOResults_MF.df)
g1 <- g1 + ggtitle("Molecular function")

topGOResults_BP.df <- topGOResults.df[which(topGOResults.df$category == "BP" & topGOResults.df$Annotated < 1000 & topGOResults.df$Annotated > 30), ][c(1:15), ]
g2 <- scRNA_TopGO_plot2(topGOResults_BP.df)
g2 <- g2 + ggtitle("Biological process")

grid.arrange(g1, g2, ncol = 2)
```

#### 2-2. Cluster 14 specific genes

```{r  warning=FALSE, message=FALSE, fig.width=13, fig.height = 12}
f1 <- FeaturePlot(data9set_subi.SO, features = "Gpc6") + xlim(c(-15, 10)) + ylim(c(-5, 18))
f2 <- FeaturePlot(data9set_subi.SO, features = "Cdh20") + xlim(c(-15, 10)) + ylim(c(-5, 18))
f3 <- FeaturePlot(data9set_subi.SO, features = "Gm2164") + xlim(c(-15, 10)) + ylim(c(-5, 18))
f4 <- FeaturePlot(data9set_subi.SO, features = "Pdzrn3") + xlim(c(-15, 10)) + ylim(c(-5, 18))

grid.arrange(f1, f2, f3, f4, ncol = 2)
```

```{r warning=FALSE, message=FALSE, fig.width= 15, fig.height= 5}
# genesOfInterest <- as.character(rownames(c14.markers[which(c14.markers$p_val_adj < 0.01 & c14.markers$avg_logFC > 0.5 & c14.markers$pct.2 < 0.5), ]))
# 
# topGOResults.df <- cluster_GO(genesOfInterest, geneUniverse)
# 
# write.csv(topGOResults.df, file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200618_subiculum_neurons_analysis_GO_cluster14.csv")

topGOResults.df <- read.csv(file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200618_subiculum_neurons_analysis_GO_cluster14.csv", row.names = 1)

datatable(topGOResults.df, options = list(pageLength = 5, dom = 'tip'))
```

```{r warning=FALSE, message=FALSE, fig.width= 15, fig.height= 7}
topGOResults_MF.df <- topGOResults.df[which(topGOResults.df$category == "MF" & topGOResults.df$Annotated < 1000 & topGOResults.df$Annotated > 30), ][c(1:15), ]
g1 <- scRNA_TopGO_plot2(topGOResults_MF.df)
g1 <- g1 + ggtitle("Molecular function")

topGOResults_BP.df <- topGOResults.df[which(topGOResults.df$category == "BP" & topGOResults.df$Annotated < 1000 & topGOResults.df$Annotated > 30), ][c(1:15), ]
g2 <- scRNA_TopGO_plot2(topGOResults_BP.df)
g2 <- g2 + ggtitle("Biological process")

grid.arrange(g1, g2, ncol = 2)
```

#### 2-3. Cluster 15 specific genes

```{r  warning=FALSE, message=FALSE, fig.width=13, fig.height = 12}
f1 <- FeaturePlot(data9set_subi.SO, features = "Cdh12") + xlim(c(-15, 10)) + ylim(c(-5, 18))
f2 <- FeaturePlot(data9set_subi.SO, features = "Tafa1") + xlim(c(-15, 10)) + ylim(c(-5, 18))
f3 <- FeaturePlot(data9set_subi.SO, features = "Kcnt2") + xlim(c(-15, 10)) + ylim(c(-5, 18))
f4 <- FeaturePlot(data9set_subi.SO, features = "Fstl4") + xlim(c(-15, 10)) + ylim(c(-5, 18))

grid.arrange(f1, f2, f3, f4, ncol = 2)
```

```{r warning=FALSE, message=FALSE, fig.width= 15, fig.height= 5}
# genesOfInterest <- as.character(rownames(c15.markers[which(c15.markers$p_val_adj < 0.01 & c15.markers$avg_logFC > 0.5 & c15.markers$pct.2 < 0.5), ]))
# 
# topGOResults.df <- cluster_GO(genesOfInterest, geneUniverse)
# 
# write.csv(topGOResults.df, file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200618_subiculum_neurons_analysis_GO_cluster15.csv")

topGOResults.df <- read.csv(file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200618_subiculum_neurons_analysis_GO_cluster15.csv", row.names = 1)

datatable(topGOResults.df, options = list(pageLength = 5, dom = 'tip'))
```

```{r warning=FALSE, message=FALSE, fig.width= 15, fig.height= 7}
topGOResults_MF.df <- topGOResults.df[which(topGOResults.df$category == "MF" & topGOResults.df$Annotated < 1000 & topGOResults.df$Annotated > 30), ][c(1:15), ]
g1 <- scRNA_TopGO_plot2(topGOResults_MF.df)
g1 <- g1 + ggtitle("Molecular function")

topGOResults_BP.df <- topGOResults.df[which(topGOResults.df$category == "BP" & topGOResults.df$Annotated < 1000 & topGOResults.df$Annotated > 30), ][c(1:15), ]
g2 <- scRNA_TopGO_plot2(topGOResults_BP.df)
g2 <- g2 + ggtitle("Biological process")

grid.arrange(g1, g2, ncol = 2)
```

#### 2-4. Cluster 17 specific genes

```{r  warning=FALSE, message=FALSE, fig.width=13, fig.height = 12}
f1 <- FeaturePlot(data9set_subi.SO, features = "Gm20754") + xlim(c(-15, 10)) + ylim(c(-5, 18))
f2 <- FeaturePlot(data9set_subi.SO, features = "Cntnap5b") + xlim(c(-15, 10)) + ylim(c(-5, 18))
f3 <- FeaturePlot(data9set_subi.SO, features = "Gpc5") + xlim(c(-15, 10)) + ylim(c(-5, 18))
f4 <- FeaturePlot(data9set_subi.SO, features = "4930509J09Rik") + xlim(c(-15, 10)) + ylim(c(-5, 18))

grid.arrange(f1, f2, f3, f4, ncol = 2)
```

```{r warning=FALSE, message=FALSE, fig.width= 15, fig.height= 5}
# genesOfInterest <- as.character(rownames(c17.markers[which(c17.markers$p_val_adj < 0.01 & c17.markers$avg_logFC > 0.5 & c17.markers$pct.2 < 0.5), ]))
# 
# topGOResults.df <- cluster_GO(genesOfInterest, geneUniverse)
# 
# write.csv(topGOResults.df, file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200618_subiculum_neurons_analysis_GO_cluster17.csv")

topGOResults.df <- read.csv(file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200618_subiculum_neurons_analysis_GO_cluster17.csv", row.names = 1)

datatable(topGOResults.df, options = list(pageLength = 5, dom = 'tip'))
```

```{r warning=FALSE, message=FALSE, fig.width= 15, fig.height= 7}
topGOResults_MF.df <- topGOResults.df[which(topGOResults.df$category == "MF" & topGOResults.df$Annotated < 1000 & topGOResults.df$Annotated > 30), ][c(1:12), ]
g1 <- scRNA_TopGO_plot2(topGOResults_MF.df)
g1 <- g1 + ggtitle("Molecular function")

topGOResults_BP.df <- topGOResults.df[which(topGOResults.df$category == "BP" & topGOResults.df$Annotated < 1000 & topGOResults.df$Annotated > 30), ][c(1:15), ]
g2 <- scRNA_TopGO_plot2(topGOResults_BP.df)
g2 <- g2 + ggtitle("Biological process")

grid.arrange(g1, g2, ncol = 2)
```

#### 2-5. Cluster 19 specific genes

```{r  warning=FALSE, message=FALSE, fig.width=13, fig.height = 12}
f1 <- FeaturePlot(data9set_subi.SO, features = "Cntn6") + xlim(c(-15, 10)) + ylim(c(-5, 18))
f2 <- FeaturePlot(data9set_subi.SO, features = "Kcnip1") + xlim(c(-15, 10)) + ylim(c(-5, 18))
f3 <- FeaturePlot(data9set_subi.SO, features = "Grik1") + xlim(c(-15, 10)) + ylim(c(-5, 18))
f4 <- FeaturePlot(data9set_subi.SO, features = "Sntb1") + xlim(c(-15, 10)) + ylim(c(-5, 18))

grid.arrange(f1, f2, f3, f4, ncol = 2)
```

```{r warning=FALSE, message=FALSE, fig.width= 15, fig.height= 5}
# genesOfInterest <- as.character(rownames(c19.markers[which(c19.markers$p_val_adj < 0.01 & c19.markers$avg_logFC > 0.5 & c19.markers$pct.2 < 0.5), ]))
# 
# topGOResults.df <- cluster_GO(genesOfInterest, geneUniverse)
# 
# write.csv(topGOResults.df, file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200618_subiculum_neurons_analysis_GO_cluster19.csv")

topGOResults.df <- read.csv(file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200618_subiculum_neurons_analysis_GO_cluster19.csv", row.names = 1)

datatable(topGOResults.df, options = list(pageLength = 5, dom = 'tip'))
```

```{r warning=FALSE, message=FALSE, fig.width= 15, fig.height= 7}
topGOResults_MF.df <- topGOResults.df[which(topGOResults.df$category == "MF" & topGOResults.df$Annotated < 1000 & topGOResults.df$Annotated > 30), ][c(1:15), ]
g1 <- scRNA_TopGO_plot2(topGOResults_MF.df)
g1 <- g1 + ggtitle("Molecular function")

topGOResults_BP.df <- topGOResults.df[which(topGOResults.df$category == "BP" & topGOResults.df$Annotated < 1000 & topGOResults.df$Annotated > 30), ][c(1:15), ]
g2 <- scRNA_TopGO_plot2(topGOResults_BP.df)
g2 <- g2 + ggtitle("Biological process")

grid.arrange(g1, g2, ncol = 2)
```

#### 2-6. Cluster 20 specific genes

Sox2 overlapping transcript (Sox2OT) participate in progress and development in AD brains [ref](https://www.frontiersin.org/articles/10.3389/fgene.2018.00744/full)

Acute Inhibition of ErbB4 Impairs GABA Transmission [ref](https://www.cell.com/neuron/pdfExtended/S0896-6273(18)30195-8)

up-regulating of ErbB4 immunoreactivity may involve in the progression of pathology of AD [ref](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3145840/)

```{r  warning=FALSE, message=FALSE, fig.width=13, fig.height = 12}
f1 <- FeaturePlot(data9set_subi.SO, features = "Erbb4") + xlim(c(-15, 10)) + ylim(c(-5, 18))
f2 <- FeaturePlot(data9set_subi.SO, features = "Sox2ot") + xlim(c(-15, 10)) + ylim(c(-5, 18))
f3 <- FeaturePlot(data9set_subi.SO, features = "Adarb2") + xlim(c(-15, 10)) + ylim(c(-5, 18))
f4 <- FeaturePlot(data9set_subi.SO, features = "Zfhx3") + xlim(c(-15, 10)) + ylim(c(-5, 18))

grid.arrange(f1, f2, f3, f4, ncol = 2)
```

```{r warning=FALSE, message=FALSE, fig.width= 15, fig.height= 5}
# genesOfInterest <- as.character(rownames(c20.markers[which(c20.markers$p_val_adj < 0.01 & c20.markers$avg_logFC > 0.5 & c20.markers$pct.2 < 0.5), ]))
# 
# topGOResults.df <- cluster_GO(genesOfInterest, geneUniverse)
# 
# write.csv(topGOResults.df, file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200618_subiculum_neurons_analysis_GO_cluster20.csv")

topGOResults.df <- read.csv(file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200618_subiculum_neurons_analysis_GO_cluster20.csv", row.names = 1)

datatable(topGOResults.df, options = list(pageLength = 5, dom = 'tip'))
```

```{r warning=FALSE, message=FALSE, fig.width= 15, fig.height= 7}
topGOResults_MF.df <- topGOResults.df[which(topGOResults.df$category == "MF" & topGOResults.df$Annotated < 1000 & topGOResults.df$Annotated > 30), ][c(1:15), ]
g1 <- scRNA_TopGO_plot2(topGOResults_MF.df)
g1 <- g1 + ggtitle("Molecular function")

topGOResults_BP.df <- topGOResults.df[which(topGOResults.df$category == "BP" & topGOResults.df$Annotated < 1000 & topGOResults.df$Annotated > 30), ][c(1:15), ]
g2 <- scRNA_TopGO_plot2(topGOResults_BP.df)
g2 <- g2 + ggtitle("Biological process")

grid.arrange(g1, g2, ncol = 2)
```

### 3. DE in subiculum neurons

```{r warning=FALSE, message=FALSE, fig.width= 15, fig.height= 7}

# subi_DE_AD_ADp40KO <- FindMarkers(data9set_subi.SO, ident.1 = "AD", ident.2 = "ADp40KO", group.by = "sample", min.pct = 0.01, logfc.threshold = 0.01)
# write.csv(subi_DE_AD_ADp40KO, file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200618_subiculum_neurons_analysis_DE_AD_ADp40KO.csv")

subi_DE_AD_ADp40KO <- read.csv(file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200618_subiculum_neurons_analysis_DE_AD_ADp40KO.csv")


subi_DE_AD_ADp40KO2 <- subi_DE_AD_ADp40KO[abs(subi_DE_AD_ADp40KO$avg_logFC) > 0.1, ]
# remove Ttr
subi_DE_AD_ADp40KO2 <- subi_DE_AD_ADp40KO2[subi_DE_AD_ADp40KO2$X != "Ttr", ]

# only logFC > 0.1
subi_DE_AD_ADp40KO_padj <- subi_DE_AD_ADp40KO[abs(subi_DE_AD_ADp40KO$avg_logFC) > 0.1, ]
# remove Ttr
subi_DE_AD_ADp40KO_padj <- subi_DE_AD_ADp40KO_padj[subi_DE_AD_ADp40KO_padj$X != "Ttr", ]

subi_DE_AD_ADp40KO_padj <- subi_DE_AD_ADp40KO_padj[which(subi_DE_AD_ADp40KO_padj$p_val_adj < 0.05), ]

datatable(subi_DE_AD_ADp40KO_padj, options = list(pageLength = 5, dom = 'tip'))
```

```{r warning=FALSE, message=FALSE, fig.width= 5.5, fig.height= 5}

ggplot(subi_DE_AD_ADp40KO2) + geom_point(aes(x = avg_logFC, y = -log10(p_val_adj))) + 
  geom_point(data = subi_DE_AD_ADp40KO_padj, aes(x = avg_logFC, y = -log10(p_val_adj)), color = "red") + theme_classic() +
  ggtitle("Subiculum Neurons DE analysis AD vs ADp40KO") + 
   xlab("log2 fold change") + 
  ylab("-log10(adjusted p-value)") +
  theme(legend.position = "none",
        plot.title = element_text(size = 15, hjust = 0.5, family = "helvetica"),
        axis.title = element_text(size = 15), 
        axis.text = element_text(size = 12, color = "black")) +
  geom_text_repel(data = subi_DE_AD_ADp40KO_padj[subi_DE_AD_ADp40KO_padj$p_val_adj < 1e-20, ], aes(x = avg_logFC, y = -log10(p_val_adj), label = subi_DE_AD_ADp40KO_padj[subi_DE_AD_ADp40KO_padj$p_val_adj < 1e-20, ]$X)) 
```

GO of AD vs ADp40KO up-regulated genes

```{r warning=FALSE, message=FALSE, fig.width= 15, fig.height= 5}
# genesOfInterest <- subi_DE_AD_ADp40KO_padj[subi_DE_AD_ADp40KO_padj$avg_logFC > 0, ]$X %>% as.character
# 
# 
# topGOResults.df <- cluster_GO(genesOfInterest, geneUniverse)
# 
# write.csv(topGOResults.df, file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200618_subiculum_neurons_analysis_GO_AD_ADp40KO_UP.csv")

topGOResults.df <- read.csv(file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200618_subiculum_neurons_analysis_GO_AD_ADp40KO_UP.csv", row.names = 1)

datatable(topGOResults.df, options = list(pageLength = 5, dom = 'tip'))
```

```{r warning=FALSE, message=FALSE, fig.width= 15, fig.height= 7}
topGOResults_MF.df <- topGOResults.df[which(topGOResults.df$category == "MF" & topGOResults.df$Annotated < 1000 & topGOResults.df$Annotated > 30), ][c(1:15), ]
g1 <- scRNA_TopGO_plot2(topGOResults_MF.df)
g1 <- g1 + ggtitle("Molecular function")

topGOResults_BP.df <- topGOResults.df[which(topGOResults.df$category == "BP" & topGOResults.df$Annotated < 1000 & topGOResults.df$Annotated > 30), ][c(1:15), ]
g2 <- scRNA_TopGO_plot2(topGOResults_BP.df)
g2 <- g2 + ggtitle("Biological process")

grid.arrange(g1, g2, ncol = 2)
```

GO of AD vs ADp40KO down-regulated genes

```{r warning=FALSE, message=FALSE, fig.width= 15, fig.height= 5}
# genesOfInterest <- subi_DE_AD_ADp40KO_padj[subi_DE_AD_ADp40KO_padj$avg_logFC < 0, ]$X %>% as.character
# 
# topGOResults.df <- cluster_GO(genesOfInterest, geneUniverse)
# 
# write.csv(topGOResults.df, file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200618_subiculum_neurons_analysis_GO_AD_ADp40KO_DOWN.csv")

topGOResults.df <- read.csv(file = "/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/20200618_subiculum_neurons_analysis_GO_AD_ADp40KO_DOWN.csv", row.names = 1)

datatable(topGOResults.df, options = list(pageLength = 5, dom = 'tip'))
```

```{r warning=FALSE, message=FALSE, fig.width= 15, fig.height= 7}
topGOResults_MF.df <- topGOResults.df[which(topGOResults.df$category == "MF" & topGOResults.df$Annotated < 1000 & topGOResults.df$Annotated > 30), ][c(1:15), ]
g1 <- scRNA_TopGO_plot2(topGOResults_MF.df)
g1 <- g1 + ggtitle("Molecular function")

topGOResults_BP.df <- topGOResults.df[which(topGOResults.df$category == "BP" & topGOResults.df$Annotated < 1000 & topGOResults.df$Annotated > 30), ][c(1:15), ]
g2 <- scRNA_TopGO_plot2(topGOResults_BP.df)
g2 <- g2 + ggtitle("Biological process")

grid.arrange(g1, g2, ncol = 2)
```

### 4. DE by cluster

#### 4-1. DE cluster 11

```{r warning=FALSE, message=FALSE}
DE_AD_ADp40KO <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/DE_csv_files/20200117_Clustering_DE_by_MAST_each_Cluster_11ADp40KO_AD_corrected.csv")

# reverse logFC2
DE_AD_ADp40KO$avg_logFC <- DE_AD_ADp40KO$avg_logFC * -1

# only logFC > 0.1
DE_AD_ADp40KO_padj <- DE_AD_ADp40KO[abs(DE_AD_ADp40KO$avg_logFC) > 0.1, ]
# remove Ttr
DE_AD_ADp40KO_padj <- DE_AD_ADp40KO_padj[DE_AD_ADp40KO_padj$X != "Ttr", ]

DE_AD_ADp40KO_padj <- DE_AD_ADp40KO_padj[which(DE_AD_ADp40KO_padj$p_val_adj < 0.05), ]

datatable(DE_AD_ADp40KO_padj, options = list(pageLength = 5, dom = 'tip'))

```

#### 4-2. DE cluster 14

```{r warning=FALSE, message=FALSE}
DE_AD_ADp40KO <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/DE_csv_files/20200117_Clustering_DE_by_MAST_each_Cluster_14ADp40KO_AD_corrected.csv")

# reverse logFC2
DE_AD_ADp40KO$avg_logFC <- DE_AD_ADp40KO$avg_logFC * -1

# only logFC > 0.1
DE_AD_ADp40KO_padj <- DE_AD_ADp40KO[abs(DE_AD_ADp40KO$avg_logFC) > 0.1, ]
# remove Ttr
DE_AD_ADp40KO_padj <- DE_AD_ADp40KO_padj[DE_AD_ADp40KO_padj$X != "Ttr", ]

DE_AD_ADp40KO_padj <- DE_AD_ADp40KO_padj[which(DE_AD_ADp40KO_padj$p_val_adj < 0.05), ]

datatable(DE_AD_ADp40KO_padj, options = list(pageLength = 5, dom = 'tip'))

```

#### 4-3. DE cluster 15

```{r warning=FALSE, message=FALSE}
DE_AD_ADp40KO <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/DE_csv_files/20200117_Clustering_DE_by_MAST_each_Cluster_15ADp40KO_AD_corrected.csv")

# reverse logFC2
DE_AD_ADp40KO$avg_logFC <- DE_AD_ADp40KO$avg_logFC * -1

# only logFC > 0.1
DE_AD_ADp40KO_padj <- DE_AD_ADp40KO[abs(DE_AD_ADp40KO$avg_logFC) > 0.1, ]
# remove Ttr
DE_AD_ADp40KO_padj <- DE_AD_ADp40KO_padj[DE_AD_ADp40KO_padj$X != "Ttr", ]

DE_AD_ADp40KO_padj <- DE_AD_ADp40KO_padj[which(DE_AD_ADp40KO_padj$p_val_adj < 0.05), ]

datatable(DE_AD_ADp40KO_padj, options = list(pageLength = 5, dom = 'tip'))

```


#### 4-4. DE cluster 17

```{r warning=FALSE, message=FALSE}
DE_AD_ADp40KO <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/DE_csv_files/20200117_Clustering_DE_by_MAST_each_Cluster_17ADp40KO_AD_corrected.csv")

# reverse logFC2
DE_AD_ADp40KO$avg_logFC <- DE_AD_ADp40KO$avg_logFC * -1

# only logFC > 0.1
DE_AD_ADp40KO_padj <- DE_AD_ADp40KO[abs(DE_AD_ADp40KO$avg_logFC) > 0.1, ]
# remove Ttr
DE_AD_ADp40KO_padj <- DE_AD_ADp40KO_padj[DE_AD_ADp40KO_padj$X != "Ttr", ]

DE_AD_ADp40KO_padj <- DE_AD_ADp40KO_padj[which(DE_AD_ADp40KO_padj$p_val_adj < 0.05), ]

datatable(DE_AD_ADp40KO_padj, options = list(pageLength = 5, dom = 'tip'))

```

#### 4-5. DE cluster 19

```{r warning=FALSE, message=FALSE}
DE_AD_ADp40KO <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/DE_csv_files/20200117_Clustering_DE_by_MAST_each_Cluster_19ADp40KO_AD_corrected.csv")

# reverse logFC2
DE_AD_ADp40KO$avg_logFC <- DE_AD_ADp40KO$avg_logFC * -1

# only logFC > 0.1
DE_AD_ADp40KO_padj <- DE_AD_ADp40KO[abs(DE_AD_ADp40KO$avg_logFC) > 0.1, ]
# remove Ttr
DE_AD_ADp40KO_padj <- DE_AD_ADp40KO_padj[DE_AD_ADp40KO_padj$X != "Ttr", ]

DE_AD_ADp40KO_padj <- DE_AD_ADp40KO_padj[which(DE_AD_ADp40KO_padj$p_val_adj < 0.05), ]

datatable(DE_AD_ADp40KO_padj, options = list(pageLength = 5, dom = 'tip'))

```

#### 4-6. DE cluster 20

```{r warning=FALSE, message=FALSE}
DE_AD_ADp40KO <- read.csv("/data/rajewsky/home/skim/Microglia_Heppner/201912_10X_9sets/R_Scripts/DE_csv_files/20200117_Clustering_DE_by_MAST_each_Cluster_20ADp40KO_AD_corrected.csv")

# reverse logFC2
DE_AD_ADp40KO$avg_logFC <- DE_AD_ADp40KO$avg_logFC * -1

# only logFC > 0.1
DE_AD_ADp40KO_padj <- DE_AD_ADp40KO[abs(DE_AD_ADp40KO$avg_logFC) > 0.1, ]
# remove Ttr
DE_AD_ADp40KO_padj <- DE_AD_ADp40KO_padj[DE_AD_ADp40KO_padj$X != "Ttr", ]

DE_AD_ADp40KO_padj <- DE_AD_ADp40KO_padj[which(DE_AD_ADp40KO_padj$p_val_adj < 0.05), ]

datatable(DE_AD_ADp40KO_padj, options = list(pageLength = 5, dom = 'tip'))

```

```{r warning=FALSE, message=FALSE}
sessionInfo()
```








